\chapter*{\textsf{Twilight came}}
\addcontentsline{toc}{chapter}{Twilight came}

T\textsc{wilight} came.

``FALL IN!''

He filed out and fell in in his place. The assembling men clanked with equipment.

``All there?''

``Yes, Sir.''

``All right. Move on.''

Romarin was left behind, possessed by cooks and men on reserve. Shops and houses were shuttered and deserted. The profits had been carried to safe territory.

From a hundred camps by a thousand tracks or roads, and under cover of night, an attacking army was crowding forward to man the front line places of assembly. Up and down the roads could be heard the shuffle of feet on cobbles, the clank of rifles and shovels and water-bottles. And rapidly night advanced from twilight to darkness, conversation was lulled to a whisper, men grew sweaty under the burden they carried, sullen under the burden of their thoughts. Scattered guns fired random shots. They passed a huge howitzer into which sweating men who looked like fiends out of hell thrust heavy shell. Earth shook as the howitzer eructed, the fiends became pallid men in the blinding muzzle flash, the shell left as does a departing train.

They moved from roads to cross-country tracks, the sky glowering with the flashes of old guns, the shimmering of far-away flares. They passed gun-pits, hundreds of gun-pits, with guns trained at zero and the gunners asleep in the bivvies, gaining strength to feed a royal breakfast to hungry breeches. They passed piles and fields of shells with prepared fuses, so many shells that men marvelled that falling German shells missed any dump.

``Jesus, it's hot.''

``Jesus, it is.''

Night was so hot, air so still, that Guy started to sweat from the heat, from the effort necessary to carry his prodigious burden. Fortunately he was in good physical health, more likely to die of war than of the moist spot that had once wasted his lung. The moist spot had become a scar, and he had been preserved to be cannon fodder.

``Blast the bloody weight,'' someone cursed his load.

``A man's a human pack-horse.''

``Bloody Brass Hats half kill a man before he hops the bags. One hundred and twenty rounds is enough, but they give a man an extra bandolier.''

``Throw the extra one away.''

But no one threw the other one away, not yet.

``The more civilized war becomes, the worse it gets.''

``Meaning?''

``A man used to go into battle naked, carrying only a spear.''

Guy heard the voices and smiled at that idea. Yes, and the contenders used to eat one another, though who wanted to eat a lousy Hun? And then, proceeding from nakedness, he scowled at the night, while he saw civilization as a multiplication and mechanization of the arts of murder. Crashing howitzers' detonation and spurts of molten flame in the darkness gave grim significance to his conclusions.

``That's right!'' Joe said, ``that's right. The bloody artillery will start a bombardment and we'll get knocked out on the way up. Get hell before the show starts.''

``Yes. And when we want shells the bloody cows will be asleep.''

``The artillery have a jolly good time, parley voo,'' grunted a sweaty and exasperated soul from beneath his yoke.

A storm of incredibly vile swearing broke out. Someone in defiance of orders and common sense had lit a cigarette to enjoy the soothing influence of narcotic. Keen nostrils caught at the aroma, sharp eyes spotted the glowing cinder at the fag end. Everyone poured imprecation upon the offender.

``Knock the fag out of his lips!''

``Knock the bastard on the head!''

The air was too thick for the cinder to have been seen in the German lines, but the offender apologized abjectly and ground the fag under his heel. It was the last fag he was ever to smoke, anyway, but he didn't know that. The German lines were still a mile away. Low-flying planes throbbed overhead, so low-flying that they seemed to skim the tops of trees on Hill 63. The stumbling men could not tell whether the planes trying to make night give up its secrets were German or British.

``We can see them. They can't see us.''

``Sounds all right to me.''

``You don't think they have some means of seeing?''

``I think night-flying must be mostly bluff.''

Everywhere to right and left along saps and tracks men were moving forward, clanking with lethal mechanism. The pigeons on Guy's back seemed to croon once when movement was halted while guides attempted to make choice of branching routes. Guides always seemed to grow doubtful on important occasions. The company paused and started to curse the guides, but they were soon away from overland tracks and into trenches, hurrying around corners in pursuit of vanishing backs in front. Trench movement was like that. The head of the column got stretched out before the rear got moving, and the rearward men had almost to run to keep the back in front in sight, or were jerked to immobility by a sudden stoppage.

``What's that?''

``The throb of a motor.''

A tank lurched out of obscurity into sight, thrusting and waddling its way overland, rolling down wire, crushing out obstacles, charging and levelling down slender trees, rearing from the horizontal and leaning against the trunks until they snapped like brittle carrots, waddling away at a tangent into obscurity again. Tanks were ugly ducklings.

The company got lost again or was ahead of its schedule, and a whispered command came back for all to rest. They were in the open again, approaching the front overland, and if all had gone, quiet night sheltered too many hazards for anyone to want to rest outside of a trench. A premature explosion might litter the country with dead. When thousands of guns are laid on targets, when a hundred thousand men are moving to the ready, complete quietness is too sinister.

``Sooner we hop the bags the better.''

``Yes.''

``If Fritz got wind of this ------''

``If he did.''

The conversation came from the gloom, from anybody, belonged to everybody.

``All right,'' a voice ahead said. ``Move on again.''
