\chapter*{\textsf{War logic}}
\addcontentsline{toc}{chapter}{War logic}

E\textsc{taples} was cold. Frozen snow lay on its sands. Sixteen men 
packed in a bell-tent aided warmth if it did cause the lice to bridge 
from body to body. Etaples was too near England. Patriots encouraged 
school children to comb English forests for chestnuts. The patriots ate 
bread and the chestnuts were fed to hungry troops in base camps. Unshelled 
chestnuts were used to thicken stew, and the stew was dished up tannin 
blue and half solid. The stew got eaten because there was nothing else 
to eat. Nevertheless it was one thing to patriotically gather chestnuts 
and another to eat the stew. Officers and gentlemen wisely kept to cutlets 
and mashed potatoes.

John Guy thrust his bayonet into bags of straw. He learned to jab with the 
point at close quarters. He was taught the value of a rifle butt.

``When you get too close to use your bayonet kick your opponent in the 
stones.''

Men grinned and swallowed the sound advice.

``Don't be afraid of hitting below the belt or kicking below the belt. No 
Marquis of Queensberry rules. Kick him in the stones, the guts, the shins, 
or under the chin if he's lying on the ground getting ready to have another 
go. That's right, isn't it?''

Men grinned. That was right.

``If your bayonet sticks in his throat or guts put your foot on his face 
and pull it out. It doesn't matter what comes with it. War isn't marbles.''

Great advice, Sergeant.

``The only atrocity I know in war is letting the other fellow do it first. 
No time to take off your helmet and say `Beg pardon.' Isn't that right?''

All were convinced.

For how sound the instruction was. To shoot, to bayonet, to scatter a face 
with toe-plate and muddied hob-nails, to kick at the stones, above all to 
race the other fellow in doing all these. Playing for fun, rules were 
necessary. Playing for life, rules were twaddle, except where they hampered 
the other fellow. Chivalry was supposed to exist in a clean bullet and 
atrocity in a blunt-nosed one, but a soldier knew better. All was atrocity, 
and the game was to survive.

``Don't bother about prisoners. Dead men eat no rations.''

Sound advice that too, although prisoners were taken. Soldiers were more human 
than they were supposed to be. Humanity threw down the rules.

``A wounded man can shoot you in the back. He can't when he's dead.''

Logic, logic, irrefutable logic. Not terminology for politicians' orations, 
but logic. Bitter verbal pills for squeamish souls. What common sense was in 
Bull Ring! No time to glorify militarism, but time to make killing 
expeditious. The morality of the method had no place in instruction. 
What was most expedient to personal survival was what mattered. The 
Church justified the war. The Army's job was butchery, not philosophy. 
And time was short, so there was division of labour. The parson preached 
on the morals of war, the Bull Ring rang with the physics, the mechanics. 
The heart of the enemy was the target for the bayonet, and if the thrust 
at the fleshy heart interfered with the soul that could not be helped.