\chapter*{\textsf{Rehearsal}}
\addcontentsline{toc}{chapter}{Rehearsal}

S\textsc{oon} they were on the road to St. Omer, marching in column of three and singing, listening when the throat tired and the bellows puffed.

\begin{verse}
``Mademoiselle from Armenti\`{e}res, Parley voo, \\
Hadn't been loved for forty years, Parley voo.''
\end{verse}

And then to tell how Mademoiselle was at last loved, how Mademoiselle responded to infantry gallantry, how she was left with two babies for souvenirs, parley voo.

With vigour he helped to swell the chorus of the supreme marching song which never was static, for its lines and words were fitted readily to the topical obscenity, parley voo. For the song made the pack less heavy and the cobbles less blistering in the sweaty heat of midday. He sang the song as others wanted him to sing it, and was not without a power of improvisation. The infantryman revenged himself in the song on the Army Service Corp, the Artillery, the Military Police, the Medical Corp, the Brass Hats, who all had a jolly good time, parley voo, who suffered from all the diseases of the brothel, parley voo, who stole the privates' rum, parley voo. Verse with dynamite in its content eased all the aches of the foot-slogger's soul.

And what feat was worthy of chronicle in blistering verse, and likely to be relished more as it was sung along the roads than the story of Mademoiselle?

\begin{verse}
``Mademoiselle from Armenti\`{e}res
Two piccaninnies for souvenirs,
Inky pinky Parley voo.''
\end{verse}

Battalion was off to rehearse its stunt and the band could not play all the way. There was a story of a New Zealand Colonel who, on the hot sands of Egypt, did command the band to play without a pause.

``The band will play, all the way, and I hope it plays better than it played the other bloody day.''

But the Colonel had learned that quantity of music had no relationship to quality, and had never repeated the instruction, though its unintentional but happy rhyme made him thereafter famous. If the instruction added nothing to music it added much to gaiety.

It was Guy's first long route-march in France, a route-march from the rottenness of trenches to the greenness of pasture. He learned the songs his division had brought from Egypt, where aridity seems to have made vice fertile.

High explosive verse that had the tang of curry was a counter-irritant t the sweat and the lice. English, French, and Egyptian languages were corrupted. Grammatical purity was not as essential as corrosive obscenity. Men wanted to throw down civilian life and let all the world know they were in the army now. They sought to cure spiritual blisters homoeopathically. And the Division found a loud voice which it had lost in the Somme. Preparation had quickened the fever. Concentration had renewed hope. So Mademoiselle from Armenti\`{e}res became the whore of all foulness and blasphemy and the centre of caressing softness.

The rural setting of the march eased nerves if its cobbles wearied the feet. Sight of green wheat and grass was an emotional bath. Life beyond the thunder of the guns was still to be seen pursuing a constructive cycle. Fields were unpockmarked, houses white and unbattered, men and women laboured. Retiring to be rehearsed for slaughter, the Battalion had the illusion of escape. The moments of reprieve were crowded with uncharnel impressions.

Verses of songs were adapted to the environment. The battalion that sang the music-hall version was a battalion unblooded. Sex songs, doleful chants about death, sneers at officers and sergeants who pinched rum were great restoratives to the drooping spirit. Swear-words with the tang of the world's corners or the flavour of the world's sewers had a detonant quality that drove out dull care. But the world's gutters have always been scoured for camp-fire vocabularies. And Mademoiselle by the roadside, listening to singing men, didn't blush because a different convention prevailed in France.

To judge by profanity, except in very late afternoon, marching men were fierce, swaggering mortals, but obscenity was really a veneer, a protective coloration to tenderness. Guy's companions, hating to be caught out as humans, as fathers, as lovers, or sons, pretended to be bad men. They swaggered valiantly down the roads, keeping heart-burnings obscured. Thus it was that some of the most sentimental let their voice yell what was most crude and vile. They had most to hide. Rhymed obscenity had survival value, for it kept the civilian a soldier. The soldier could spit on what the civilian yearned for.

``Bum sore Madam,'' the language was corrupted as they bawled at the side-walk.

``\textit{Oui, oui,}'' innocent response to what appeared to be a friendly greeting.

But even obscenity could dry on the lips when the midday sun burned down at steel helmet. Sight of green wheat and climbing tendrils of hops and the white peasant cottages could not slake parched lips enough for song. Battalion would shuffle along, eager only for sight of cookers drawn in to the roadside and charged with hot tea. But if parched lips could not sing there were yet throaty and spicy expedients to coax another step or two. When the band was silent and songs were silent and conversation was no more, occasionally there came a voice from somewhere amid the column of threes, an irrepressible, stentorian voice.

``Have you heard the news?''

``What news?'' The whole Battalion would straighten its back to take the response.

``They say the squire's daughter has been foully murdered.''

``Foully murdered?'' questioningly from the hundreds.

``Yes. Foully murdered. Shot herself.''

``Shot herself?'' Still the query.

``No. S----- herself,'' would come the annihilating response.

And the pack grew lighter for a moment as the weight was eased by the lifting power of community profanity. Along the road for half a mile and the voice would come again, once more recuperating sagging backs and spirits.

``Ten thousand pounds I'll give away because it is my wedding-day.''

``Hooray! Hooray! Good fellow!'' from all the voices.

``On second thoughts I think it best, I'll put it in the old oak chest.''

``You---lousy---bastard!''

Storms of laughter at the old response that was always fresh. Detonant tomfoolery. Witches have ridden over the moon on broomsticks, so why should not soldiers forget the cobbles and roar aloud profane claptrap? Constructive profanity is the pinnacle of ranker virtue.

Guy knew the virtue of straightening tired back and marching past a band and the Colonel at the end of the day, trying to look as all tried to look, good for another ten miles. \textit{Esprit de corps.} Straightening, struggling to lift the feet high, to pass with erect head, to give a sharp salute, to do willingly at journey's end what one did grudgingly in camp at any part of the day, to do for the Battalion's sake because one was part of the Battalion and shared the honour of the Battalion.