\chapter*{\textsf{Talk, talk, talk}}
\addcontentsline{toc}{chapter}{Talk, talk, talk}

T\textsc{he} platoon was a strange amalgam. Men from cities jostled men used to crashing of big timber and the song of the circular saw. Many men concealed their past behind barriers of diffidence. A dead policeman might be replaced by an ex-prisoner. Those from good homes talked the loudest about home. The more obscure mortals refused to parade their past inequality. But past social standing was no longer as important as past experience of life. Why should men tell of themselves to their disadvantage?

A glass of wine lifted the curtain only on yesterday's glamour and not on its defeats. Poverty or wealth did not prevent a private from attaining great heights in dull hours if his had been richness of life. Bank balances deflected few shells. It was more important to have lived than to have husbanded when death gathered for the assault. Confessional of sorts is always the prelude to expected death, whether in bar-room or chapel.

So it was that John Guy, with an unequalled New Zealand experience of vagabondage, found his very lack of social respectability a title to a considerable appreciation.

``You make me want to go tramping when I get back.''

``When you get back. Why the optimism?''

``The open road seems to have been a long holiday.''

``Only in retrospect. There are times when the road is hungry, lonely and wet.''

``But the amount of New Zealand you saw. I'm dying for a country I've never had a look at.''

``A rolling stone gathers no moss.''

``But I worked in a shop all my life.''

``And I always longed to work in a shop.''

``When I get back------''

``If you do you'll want a feather bed and the shop again. You'll tire of sleeping out in France.''

``In the mud, but not in clean straw.''

``Feather beds are better than haystacks.''

``But it sounds as though you had a wonderful time.''

``Looking back it seems that way. But I was never happy about it then.''

Religious dispute was frequent. The disputes had no sectarian taint, for diversity of creed was welded to unity before the coarse satire of agnostic. Catholic, Presbyterian, theosophist, joined to rebut the atheist if they did separate to worship.

``God is on our side.''

``The Germans say that too.''

``The Germans are blasphemers.''

John Guy had an idea that the Pope and a few Quakers were the only Christians in the world, for they refused to allow God to be on any side.

War-time love got talked about in realistic language. The survivors from Gallipoli talked with relish of the obscenities of the Egyptian Waazer. The man new returned from leave discussed the whore he had slept with in London or Paris. The few who winced could draw aside. Sex, like drink or tobacco, drugged the senses for a moment.

New Zealand women were not brought into the conversation. They belonged to a departed world where soldiers had been fathers and husbands and sons and brothers. Women of home were the women of cosmos, women of the streets were the companions of chaos. Their attractions were as fleeting as life itself. Men being men, and not bowelless saints, were roisterers because only the brothel was obtainable. Choosing a woman of the streets was not disloyalty to home ties, it was merely the inevitable expression of a choiceless passion. Intimate home life, except among rare friends, was not discussed. The family life of each was sealed away. Men were better soldiers when they stopped thinking about home and discussed the merits of French or British and brown or white thighs.

``Why are soldiers expected to stop at killing?''

``What do you mean?''

``Is it only holy to break the commandments when the vices are unpleasant? The parson excuses killing but still baulks at fornication.''

``Only metaphorically. Down in the V.D. hospital------'' The words were drowned in laughter.

Poor soldier. When he erred most he gained least. What he bought as like as not evoked memory of what couldn't be bought. But the illusion of sex persisted so that he plunged again and again. In a grey world any woman was a purple patch, even at ten francs a time. Soldiers rarely had time or money enough to achieve disillusion. And boys are not ascetics. No wonder he heard the talk swing back again and yet again to sex.

``Mamselle! Mamselle! Mamselle!'' He joined his voice to the vociferating chorus as they found a seat.

Mamselle ran from table to table, anxious to oblige.

``The original Mademoiselle from Armenti\`{e}res,'' Joe grinned.

``Long enough in a war zone to be devastated territory,'' Guy replied. `` `And thou beside me singing in the wilderness.' '' He nodded at her stout form, and gave Joe a dose of his own medicine.

``Jesus,'' murmured Joe at the blasphemy.

Mamselle stood at his shoulder, and he planked down the ten-franc note and essayed again.

`` `Ah, my beloved, fill the cup that clears to-day of past regrets and future fears.' '' For he had once grown drunk on Omar too.

``What you want, digger?'' she said uncomprehendingly.

``Wine, the infidel.''

``Hurry, soldier, or I'll serve the others first.''

``Beer,'' Joe said. ``Beer. Beer. And wine to put punch in it.''

She went on her way.

``As Omar said, `Beer, beer, glorious beer.' ''

``Don't be a bloody fool.'' Joe took his beloved copy from his pocket and opened it on the table amid the stains.

Mamselle was forty and Mamselle was fat, and Mamselle was really Madam, but it pleased everyone to call her Mamselle. Beads of sweat stood on her brow, and her dress was moist under the armpits as she hurried to gather the francs.

``Hurry, Mamselle, hurry. The war might end.''

``Don't be so cruel to her,'' said Joe.

`` `She'll make the most of what we yet may spend before we too into the mud descend.' ''

``Oh, hell!'' Joe was growing annoyed.

One could be ironical about war and religion and women, but Omar had to be approached ecstatically. They drank a little from their glasses of beer and fortified the remnant with wine. They sipped and beat time with their feet to the organ.

``Omar's right. Woman, wine, song.''

``How do you know?''

``I don't. That's how I know he's right. I've sold tea, taught Sunday school, been pure all my life, and now I'm going to get killed. I was a simple Christian patriot.''

``And now?''

``I'm nothing. I don't believe in God any longer. I don't want to sell tea. I could kick myself for all my lost opportunities. I want to drink wine instead of cocoa.''

``It's never too late to mend.''

``I've done my worst, but what chance have I got? I don't want to die innocent.''

``Funny fellow,'' Guy laughed.

``I used to be a Tory and now I'm a Socialist.''

``Deat-bed repentance.''

`` `And lip to lip,' '' Joe was quoting, `` `it murmured while you live, Drink, for once dead you never can return.' ''

``Gently, brother, gently, pray.''

``All my life I've been in a mental cage and now I'm out of it, but I'm going to get killed before I get a chance to be a healthy sinner. I had to get into this bloody uniform to get out of my cage.''

``And all my life I've been in the open and I've been wanting to do most of the things you want to escape doing.''

``More beer, Mamselle. Come, fill the cup------''

``You surely don't want to be a civilian, to get drunk, to go whoring, to deny God, to stand on a soap-box?''

``I haven't got it that bad. But I want to be me. I want to lounge on a beach on Sunday instead of teaching Sunday school, I want to take a girl to a dance instead of to church. I want to walk up to a bar for a whisky and soda instead of running to a bun-fight. I want to live during this life and leave eternity to mind its own business.''

``You're drunk.''

``Not that lucky. Mamselle, more beer. And you, Guy, \textit{apr\`{e}s la guerre}?''

`` `The ball no question makes of ayes and noes,' '' Guy read seriously from Joe's open book.

``That's how you feel.''

``Joe, I was a dead-beat before the war. Who knows but I may continue to be one?''

``A dead-beat? You always sound as if you had an exceptional education.''

``I am a shrewd pretender. I was a vagabond most of the time.''

``A swagger?''

``Without a swag. No trade. No job as often as not, no home.''

``No bloody nothing.'' Joe's semi-drunken face grew broad.

``No. No bloody nothing.''

`` `Then to the rolling heaven itself I cried, asking what lamp had destiny to guide.' ''

Guy was getting too drunk to make fun of his friend's reading.

``More beer, Mamselle.''

``If you get back you'll be lucky.''

``Lucky?''

``You'll be free.''

``Free! Free to look for a job I won't be able to find.''

``But you've got a girl.'' Joe knew he had a reason for chastity. ``You must think a lot of her. She must think a lot of you.''

``Got! You mean had. I can only go home via Berlin, and it's a long way to Tipperary.''

``But if I get back I'll have all my people waiting for me to go back to church and thank God a bloody German didn't stick a bayonet in my ribs.''

``Never mind the afterwards. The war is young.''

``Yes. First ten years are the worst.''

``And they can do anything to a soldier except put him in the family way.''

They laughed at their use of the hackneyed phrases, drunk enough to roar at their own wit. But Joe was not drunk enough to desert his major intoxication.

`` `Unborn to-morrow and dead yesterday. Why fret about them if to-day be sweet?' Beer fills me with irresponsibility. I think it stops the lice from tickling. They get drunk on the sweat. And the war's only outside the door. Good fellows, all.'' He waved his glass at the crowded tables.

``They are good fellows.''

Both nodded because each meant the term ``good fellows'' to incorporate the speaker.

`` `One moment in annihilation's waste------' ''

``More beer, Mamselle.''

They fell to reading and quoting, each a listener to have audience for his own quotation. For the beer and the good fellowship renewed some of Guy's one-time worship too. They declaimed across the glass tops in confidential tones.

``Mamselle, more beer.''

`` `I often wonder what the vintners buy,' '' Joe started.

``For God's sake get the water Gunga Din,'' Guy finished.

``Good fellows.''

``Yes, good fellows, Joe.'' Almost they purred over the noble chaps they were.

`` `And when the angel with his darker draught draws up to thee, take that and do not shrink.' ''

``But if the angel is a Saxon with a sharp bayonet?''

``Oh, hell. You're growing impossible again.''

As Mamselle hurried to and fro with her beer, men discussed aloud her attractions as a bed-mate, her arms, her thighs, her over-weight bosom, the sweat that streamed as she hurried so that all the profit might be taken before the war should end. Voices hailed her with open suggestion, and the unfailing answer came back:

``\textit{Apr\`{e}s la guerre}.''

They could have their verbal licence if they accompanied it with orders for beer.

``More beer, Mamselle.''

``\textit{Oui, oui,} digger.''

Always Mamselle was behind, always she was driven along by fear of profits that would escape unless she was nimble. Conversation grew to pandemonium, rivalling the organ. Voices acquired loudness as stomachs knew beer.

``Say, aren't you listening?''

``Of course I am.''

``What lines was I reading?''

``Oh, hell!''

``How many francs to-night, Mamselle?'' a drunken voice boomed.

``\textit{Apr\`{e}s---la---guerre,} digger.''

Omar was gathered up, wiped of its drops of beer, returned to Joe's breast pocket. They set off to billets together arm-in-arm, moonlight, alcohol, snow, making the village restful. Chug! chug! chug! a far-away German machine-gun was going.

``There's a verse about the moonlight. How does it go? `Ah, moon of my delight------' '' He started to fumble for the book.

``For Jesus sake leave it where it is.''

``The best verse is about the woman in the wilderness.''

``You're drunk.''

By the door of the billet they paused.

``Say, Guy, I've a favour to ask of you.''

``As long as you don't want to borrow. I'm broke.''

``Will you write home for me if I get killed?''

``Sure.''

``Will you tell my people what a good fellow and all that rot I was?''

``Sure.''

``You don't think I'm foolish?''

``No. Only drunk. There's a verse about writing. `The moving finger writes, and having writ, moves on, nor all the piety or wit------' ''

``Let's go in.''

After Messines Guy was to write the letter. Joe was a good fellow, but the shell had no apologies to make on that score.