\chapter*{\textsf{Klink}}
\addcontentsline{toc}{chapter}{Klink}

H\textsc{e} lay on his back and surveyed the galvanized roof. The 
Sergeant of the guard came and yarned to him through the peep-hole 
in the cell. The Sergeant was a good fellow, convalescent from 
wounds.

``Hard luck, but don't worry. The first seven years are the worst.''

``Angley's a swine,'' Guy replied.

``I don't mind if you sing it. We all think that.''

``He's a bastard.''

Guy used the term not for its dictionary meaning, but because 
among New Zealanders no term expressed greater contempt. The 
dictionary gave the word meaning that was accidental. Guy saw 
Angley as a conscious seeker for the rotten.

``Take it with a grin, boy. You'll find a few like him in any 
division.''

``One too many, anyhow.''

``I used to think I'd loaf in Sling, but I'd rather hurry back 
to rotten conditions than be bossed by rotters.''

``You're a good fellow, Sarge.''

Left alone, Guy could have cried with helpless rage as he saw 
the case against himself fashioned out of triviality. He did not 
sense that his attitude rather than his crimes had been the fault. 
Into his thoughts came a voice from the next cell.

``Say, digger, how long?''

``Two days,'' he answered.

``Absent without leave?''

``No. Insubordination. Didn't have enough brains to shut up.''

``Man's a fool to argue.''

``That dirty b------ Angley.''

``Came here drunk the other night and pulled a fellow out of 
klink in his shirt-tail, and threw a bucket of cold water over 
him. Fellow's gone sick. Perhaps he's got pneumonia.''

``Why didn't the chap report him?''

``And get charged for malingering.''

They were the only two in the guard-room and were marshalled 
for strenuous pack drill. But the Corporal in charge wasn't a 
bad fellow and nursed them so that they survived even Angley, 
who did not arrive until the end of the day.

Guy's companion was a man with two wound stripes, a man recently 
discharged from hospital. He was an heroic militarist with a 
tendency to violence in his cups. Drunkenness was not an offence, 
but he had refused to stagger home quietly. Returning exhilarated, 
he had brushed against a man who was very much of a Lieutenant.

``Why don't you salute?''

``Go and play marbles.''

The Lieutenant had tried to arrest the drunk. The drunk had succeeded 
in uppercutting the officer and gentleman, who had acquired a black eye 
and a bleeding nose. Alas there were no marksmen's badges for target 
practice of that sort. And the offender had mauled a superior once 
before. So he was awaiting transfer to a military hell for a short time, 
a place where he would be drilled until he collapsed and kicked until 
he got up again. Military prisons had to be that bad to keep toughs 
in the trenches.

``Break your heart, they say, and break a few ribs to get at it.''

Bit exaggerated, thought Guy. In time he knew the statement for near truth.

When the day's drill was done they were relieved of their clothing and 
given two blankets and the board floor with which to defy freezing night. 
Neither slept. Guy's mate was used to the sheeted bed of a hospital. Guy 
turned and turned from sore side to sore side. They exchanged opinions 
through the night.

``By God, it's cold!''

``By God, you're right!''

The Sergeant of the guard gave them hot cocoa while the Corporal kept 
watch for Angley.

``I'll give you more blankets when that swine goes to bed.''

They lay with their teeth chattering, trying to pretend they were 
tough fellows.

``Are you asleep?'' from the next cell.

``No I'm drilling. Can't turn the right way.''

``Me too. Can't get into step.''

``We're both in step and out of step.'' He hazarded a joke.

``Cheer up. The first seven years are the worst.''

``All right boys?'' the Sergeant asked.

``Dreaming I was with a Flossie,'' said the man next door. 

``Jolly as a sand boy,'' chattered Guy.

``I think Angley's gone to bed. I'll give you more blankets.''

``You're a good sort, Sarge.''

In the morning Guy was sure he had a severe chill.

``I'm not well.''

``Report sick,'' came through the wall.

``Get hell if the Doctor says no. I'll go through with it for the rest 
of the day.''

They were drilled by an ambitious corporal who wanted to demonstrate 
his fitness for promotion, and they were kept near the window of the 
orderly room, near Angley's eye.

It was Quick March! Halt! About Turn! Left Wheel! Right Wheel! Pick 
up those feet! Left, left, left right left! Double! Halt! Mark Time! 
up and down, up and down the barrack square until the pack became the 
whole world, a burden that pressed the body down, that subjugated the 
spirit, that cut at the shoulders. They wheeled, moved, sweated until 
the frame was dried out, until the feet shuffled, until the mind reeled, 
until the ears ached from the torture of a raucous voice---voice of 
the inhuman corporal who marched to promotion over the bodies of his 
vanquished comrades. Around and about, around and about, until the 
mental chaos reigned and flesh grew pitiful. At last murmurs of protest.

``For Christ's sake ease up.''

``Give us a spell.''

The appeals only caused the Corporal to raise his voice higher so that 
Angley would not hear them, and so believe that the Corporal's squad was 
getting out of order. And Angley did bounce out to drill them that day.

``Ah, Corporal. Leave them to me.''

Angley would show the Corporal how he could apply the goad, humiliate the 
flesh. He would show what a whale he was to the Corporal's minnow. Worms 
are nothing to whales, but heavy diet to minnows.

``What a nice pair of soldiers you have here, Corporal. Think they can get 
past the army, but they can't even get past me. Why don't you smarten them 
up? I'll show you how. Pick up those feet, up, up, up. Don't slouch along. 
Hold up those feet, up, up, up. Hold up that head. Tired? What bloddy right 
have you to be tired? You will get tired only when you're told to get 
tired. Tired by numbers, ho, ho, ho, ho. Sleepy, sleepy, I'll wake you up. 
Pick up those feet. This is the army, not the Salvation Army. About Turn! 
Save the Empire with rotters like you. God help the Empire. Double! Time 
you washed your ears and listened. About---ah, got you. What are you turning 
for? About time you washed your ears. About Turn! What the hell did you 
wait for?''

Around the square they were driven under his jeering, rat-eyed surveillance. 
Driven until they desired to jab him in the face with a gun, with a fist, 
with a boot.

``Double! Double! Double! We can do anything except put you in the family 
way, though you look like a pair of sissies. Double in front, Corporal, to 
show them how it's done.''

The Corporal was fresh and unencumbered by a pack, and sought to excel. 
Angley manoeuvred them into a quiet corner of the square where he could 
indulge in his passion for obscenity. Into their ears poured a stream of 
filth. Up and down they drilled, until Guy's mate, green from hospital, 
stumbled, fell, erected himself, decided.

``I've had enough.''

``You filthy rat. That's mutiny.''

``Me too,'' said Guy.

``Mutiny, I say.''

``I'm tired of being called a bastard,'' said the man with two wound stripes.

``Double mark time, you curs!'' Angley's face was congested with wrath.

``Can't be done.''

``You'll double when you come from jail. They tame lions in those places.''

``We eat 'em in Wanganui.''

Guy felt a thrill of pride at the desperation of his mate. This surely was 
a brave fellow.

``I'll fix you,'' Angley said. ``I've got a friend where you are going. I'll 
tell him to make it hot. When you come back you'll lick my feet.''

That threat to his mate cost Guy his control.

``That's what a dirty rat like you would do.''

Guy lapsed into silence, afraid of what he had said as he said it. He could 
have bitten out the tongue that had been so foolish. Angley frowned but 
dropped the matter. He had been indiscreet in his own utterance. He had a 
more satisfactory way of settling these affairs than dragging them before 
high commands.Out of his congested face he spoke with a tone that was 
suddenly subdued.

``Your say now. Corporal,'' looking at his wrist-watch, ``take over the squad. 
My turn next.''

Thrilled at Angley's retirement, which puzzled the Corporal, the squad turned 
its vocal chords on that worthy.

``You rat,'' said the man from Wanganui. ``I suppose you sleep with Angley.''

They slouched across the barrack square until sight of the Colonel spurred 
themselves and the Corporal to heroic effort. The victims and their tormentor 
formed an entente before the mightier power.

In the background, under the veranda, Angley stood grinning maliciously. He 
was a ``filthy rat,'' was he? He'd show 'em how to swallow words before they 
were much older.