\chapter*{\textsf{Sling}}
\addcontentsline{toc}{chapter}{Sling}

J\textsc{ohn} G\textsc{uy} could feel the rolling of the sea as he stood on parade on Salisbury Plain. 
The effect of the weeks of voyaging took more than a night to sleep off. He turned 
to whisper to his comrade, but that worthy beat him to the sentence.

``I can still feel her rolling.''

``Me too,'' he replied.

But even a careless breath would have excited the Sergeant-Major in front.

``SILENCE!''

What a snarl was in that voice to soldiers from New Zealand, soldiers used to the tone and method of
volunteer R.S.M.s and officers. In New Zealand militarism had been endured for war's sake, on 
Salisbury Plain it was worshipped for its own. In New Zealand men were civilians who 
had volunteered. In Sling they were raw material to feed to an impersonal machine. Late 
the previous night the reinforcement had stumbled into camp to find the civilian bond 
which welded officer to man severed. No longer were officer and man equal adventurers. 
At once the reinforcement became Other Ranks, automata. The officers attained an 
unapproachable elevation. The civilian army had to be put to death and out of it 
automata had to be resurrected.

Some few men shuffled.

``Silence! Still, I say!'' The voice had a wolfish ferocity.

Curious eyes watched the sentry on guard duty at the orderly room as he sloped 
his rifle. The three exaggerated movements rang across the barrack square.

Click! Click! Click!

The Sentry turned left.

Click! Click!

Marched five paces and halted.

Click!

Turned front again.

Click! Click!

Ordered his rifle.

Click! Click! Click!

Stood at ease.

Click!

Gazed blankly at the assembly as though it was not a reinforcement from his 
homeland containing friends. He was a piece of military furniture. Intelligent 
man was gone and automatic mercenary reigned.

``Jesus!'' Guy breathed.

``If he shuts his eyes they'll click,'' whispered a neighbour.

``If 'e's a soldier I'm not ambitious,'' another breathed.

``SILENCE! I SAY!''

The way the R.S.M. said ``silence'' made the whisperers feel guilty. He seemed 
to have the stealthy faculty of sensing sounds he couldn't hear. Maybe the 
word was pretence. The threat of the voice was profoundly bestial.

``I hate the bastard already,'' Guy whispered.

Near the sentry stood an officer. He stood as a human being would stand, 
relaxed, indifferent, bored, swinging a cane, apparently unconscious of 
the sentry's heel-clicking. He was obviously great. One of those 
privileged almost to remain a human being. He straightened himself, 
limped forward.

``Parade, SHUN!'' the R.S.M. roared.

The great man spoke.

``Men, I welcome you to camp. You are no longer civilians in uniform. 
Here you will find out you are soldiers in training. We give orders 
once only. When men disobey, correction is swift and severe.''

``God save the King,'' whispered John Guy's associate.

John Guy squirmed under the unnecessary threat. He hated militarism and 
was committed to it body and soul. He loved freedom and open spaces 
and had enlisted in the spirit of adventure. His soul was at ease 
beside the crash of falling water, the song of birds, the hushing 
of dense forest. And the more they tried to make him automaton 
the more he became individual. He was dark, slender, dreamy, 
unsoldierly. Hostility at the threat in the suave voice welled 
up in him. But he knew that the power behind that voice was 
a juggernaut that crushed human foibles.

``There will be London leave for those who can afford it.''

A quiver of appreciation ran through the ranks. The R.S.M. caught 
someone talking.

``Stop his leave,'' said the great potentate.

``Yes Sir.''

``When leave is over there will be an end of slacking. March away 
the parade.''

Platoon by platoon they returned from the barrack square while 
the commanding officer limped away.

``What's the good of leave to me?''

``What's up, Jack?'' Robbie of the golden smile asked.

``Broke. You know I lost my wages for the trouble in Capetown.''

``I'll send round the hat.''

``No. No.''

``But you must.''

``Well, only my train fare.''

So after being a penniless vagabond in New Zealand he was going 
to London to be a penniless deadbeat in the heart of the Empire. 
From the ends of the earth he had come to make a bed of embankment 
granite. Macaulay's New Zealander philosophically contemplated 
the ruin of London, but as a lesser New Zealander he would sneak 
under the bridge to sleep. Vaingloriously he sang:



\begin{verse}
``Soon we'll be in London Town.

\ \ Sing my lads, yo ho.

\ \ See the King and the Golden Crown.

\ \ Sing, my lads, yo ho.''
\end{verse}

``You are featherless but happy,'' Robbie averred.

``I intend to paint the town red,'' adding under his breath, 
``on four bob.''

``Say, Bill,'' yelled a voice, retailing an old chestnut.

``Who keeps the pub in London Town?''

The Orderly Sergeant put his head in the hutment door.

``Men going on leave will attend compulsory lecture on 
Venereal Disease.''

``Is London that bad?'' a New Zealander sneered at an 
immigrant Cockney.

``For sissy boys,'' the Londoner turned New Zealander 
replied.

``Boys,'' said John Guy, ``vice has got to be made safe for 
democracy.''

The Colonel was notorious for his Rabelaisian lectures. He 
surrounded sex with such obscene glamour that men went 
more than once to laugh uproariously. His crude zest 
would have shocked the sensitive if their diffidence and 
disgust had not been exploded by the sheer humour of his 
brothel stories. Officers and men attended; a padre on 
the prominence of the platform was the butt of asides that 
rocked the house. The Colonel did not forget it was a 
military lecture and fitted the lore of the brothel 
into martial terminology. He was not concerned about the 
morals. He took it for granted that all intended to err, 
and he only aimed at preventing error from impeding 
military efficiency.

Probably the glib Colonel attained a success that would 
have been beyond the power of a delicately tuned soul. 
Ribaldry was of the soldier. Humility had been sloughed 
off with civvy clothing. The audience felt they listened 
to a man of the world, strong enough to sin, shrewd enough 
to avoid the consequences. The audience, boys mostly, 
would heed the apparent rake but ignore the prude. The 
Colonel knew that ``Don't'' was futile. The imperatives 
of home were no more. He made vice alluring to secure 
their goodwill, and then used their goodwill to make vice 
safe. He cheapened sex to the commonplace so that boys 
would not conceal their ailments but hasten for cure. There 
was a war on. Militarists were not concerned about morality 
or immorality. They were concerned about the military attribute 
called health. It was best to condone lust and make it safe. 
Bodies were cannon fodder only if physically fit. A diseased 
soldier wasn't as good a target for high explosive as a fit 
soldier. Disability by a bullet was a military risk, by a 
woman unscientific management.

John Guy walked from the lecture-room with sides aching from 
laughter. He would have liked to have shown disgust, but 
the crude impact of droll stories had swept away any diffidence. 
He wondered if he was in mental harmony with the lecturer and 
shuddered. But as he shuddered the military terminology came 
to his mind again and he roared aloud. Birth and life and 
death were laughable in war.

``Ho! ho! Ho! ho!''

``The Colonel's a good sport.''

``Didn't he rub it into the parson?''

The reinforcement paraded for pay, for railway tickets, for 
a box of prophylaxis.

``Not me,'' said John Guy.

``Goody, goody,'' muttered the Quartermaster.

``I'm bloody well broke,'' Guy answered.

``Hard luck, old man.''

But John Guy did not want to look for Flossies. Poverty was 
only his excuse so as not to seem puritanical. He had a chaste 
love for a fair-haired girl he had left sobbing in Wellington, 
New Zealand. Suddenly, however, he was curious. He wanted to 
see the contents of the box.

``Give me one.''

``That's right. You never know. An emergency ration.''

He was not a prude but merely a boy in love. Absence made the 
heart grow fonder, temporarily, he was to find out. For absence 
and despair could at last strangle love to death.

Four shillings did not buy much red paint. The most ardent 
sinner can be a saint on four shillings even in peace time. 
War time brought a horde of swarthy strangers to London and the 
flesh market appreciated. Supply and demand economist's describe 
the process.

Out of tall Southern tussocks he came to doze in London parks, 
to sleep by the lapping Thames. In four days and nights he saw 
more of London than he would have in four years with cash in 
his pocket. He communed with the shuffling derelicts of night. 
With a sordid wretch he stood in the shadow of a mansion and saw 
elaborately gowned wives, and mistresses of pillars of Empire, 
hurrying from taxi. Wealthy London was wallowing in its war 
profits.

``Good night, New Zealander.'' How many girls came to sell 
themselves? Blessed are the poor for theirs is but temptation.

``No! No! No! No!''

``Come on. I've got a nice room. A soldier wants a bit of fun.''

``I'm broke.''

``Oh, sorry. Well, better luck next time.''

Girls didn't argue once he disclosed his financial state. War 
had brought rich harvest up to London Town. Sweethearts and husbands 
were far from lovers and wives, and vice was made safe for democracy. 
When the young had gone with the sturdy the blear-eyed hags had a 
chance to scavenge for the remnants.

``Good night, soldier.''

``Jesus!'' He recoiled from many a one with that word on his lips.

``Listen, soldier------''

``Let me alone.''

``But------''

``Go to hell!''

Perhaps he wondered whether profiteer mansion or frowzy hag represented 
the Empire he had come to fight for. With the sky for roof he saw a Zeppelin 
shot down at Potter's Bar. Unencumbered with wealth he saw the London social 
gutter, deeper and wider than the Thames itself. He was glad to return to 
camp for a hot meal and warm sleep.

``Had a good holiday, Guy?''

``Marvellous.''

``See the Tower?''

``No.''

``The museum?''

``No.''

``Go home with a girl?''

``No.''

``The Bing Boys?''

``No.''

``What the hell did you see?''

``Nothing much.''

``Some fellows come across the world and see nothing.''

``That's me. Born blind. Only got a whiff of the sewer.''

To an enthusiastic militarist Sling was bitter enough. To an outlaw in 
khaki Sling was hell. Wounded men rejoiced to be redrafted after a spell 
in Sling, and camp had to be hell to have that effect. In Sling patriotic 
illusions were rolled in mud. Militarism was using war to dig in. The 
amateur soldier could not be permitted to win the war. More 
important that war be used to make the civilian seem a professional 
soldier.

Sling had its beauty as well as its bestiality. Rolling Salisbury Plain 
had historic landmarks that the pedestrian could reach when the war god 
relaxed. The deciduous plantations looked as old as British civilization 
though younger than primeval New Zealand forest. The whole world was different 
from Down Under. Moist, hazy air blended all into one picture, whereas in 
full-toned New Zealand landscape was a multitude. Smoke rose from a 
chimney and lost itself in the haze where on a still New Zealand day it 
might ascend in a straight pillar to heaven. Used to open air and skilled 
at appreciating contrasts, England's special charm delighted the outlaw. 
Snow came and fog came, frost came and made the fog a crystal icing on 
tree tendril, so that the winter to an inhabitant of the subtropical 
became fairyland. But the fairyland was peopled with evil spirits who 
wanted John Guy, individual, to become John Guy, automaton.

After the sea voyage drill was a tough hardship but accepted without protest. 
Authority in the form of a swaggering bully was not as easy to swallow. If 
the body toughened the spirit revolted. The outlaw wanted to use the butt of 
his gun on the bully's nose. He didn't. He had grown wise on past defeats. 
When officers on the road to France were in charge drill was a challenge to 
excel. Partners in adventure demanded what they gave themselves. When the 
bully swaggered he gruelled others for self-glorification and not to win the 
war.

John Guy should never have been a soldier. He hated regimentation. Because 
his countenance reflected the prejudices of his soul, regimentation soon 
hated him. Not that he was sulky or inefficient. He was able but could not 
dissemble his militant hostility at what he considered injustice. He rebelled 
against the attempt to fit his clay to a common mould. He had qualities that 
might be of service in France, if he got that far. For the R.S.M. who dominated 
the camp John Guy conceived a hatred akin to that an animal feels for a trap. 
When the R.S.M. snarled a pugilistic something possessed Guy's jaw and fists. 
He hated to hear the inefficient bated. Fool that he was, he thought that a 
soldier still had a right to be a human.

What grist war brought to the camps from frozen south and northland, from tropic 
and sub-tropic. White, yellow, red, black, all offered their bodies to a machine 
that was not concerned with individuality. With such a mass who was concerned 
about capricious nonconformists? If a vagabond became a soldier he must accept 
the whole yoke, else vagabond and chief bully were bound to clash. They magnetic 
attractions. John Guy was to feel the heel of militarism on his carcase at an 
early date.

``Look out, Guy, you are taking risks.'' It was Robbie who advised.

``Do you expect me to laugh at that bastard's jokes?''

``Try anyway. Or do what I do when in doubt.''

``What's that?''

Robbie smiled. The smile of a radiant boy on the verge of handsome manhood. The 
smile had the earthiness of life that expanded in the sunshine. When Robbie smiled 
shadows vanished. He broke up bitterness as the sun breaks up mist.

``You always seem to reflect ecstasy when you smile, Robbie.''

``Funny. I smile automatically when I'm in trouble. Folk think I'm smiling at 
them and don't press me too severely.''

``Really?''

``I know it's funny. I always smile when I'm ready to cry.''

``And keep others happy. But surely not only then, Robbie?''

``Not always. I shouldn't have told you. You'll get to be able to distinguish 
the different sorts.''