\chapter*{\textsf{Gas}}
\addcontentsline{toc}{chapter}{Gas}

T\textsc{hey} stood up and moved on again. A heavy fog had come down and the shimmer of flares could no longer be seen. The few shells that wobbled wearily and high were bound for a long way back. But if they imagined night was going to be generous all the way, they were to know disillusion. A peculiar soft, moist scream came out of the air, an impact.

PHUT!---a soft impact.

``Gas!'' a dozen voices said.

Phut! phut! phut!---away in front.

``Gas!'' voices reiterated.

Phut! phut! phut!---to the right.

``Gas all right.''

Phut! phut! phut!

The air around started to rain shell possessed of the liquid impact. Fritz was drenching the country. A faint aroma was borne to nostrils, the aroma of gas diluted by the fog. No other shell had that wobbling scream, none other that soft impact, and there was no doubt about the odour.

``Gas. Put on helmets.'' The Gas Corporal made the decision.

John Guy adjusted his helmet and stumbled along in the gloom, chasing his predecessor's back. To sit in peace dressed in a helmet was bad enough, to be compelled to adjust a helmet and struggle on when hot night and a heavy load made breathing laboured, was to stifle. The lungs wanted much air, and they were only permitted to have a little. The nose-clip of his respirator pressed heavily on his nostrils, and the eye-pieces were dimmed by the warmth of his face, by the fog which gathered on the discs. He could not suck in sufficient air, and expulsion was difficult.

Fighting for air was enough of a task without pressing along with a bloody ton on his back. And the weight grew as the breathing was restricted. Still, he had to go on, to follow that blasted back in front, to keep out of the way of the man in the rear, who also had his difficulties. On, on, laboriously on to the trenches which had to be occupied at zero, on and through and beyond the gas. Through his dim eye-pieces he saw a figure coughing and spluttering, turning back, throwing down all his harness as he hurried to the rear, tearing away from his face a mask that had failed, or that had not been carefully adjusted at the crucial moment. For a thoughtless instant Guy was sorry that his mask had not also failed so that he might let the cooler and foggy air wash against his face. Fresh air could be worth risking even pain for.

He thought of his basket of pigeons and wondered how they fared. But there was a touch of selfishness in his thoughts. Were the birds lying stricken, and was he carrying into battle two useless birds? Should he lighten his load? Dead birds were no bloody good. Why should birds be carried while he could hardly fight along himself?

On, pushing on, over a blasted duckwalk that was a snare for unencumbered feet and eyes even in daylight, a duckwalk which, now that men wore masks and were heavily weighted, seemed to trip the feet at every step. On, on, on, on with the soft shells going phut! phut! phut! but no linger alongside. On and on and the steel spring of his bloody nose-clip was so strong that it would cut the bloody end off his bloody nose. On, on, the respirator growing hotter and warming the air he inhaled, on, on. Jesus, he would strangle, be gassed by his own stale breath. His mind was full of the expletive bloody. Bloody nose, bloody nose-clip, bloody hot air, bloody duckwalk, bloody opaque eye-pieces, bloody fool walking about with a bloody sack on his bloody head.

Oh for a few breaths of cool air, the air that had seemed only a few minutes before so hot and oppressive. The moisture blinded his eyes so that he scarcely saw the figure in front that started to gasp and splutter, that discarded equipment, that gulped down lethal air that seemed so cool and refreshing when it did not come via a helmet. He bumped past the figure as it turned back.

``Hospital for me.'' There was exultation rather that agony in the voice. Actually it was the excitation of the night, the incredible burden being carried, the necessity of getting on that crushed. In ordinary trench-kit and reclined against something, the difficulties would have been made light of. But minutes of enforced activity without adequate air stifled more surely than hours of immobility. Damn the bloody nose-clip. He would relieve his nose from the moth-piece. He relaxed his nose and unwittingly, from his nostrils, filled his helmet with moist, stale air and lost his sense of direction completely. The moisture absolutely put out his eyes. His feet hesitated and his follower stumbled against him and fell to his knees. He heard a gurgled protest from the rear helmet. Why the hell had he liberated his nose in defiance of the rules? Damn, damn! He was a dirty, bloody idiot.

He couldn't see his way and he couldn't feel his way, his hands and arms had to control his equipment. And he had to keep moving. And if he wasn't gassed he'd soon be killed anyhow. Why should he choke in a bloody sack? Why should he suffocate because a bag was tied round his bloody head? When the trench turned, for they were back in the trenches, he ran straight against the revetting and knocked the remaining wind out of himself. Bump! Someone floundered on his heels again.

``Get out of the bloody road.'' He was sure that was what a voice was trying to say as it sucked at a mouthpiece.

He made a decision.

``Here's where I become a bloody casualty,'' he said, spitting out the respirator and tearing the helmet from his face.

Cool, foggy air that once had seemed so hot fell against his face. His lungs gulped at the refreshing coolness. What a miraculous quality cool air was. The odour was faint, might be imaginary, the soft phut! phut! phut! was away to the rear. He would continue to take risks in the open rather than die like a cat in a bloody bag that he had tied around his own head. He would take the chance. And was so easy once one's lungs knew fresh air. The progress of the helmeted was halting, pitiful, if it seemed tremendous to them. And nothing happened except that cool air washed against the streaming sweat of his face.

In front and in rear were stumbling, swinish figures. For the masks on the faces had destroyed all the humanity. The figures were bulged with mechanism from head to foot and the face no longer remained to recall the human. For the face was blotted out, the nose was clipped by a swinish snout, and the eye discs sere without felling. The rubber tube that hung from the snout was half a trunk. And the shambling men clanked and clinked and creaked. The god of war had fashioned men in his own image, swinish snout, cuttlefish eye-pieces, bestial trunks, figures that groped and ambled like blinded bears. Guy had plenty of time to note and meditate, for he breathed at will and saw where he was going.

``When will the alarm be declared off?'' a masked voice gurgled.

``I've declared it off already.''

``I'm bloody well with you.''

So another head emerged and another sweaty face gulped toward cool air. head after head of the more incautious appeared and carefully tested the air with the nostrils.

``Fog must be washing the gas out of the air.''

``Suppose so.''

``Helmets off.''

The timid were pleased when the order came back.

``You took yours off too soon.''

``That bloody broken nose of mine! The clip gives it hell. The bones are not plumb. At the gas school they told me to go to hospital and get the bone straightened or removed.''

``Why didn't you?''

``Too many bush carpenters and vets in the Medical Corps, but hospital would be all right until after Messines.''

``Too late now.''

``Fritz will operate when we jump the bags.''

``Yes. Without chloroform.''

``Quiet,'' the man in front whispered back. ``Quiet. We're near the front line.''