\chapter*{\textsf{Nerves, nerves, nerves}}
\addcontentsline{toc}{chapter}{Nerves, nerves, nerves}

A\textsc{ll} through the night men crept back to the Club's dormitories. Some came with the grey dawn and the milkman. At breakfast-time fully half the beds were full. The missing fifty per cent were wisely greeting dawn and breakfast from the soft pillows of double-beds, reaching out arms to caress fair bed-mates.

Guy awakened early and lay looking at the ceiling and sensuously writhing his body against the unaccustomed warm smoothness of clean sheets. He appreciated the luxurious resiliency of a mattress, the lice-free underclothing, the silky friction of a pillow against a cheek still abnormally smooth. That pillow caressed deliciously after the course fibre of his pack with its buckles and lumpy assortment. He thought of the breakfast he would like to order. He would like to emulate the Colonel. Two eggs soft-boiled, a plate of buttered toast cut in long and plenteous fingers, a jug of coffee rich with sugar and milk. And all served by the tall fair creature who had caught his eye in the previous evening. ``Ah!'' he could sniff appreciatively at imagined aroma. Of course he didn't have to bother about breakfast. He couldn't afford that luxury. He was facing day and London leave with eightpence. He was face to face with grim hazards, and his nerves were playing him the childish tricks all nerves play before zero.

In the next bed was the sleeping face of the last man to come home to bed, a man heavily and audibly asleep. Guy decided that since he must wait on the goodwill of the pay-office there was no urgency about getting up. He might as well try to sleep again. He had paid for the bed.

He could not rest without challenge. When he awakened a second time sheer joy of bed's cleanliness caused him to rid himself of his singlet so that he could know joy of sheets against a naked body. He twisted from side to side to revel in the electric tingle of the chafing, as though he could brush the quality he felt out of the bed and into his blood and sinew, as a brush on a wheel extracts electricity from falling water and converts it to power.

``Christ, I can't say I saw much of London yesterday.'' The voice was that of a soldier who had staggered or who had been brought back the night before. ``I got into a pub, and that's all I remember.''

Guy heard someone answer.

``Man's a fool to get drunk. Picks up any old tart. Forgets the dreadnought. Ends up behind barbed wire.''

A very early riser who had gone to bed, a superior soul above the allurements of women and liquor, looked upon the dormitory and listened to the conversation with obvious disgust, a saintly pilgrim nauseated by the swinishness of the herd. Guy revelled in physical purity if his mind was impure. He ran his hands over his lice-free legs, over his smooth,  sweatless chest, and sighed with a rich animal contentment, sighed as a surfeited cat might purr. He thought of his body of the previous evening, grimy, lousy, encased in sour clothing, he thought of a chin decently shaved that in France was regularly agonized by a dull razor-blade. His body had gone at a bound from trench rottenness, from the sound of the guns, from a dusty or muddy couch, to a bed clean and springy. But his soul had not escaped, although he forgot about that for the moment. No one shoved a boot against his ribs and said ``Get up. Your turn on the fire-step.'' Half the world murdered the other half, but he reclined in clean sheets.

His spirit exulted and he was overcome by a tidal wave of joy. He was a pagan adoring a clean skin. Men could stand on platforms and be eloquent about hatred, but he wanted only to run down Southampton Row naked, shouting ``Clean! clean! clean!'' lording his purity. Yet if he did that his body would be abandoned by Mars and surrendered to a madhouse, for his fanaticism would be unconventional. Every nerve within him was a separate entity, enthusiastic and vain about his noble cleanliness. Soap, clean water, sheets, respite from concussion, these simple amenities were the great things of life. His body was free. Alas, his nerves and his soul were in chains. The gladness in his heart had issue suddenly and unexpectedly in copious tears. Tears ran down his cheeks, then a sob of a child racked his body. The tears gave easement, were refreshing like a spring rain on parched pasture. These wellings up, these unaccountable and uncontrollable sobbings that had periodic return seemed in themselves to yield a relief from some intolerable inward stress. He was clean and free, but had tears in his eyes because the man had disintegrated amid the thunder of the guns.

``Say, what's the matter, old man?''

The voice of the soldier who had been drunk came to his ears, and he remembered that he was sobbing in a dormitory, not in solitude. That did not mean he could control himself. It did mean that as he regarded the drunk he disregarded himself a little, and the tears lessened.

``Don't know. I've been doing that a bit lately. First time I've been caught.''

``Well, I do know. I've seen that before. Your nerves are gone. Go and see a quack.''

``Leave is too precious.''

``You might get a long leave.''

``Report that I have been bloody well crying.''

``It does sound a bit funny.''

Aches, pains, bloody wounds, trench feet, all these he could explain to a quack. Of all these one would have discernible evidence, but if he went to see a quack he wouldn't be able to weep to order. That's just the time he would laugh, and the quack would laugh at him too. If the quack was a good fellow. If he wasn't he might charge him with malingering. He was suffering from too much trench, too much war, too much concussion. But so was everyone. How could he cause a quack to make head or tail of that? And his holiday would set him up. There were golden days and nights before he donned his verminous rags if only he could vanquish the pay-office. Against the rules and with the aid of bribery, the drunk of the night before brought him coffee and food to the bedside. His tears had been of strategic value.

``Now buck up. I can't stay. My head's big. I've got to get a booze somehow before the pubs open. When they do have a stiff one yourself. It'll buck you up.''

Guy was profoundly grateful as he thought of his circumstances. He swallowed the food and drained the coffee to the last drop, and jumping out of bed and garbing himself in his raincoat, he descended into the regions of the the shower-baths. He left his coat and wallet on the bed until he returned. He was poor enough to disregard the printed notice that warned him to be careful for his valuables. Perhaps some crook would note his poverty and add a few bob. He smiled with self-satisfaction at that impossible idea. Under the shower he sent streams of water down his body, laving himself from head to foot with soap, more to be luxurious than to be clean, for he had been scrupulous the evening before. Other men were amusing themselves in the same manner, sending gutters of soap bubbles down to the drain. They beamed and sang as they soaped and laved and as the filth and sour skin was rolled away. A shower-bath with its frictional pelt and flow of drops was a miracle after the slimy sousing of the Point de Neippe vats. There the very water in its slimy alkalinity at last reeked with sweat and trench odours.

Soldiers, like boys in ponds in summer, were grown body proud, applied friction until the dirty sallow hide flushed athletic pink. Cleanliness was godliness. The ranker under the shower became God.

``Oh, madame, have you any good wine, Parley Voo?'' one sang.

John Guy rubbed himself down thoroughly. The night before he had been able to rub off little rolls of dirty skin. Now he was clean. Ready for the attack on the pay-office, and then------He sang his thoughts.

``The infantry have a jolly good time, Parley Voo.''