\chapter*{\textsf{Torches and meteors}}
\addcontentsline{toc}{chapter}{Torches and meteors}

A\textsc{bove} Romarin the air was always throbbing with planes. Like huge dragon-flies they soared and circled and stung and spat at other planes that swooped down out of the ether to sting and spit. Engines throbbed and droned and buzzed, planes were driven down by tracer bullets shot from enemy planes, bullets that threaded a silver streak in the twilight sky. Planes were driven up into the blue away from the angry burst of aircraft shrapnel, planes flew low, defying point-blank fire of machine-gun and hundreds of trench rifles, to map out for enemy artillery. Romarin was a hive of activity that radiated toward the German line, so over Romarin hung the eyes of two armies. Mankind climbing in the air caused mankind to grovel and burrow feverishly deeper into the safety of mud.

And for a terrible moment Fritz was Emperor of the Romarin heavens, and set those heavens aflame with the burning bodies of his enemies, illuminating his kingship. Above Romarin a captive balloon hung and feasted its greedy eyes upon the German preparations in Messines.

``That balloon,'' Guy heard and Intelligence officer remark, ''must make Fritz feel naked.''

So out of the clouds above a Hunnish hawk must dive, firing inflammable nickel pellets at the balloon. Down on earth curious eyes could follow the track of bullets from the muzzle of the gun to the big gas envelope. Beautifully the hawk circled the balloon, spitting, spitting, spitting. The gas envelope became a torch in the sky, heeled over like a sinking ship. The observer came jumping out of the basket, to float down to earth in his parachute. So the hundred thousand on earth held their breath, apprehensively upward.

``OOOO---O.'' What else was there to say?

Greatly daring, the hawk swooped again and circled the parachutist, trying to weight the swinging body with lead.

``OOOO---O.''

The hundred thousand men on earth were spellbound with the drama, torn between admiration at German pluck and apprehension for a lonely swinging observer, a helpless mid-air target.

``OOOO---O.'' Eloquent syllable.

For war in the air had the skill and romance of old-time personal conflict. Death could come from the hands of the strong and the brave. On earth killing was mechanical, shell knew no distinction between coward and hero. But even as men watched from below mechanical ingenuity was commencing a triumph over personal skill and valour. Duels of individuals were giving way to weight-carrying competitions, to efforts to reduce petrol consumption per mile, so that tons of bombs might be dropped on defenceless targets.

The hawk winged his way over Hill 63 to his own lines. The parachutist made a safe landing. The burning balloon hung like a torch in the heavens and fell to earth. And the hundred thousand turned their eyes to earth adn went about their allotted tasks.

`` `The moving finger writes and, having writ, moves on.' '' Joe was becoming impossible in the frequency and irrelevance of his quotations.

``Well, that's that. Fritz feels naked no longer. Peeping Tom has gone.''

``Nakedness is psychological, then.''

``What the hell's that?''

``There'll be another balloon there by morning,'' he avoided.

British and German ascendancy fluctuated. Neither would yield, which for the moment meant an increase in British dead. There was a new balloon above Romarin in the morning. Expeditious killing necessitated eyes in the heavens. Under the lash of high explosive the human family was learning to inhabit the clouds. War in the air was prodigious in its moral effect. Each combat was staged to intimidate an army. Earth worms looking heavenly portent knew exaltation or dejection as victory fluctuated. A million could be pulped in the mud and scarce be noticed. Only a trembling neighbour visualized such death as he crouched against a parapet or low in a shell-hole. Happy to survive, he shrugged his shoulders and went his way when danger was past. In the mud death was an event. Next morning the daring hawk swooped again, striking out the new eyes. Another torch in the sky, red sky at morning.

``OOOO---O.''

``We're losing the bloody war.''

Yet evening saw another pair of eyes gazing from Romarin towards Messines. For the third time the hawk swooped and again twilight had heavenly illuminant. The New Zealanders were thrilled and sobered by the German success. What sort of a nation was this that, outnumbered and outmaterialed, was having this moment of success, striking eyes out of heaven with nickel talons?

``Germany isn't beaten yet.''

``Fritz is king.''

For out of another twilight, as though from another planet, came a meteor leaving a trail of flame, a living, vivid rocket, for a British plane was hurtling to earth. A charred cinder of a body was borne away for burial, and while earthworms still shuddered a second British plane came crashing down.

``We're losing.''

``By Jesus, we're losing.''

Gloom for a moment predominated. And yet only two men had been killed, one shot, and thousands were shot on earth and not bagged magnificently in the sky, one burnt alive in oil, petrol, and many men were burnt with flame-throwers on land. None had heard his shrieking, but all had imagined and shrunk back from his suffering as, a living and dying meteor, he had flashed brilliantly across the sky, calling all the world to pay tribute to his death agony.

``OOOO---O.'

When infantrymen died no comets were seen. None bothered when men were churned in mud. But the contest in the air was symbolic: men read alarming portent into visible success. Fritz was winning. The heavens had spoken.

And then at last John Guy saw a German crash down, shot down while trying to limp a damaged plane back to German trenches. A speck of shrapnel or a tracer bullet had felled him from the blue, but the pilot regained control in time and tried to win back. Desperate effort. For machine-gunners and riflemen in trenches tried their luck like sportsmen and the plane came down. For a moment the trenches rang with the sound of cheering, as men might cheer at a football match, a race.

``Hurray------!'' A human note.

The man on earth was proud. He had bagged a man bird. Thousands cheered at the crippled eagle, an eagle that was a limp, smooth-faced German boy lying forward dead. He had gone to help pay the debt his country owed to an incinerated Britisher who had a few nights before flashed from life into death.

In a few days two far-away German balloons were torches at twilight, fired by a deluge of spreading liquid fire dropped from British planes. Men marvelled as terror succeeded terror. Life climbed higher into the blue that death might penetrate farther into the earth. And life in the air was being regulated. Even in the clouds ideals and courage were wilting before superior mechanization. High explosive and engineering technique were making light of high purpose. Science must win though every human value perish.

For twenty-four hours the new liquid flame that spread fanwise when dropped on a balloon was discussed excitedly in billet and \textit{estaminet}, a detail that was set against temporary German supremacy.