\chapter*{\textsf{Nerves}}
\addcontentsline{toc}{chapter}{Nerves}

T\textsc{hey} arrived at their jumping-off place in the trench and crouched low to wait for dawn. There were a few hours to go. Overhead the swish! swish! screaming of has shells could still be heard searching too rear areas. The springs were oiled and set for zero and for a little everything stilled itself, even the men, who were tired at the exhausting assembly. Ten thousand guns were quiet because ten thousand guns were ready to belch. The company rested in the second assembly trench, for it was playing the part of the second wave.

For miles along the front, from Plugstreet toward Ypres, men were crouching ready in trench bottoms, worried no more about yesterdays but appalled by thought of the fury of the day that was coming. Tired from lack of sleep, from weight of harness, from nervous strain, all the men sank close to the damp earth. For they were not in the ordinary duckwalked trench-line but in emergency trenches. Guy found himself marvelling that men could be expected to erect themselves once more and stagger up Messines Hill with such a load on their shoulders and crushed by a nervous load too. He did not know the lifting power of zero.

Someone nudged his side and a voice whispered an order in his ear. The incongruity of whispering, with the German army two or three hundred yards away, did not suggest itself to him. the imminence of catastrophe had created spiritual tension.

``Captain Guthrie wants Lieutenant King.''

He handed the message along in as tense a whisper.

``Captain Guthrie wants Lieutenant King.''

Along the trench from him went the whisper. The Lieutenant came stumbling over legs. A succession of messages came and were passed along. The tank they had passed came slowly waddling out of the night and took up a position on the trench top to Guy's left. The tank made a noise that made the whispering of men laughable, and yet, by adding to alarm, caused the whispering to become still more subdued.

``Thank Christ it's dark. If the Hun saw that tank we'd get hell.''

Fog had its uses. The low-flying planes still throbbed above, surely blinded by the night, by the vapour-impregnated darkness. Guy leaned to his neighbour to the right and found that mortal chewing gum, but there was a slight tremor in his body. He leaned to the left and the thigh he touched had a violent, almost an epileptic trembling. Listening, he heard distinctly the chattering of his neighbour's teeth. It was pitiful. He grew aware that he had a slight jump too, scarcely perceptible, but it was there. Steady, he thought.

``Steady,'' he said quietly to his neighbour.

``That's all right. It's only the damned waiting.'' The teeth continued to click.

With cynical humour and from lack of anything else to do, Guy started to paint the horrors of the coming dawn. He painted a ghastly picture with amused resignation and his neighbours grew interested. They joined him in his efforts and competed horror against horror as they poised on the brink. He cured his own and his comrade's fear by exaggerating it and laughing at the exaggeration. As he sneered he lost his own slight tremor and lessened a little the clicking of his neighbour's teeth.

``But you haven't been over before?''

``No,'' he admitted.

``It's different when you start to move. The shake goes.''

``They can only butcher a bloody man once, anyway.''

As they swopped gruesome phrases a whizz-bang came from a German gun and landed in the trench. The explosion was succeeded by a verbal clamour which contrasted with the grotesque whispering that had become habitual. And then silence fell and a groan came back up the trench.

``Stretcher-bearers. Lieutenant killed, two men badly injured,'' came whispering along the trench.

A second whizz-bang fell haphazardly out of the night and into the tank a few yards away. The explosion set fire to the petrol. The tank became a red-hot stove and acted as a grill to the bodies of the wounded tank crew. The smell of burning flesh came along the trench. The heavy fog screened the burning tank from observation from the German lines. A fine night and a burning tank on the parapet would have ushered in hell before zero.

But the excitement engendered by the shot soothed Guy's nerves, horror at the human grill stilled his tremblings. Placidly he resigned himself to anything the lingering remnant of night held. The machine would get everyone, and it didn't seem to matter whether one was torn to pieces in a trench, riddled in the open, or grilled like a cutlet in a red-hot stove. There was not much use fretting about one's fate. Whether men shivered or not they were ground or pulped, whether they laughed or cried.

``Don't shiver,'' he gruesomely advised his mate. ``If you shake too much when you jump the bags Fritz won't get a decent shot at you.''

Grinning at his own sombre humour, he saw another whisper coming as head after head turned, listened, turned, passed the message along. And when it came the message knocked the bottom out of placidity, and set him again trembling. His vaunted indifference was shattered to pieces.

``An hour to go.'' He turned and passed it on.

``An hour to go.'' The message went from him.

And he was not quite so tired. That message seemed to have caressed him into a vigour. And in a few beats of the heart, for how time flew once life had a limit, it came again.

``Half an hour to go.''

``Half an hour to go,'' he hurried it along.

``Half and hour to go.'' The whisper was thrown from lip to lip.

``It's getting grey.''

Jesus! how his neighbour's teeth were dancing.

Half an hour to go. Proximity of the unexplainable, of the knowable, for each hell had its individuality, drove away tiredness. The equipment which had been dragging him against the trench bottom was losing its weight as morning and muscles grew tense. German machine-gunners expecting an offensive some morning were crackling and whistling their bullets over the trench top. Tired very faraway guns could be heard chugging too. Only a few decent breaths and the whisper came again. Time hurled its victim toward eternity.

``Twenty minutes to go,'' came to his ear.

``Twenty minutes to go.'' The words sped down the trench away from him.

Without being aware of it all voices had grown louder, so that ears caught messages from lips a few paces away. Bloody few heart-beats knocked the guts out of twenty minutes. And he was trembling and twitching not with fear, for fear had gone, but with excitement, with muscular as well as emotional excitement. His eyelids were dancing, his lips quavering, his knees jigging. The heroes might have seemed epileptics to dispassionate observers, but there were no dispassionate observers. Twitching eyes saw twitching comrades. Night was going in truth with the speed of light, peace and ordered life were falling away. Dawn was coming, coming. Morning and chaos. Red morning. Red morning from the flickering guns. Red sky at morning a soldier's warning. The fog was suddenly no longer part of the darkness but a blanket in a growing light.

``Five minutes to go.''

``Five minutes to go, by Jesus.''

``Five minutes to go, by Jesus.'' The excited man caught at hi whole phrase.

It went along the trench that way. ``Five minuted to go, by Jesus.'' He was gripping his rifle, his wonderful bayoneted rifle, and his equipment had ceased to drag him down, sat ridiculously on his shoulders as a child's balloon. He was crouching like an athlete, he was already half-fiend, blood and iron ready to be amalgamated with blood and iron. He would ride a wave of molten metal into Messines, onto the very gates of hell. He had lost all concern for his person. He was charged with dynamic exaltation. His fleshy limbs were limbs of steel, his heart was beating with the steady strike of a mighty engine, He did not know it, but far above the ordinary man he had the power of emotional and fanatical adaptation in crisis. He was ready to embrace the devil and rollick with him in hell. He had lost all concern for his comrades, had forgotten them as he tensed himself to spring. He didn't know, he did not care if they were possessed of divine hysteria. Some were. Some were still brittle humans too aware of their frailty.

Compassion, humanness, had gone from him as he crouched. He was eager to be up and away, driving his bayonet at German breasts. Eager. Superlatively eager. He would carry his point with headlong vigour into a Hunnish chest. He strained at the leash like a hound of hell. ``Ah------'' In his ears rang the death music of the men he would kill, who would wilt away from him. For he was sure of invulnerability. He was mad if he thought he was infected by a new and magical sanity. He saw red morning with red eyes. He blended. He was Martian. Upon him sat the spirit of the bayonet. By the Almighty he had suddenly become a great soldier. And that was funny, for of all his Battalion he resisted most of the ideology of war. But his emotions had gone down before a more powerful infection.

``One minute to go,'' to his ear.

``One minute to go,'' from his voice.

``One minute to go,'' on down the trench.

The words were spat along as though from a machine-gun. Fully sixty seconds. They were all on their feet clutching at the front of the trench as the second-hand ticked on, cutting away the restraining lash. For miles along the  front thousands crouched ready to spring, some fearing, some sober humans, a smaller number red-eyed and fiendish, ambitious only to get at German throats, to ho out of the trench, to ho upward, forward, onward, with feet caressed by chaos.

In thousands of gun-pits hands were on keys, gunners stood with a second shell poised for the hungry breech. Somewhere an engineer stood ready to press a knob and fire a dozen mines. War correspondents adjusted their pads and poised their fountain-pens, ready to fit prepared terminology into the agony that was dawning. Everything was to plan. For Messines was the ideal of the planned offensive, the pinnacle of attrition. And even the German plans were ready. British intention had not been secretive, and Fritz had planned bloody reprisal. Doctors and orderlies laid out sharp knives and piles of dressing. Ambulances came forward to collect the wreckage. And German parsons prayed for German steel to spit British briskets, and British parsons prayed for British steel to cut through German briskets. Everything had been accumulated to destroy. Directing intelligence was ready for beggar-my-neighbour. Victory would go to the last surviving private soldier if all others consented to die meanwhile.

Of all this he was unconscious. He was conscious only of the desire to be hurled forward by the muzzle velocity of barrage. He wanted to move over an earth that squirmed and writhed at the thunder of the guns. He couldn't hear his neighbour's teeth dancing. The stockwhip crack of German bullets overhead was an orchestra accompaniment to his emotions. He wanted to poke his head up into the nickel barrage to test his invulnerability amid the hail. Flesh and blood were as unconquerable as his resolution, said his madness. Therefore was he armour-plated.

``One bloody minute to go.''

It was gone as he savaged that phrase, and the day of all bloody minutes had arrived.

Oh, lucky man, that his heart always beat at steady stroke when his feet were amid chaos.

