\chapter*{\textsf{Raid on the left}}
\addcontentsline{toc}{chapter}{Raid on the left}

P\textsc{reparation} made staffs nervy, increased efforts on both sides of the line to obtain information.

He stood on the fire-step in front of Plugstreet in the evening twilight of a perfect day. The air was of amazing stillness, shaken only by far-away detonation of explosives. Messines was on the way to dust. Shells hammering into the town blotched the darkening horizon red. The Sergeant came to the bay with a message.

``No patrol or working-parties to-night.''

``Why, Sarge?''

``We are going to have a raid on the left. The chaps on the left will move over to our trenches so as not to get knocked about by the counter-barrage.''

``Fine night for a murder, Sarge.''

``Yes. Fine night for a bloody murder.''

So there was to be a raid. He listened to the noises of evening, to shells bursting in Messines, to far-away chug! chug! of heavy German machine-gun, to the crack of a flare pistol immediately opposite as the first flare shimmered up and down.

``A bloody raid,'' Joe said in Guy's ear.

``A bloody one,'' Guy said ironically.

The voice of Joe answered him as hundreds of thousands of questions were answered, by declaiming aloud to all the airs in the bay:

``Dear Mother, we are about to have a heavy bombardment on the Western Front.''

``Glad I'm a spectator,'' a voice in the dark asserted.

``Me too,'' someone assented.

``What's a little raid between friends?'' Guy asked.

``As for me,'' a voice started to repeat a dialogue that had a million authors and no copyright, ``I'd end the war. Put Lloyd George and Bottomley and Hindenburg and a few German bastards in no-man's-land and keep 'em lively with a barrage or two until peace was declared.''

Surely every trench soldier retailed that prescription occasionally.

``All very well to talk,'' said a patriotic one into whose soul disillusion had not penetrated. ``But all Germans are swine, and they deserve all they get.''

``But must we get the same to give them what they deserve?''

``What wrong has Fritz done''---Guy was pugnacious---``that all the nations have not done? The German worker is a victim like ourselves.''

``Well, we'll make him free.''

``By sticking a bayonet through his navel?''

``You're a bloody pro-German and Bolshevik.''

``Talk sense.''

``Well, how can you end the war without ending Fritz or lying down on your belly?''

``We've got our own war gods to kill,'' Guy evaded.

But his opponent was not shrewd enough to force his advantage.

``The worst Britisher is better than the best German.'' He added an afterthought: ``But they tell me they have great Fr\"{a}uleins.''

``But surely you wouldn't sleep in an unpatriotic bed?''

``Wouldn't I? Helmet, puttees, gas-mask, mud, full fighting-kit, I'd show what British blood was like.''

``For me I'd prefer soap and water, nakedness, clean sheets, no lice.''

``Well, who the hell wouldn't? A man must be allowed to exaggerate a little.''

Argument, serious argument, as ever, wandered into obscenity. The serious patriot who wanted to kick the Kaiser in the stones was only aggressive on the verbal front. He was a good-natured, passive fellow actually. And then silence again conquered garrulity, conversation, like machine-gun fire, coming in fitful spurts.

``Fine night for a murder.''

John Guy leaned his chest against the sand-bag that topped the parapet and searched the cool night idly, seeing and yet not bothering because he was not on guard, or seeing objects distorted, as they are seen reflected in a slightly rippling but clear pool. Responsibility sat on other eyes, so that his eyes saw everything through a screen of internal thought. He saw merely because his eyes were open and not because he sought to take cognizance of what was about.

``But the Tommies waiting to raid won't feel happy.''

``Cursing their bloody luck.'' The words dropped from his lips for any pair of ears.

``A man does everything he's got to and blasts everything he does.''

``Um. Rushes to enlist and then puts the brake on.''

``Doesn't know what a fool he's been until he can't get out.''

``I wonder if I'll ever see New Zealand again?''

``Yes,'' said Sergeant Bill. ``If we get a leg or an arm hacked off, or get a shot in the guts, or get our eyes knocked out.''

``See it if we got our eyes knocked out?''

``You know what I mean.''

``How quiet it is.''

``Quiet, yes. But after the Lord Mayor's Show comes the sanitary cart.''

Men spoke aimlessly because the silence of waiting for the bursting of the barrage had to be filled in somehow. They were all weary from a spell in the line.

``The men going across will have loose bowels now.''

``Yes. And be bloody heroes the moment they start moving. It is difficult to be steady sitting on your backside waiting.''

``How quiet it is.''

For how quiet it had really grown.

Guy tried to imagine the raiders crouching in trepidation, or sneaking half-way across no-man's-land to be ready. He had heard of the trembling that preceded fiendish hysteria but had not yet been lauched into chaos. His mind let him see those timid Tommies grow suddenly to demonic proportions as they surged down on crouching defenders. Coming out of the night on the heels of the barrage they would seem giants, so that German hearts would quail before them as they dropped down in the trench. The shock of such berserk advent would paralyse any defence.

And he thought about the men to be raided. The black night was so mild that he thought generously of them. He saw a man in grey dreaming of Fr\"{a}ulein's caress as the velvety smoothness of night caressed an unshaven cheek, leaning chest against parapet, happy at absence of bombardment. Although, if silence long continued, exhausting alertness would supplant dreaminess. For the man in grey would listen for hostility and not to dreams, would listen for soft foot-falls and creaking wires and pour his nervous energy into the lap of night, forgetting he had ever wanted to erupt it into the lap of Fr\"{a}ulein. For the disciples and followers of blood and iron knew they were so brittle that they smeared easily against a trench.

``How quiet it is.''

Quiet enough to dream when one had not to stand and stare. Quiet enough to understand that grey-clad hell-hounds two hundred yards away loved and dreamed, knew music and art and science and appreciation of beauty. If night cloaked only kindliness, and terror was banished, how opposing armies would unbend. The trench soldier would make a bonfire of duckwalks and trench props, and sing harmonies in red light that was a social and not a killing light. They would forget the brothel because home had suddenly drawn near. They would ``Hello, Fritz'' and ``Hulloa, Tommy,'' and drop in on one another.

A Very pistol cracked across the way, and Joe saw Guy's face in the silver light.

``What are you smiling at?''

``A stupid fantasy.''

``All dreams are freakish.''

``What did you dream?'' asked Bill. ``Steak and onions, or steak and eggs? Madame or Mademoiselle, or a nice little wound, or a double rum ration?''

``I dreamed I was shaking hands with Fritz and asking him how he liked the air down here.'' Enjoying the opportunity, Guy added: ``And we all joined hands, sung `Auld Land Syne,' and went home.''

``And then you bloody well woke up.''

``Not much of a dream that. If it had been Miss Fritz I would have envied you. But you'll throw your arms around his neck and try to kiss him when he runs at you with a bayonet, if you don't look out. You should have gone to bed with one of 'em. Love at first sight, you know.''

Bill was pleased with himself.

Guy laughed as heartily as anyone. How quiet the world was except for the subdued laughter.

``When I meet Fritz I'll beat him to the thrust.''

``Why not keep on dreaming? Why not say, `Pardon me, and put a little iodine on the wound to keep it clean?' Why not say, `Now take a deep breath'? You and your bloody dreams.''

``How quiet it is.''

The silence and the knowledge of the gathering raid stifled attempts at garrulity. And now Guy dreamed again, but not of Fritz as a kindly fellow, but of Fritz as a fellow he would shoot, bayonet, club, kick in the face with the toe of a muddied boot. There were no rules that would govern his fury except the will to do all these things first. For millions would rip and tear at one another and rot before Fritz and his foes would see one another as brothers.

``How quiet it is.''

Dreams were stupid. Germans were foolishly dreaming of home and the murmuring Rhine. Guy did not pause to ask whether they were not being sweatily driven or came from somewhere else. Directly their dream would perish in hell, for dreams were shattered when dreamers were pulped. Fiends on the heels of barrage would shatter the power to dream with a revolver or cog-wheel on a handle or a bomb. And the flesh that dreamed of the Rhine would smear itself along the trench with the flesh that had dreamed of Petticoat Lane or Devon.

A gun exploded in the rear, a detonation which caused his mind to make grips with reality.

``A trial shot from a battery.''

The shot caused a German machine-gunner to bestir himself, and whining bullets brought all heads momentarily down.

``Where do bullets go to? I've never seen a spent one?''

Strange, unanswerable query in the night Flares soared up and flares sank. There was silence, emptiness, dreamlessness.

``Does every bullet find a billet?''

Who was it that asked such questions?

``How quiet it is.''

``I suppose they are in no-man's-land awaiting barrage protection.''

``Suppose so.''

``Would be rotten to be seen and get cut to pieces.''

``Get cut about, anyway. Raids are unhealthy.''

``The raider at least knows where he's going.''

``You mean where he's trying to go.''

``Looking for a bloody argument?''

``No.''

``How quiet it is.''

But the guns made a liar of the speaker. Night glowed to the rear, shells roared on the way, glowed and reverberated on the left front as a stream of shell deluged a line of trench. The night flashed white toward the cannon, red to the left where the shells burst.

``Some barrage.''

``Yes. Some barrage.''

British troops were hastening on and the barrage stilled a moment as it lifted, and then howled and shrieked again.

``They must be in the trench now.''

Silver flares all along the front lit the heavens. From the centre of hell a German rocket cleaved into the stars, spilled a shower of green and gold red stars, an SOS proclaiming the bloody and hysterical frenzy of defenders.

``SOS!'' Everyone in the bay said it.

Beautiful stars falling out of the sky appealing for protection, for retaliation, for that avenging mercy which, shot from the cannon mouth, is counter bombardment.

Above the hysterical gold and green and red brilliants the stars of heaven flickered with the passive indifference of eternity while men tore at each other's breasts. ``Save Our Souls!'' said man-made stars. What were a few souls to the eternal brilliants?

Counter barrage came, a hail through which raiders must return, so that returning became as great a hazard as going. Everywhere but in the area of the raid machine-gunning ceased. Guy had an idea that for miles German and Britisher was standing above the fire-step, attention concentrated on inferno, trying to pierce the veil, to guess the fortunes of countrymen. German feared for Fritz, and Britisher feared for Tommy, and none mourned for the human family. Two more rockets shone against the sky.

``Fritz wants someone to save his soul.''

``Save his dirty bloody hide.''

``Anyone would.''

``Yes. But to hell with the Kaiser.''

Somewhere men were dying, grovelling in pain, blind and numb with agony. And out of heaven to which they appealed came high explosive, behind high explosive came fiends to raid, a wave that would recede and leave broken flotsam on alien territory, that would recede bearing away groaning and battered captives. Maybe as dying man saw dying man enemy discerned a brother and marvelled. For hatred as well as love can ebb with blood.

The barrage tapered off, the raid was over. A few flares flickered up again and no-man's-land was wreathed in silver radiance. A machine-gun started traversing, and men went back to earth. Machine-gunners on each side of no-man's-land joined light-heartedly into the playing of the rhythm that the feet of people play in the theatre.

``Trrr---T. Trr---T,'' from British gun.

``Trrrt---T. Trrr---T. Trr---T,'' from German.

Guy listened to the refrain of fellowship expressed in nickel pellets.

``Me for forty winks.'' He left for the signal bivvy.

``Listen to those silly cows playing their game,'' a number of the squad said to him.

``It's a friendly war.''

``Yes. Perhaps. But there's going to be a lot of blood spilt down here soon.''

``German blood?''

``Who knows? They can do anything to a man except put him in the family way.''