\chapter*{\textsf{Le Bezit}}
\addcontentsline{toc}{chapter}{Le Bezit}

``L\textsc{e} B\textsc{ezit}, it is,'' the Lieutenant informed them.

``Yes, Sir. Where's Le Bezit?''

``That's up on the side of Plugstreet. We are going to relieve a division that's let Fritz get the upper hand. The Division's tail is down.''

``The Division's trousers are down.'' The Sergeant had been on a tour of inspection and improved on the Lieutenant.

``Yes. Suffering from mental dysentery. Backsides calloused from lying face down on Fritz's knee.''

``We'll keep our trousers up, anyway,'' a private risked.

Filthy slurry drained from high and dry German trenches and filled the British system. Parapets were low and the line was dominated by Fritz snipers, so that men could not stretch upward in day-time, and the trench bottom was full of filth so that lying down was impossible. The men in the trench had to squat like ducks. Sanitary conveniences were spotted, and men suffered agonies of restraint and insanitation rather than excrete and urinate in the filth they were compelled to slop about in. The enemy trenches were no more than fifty paces distant and only the German line was well wired. Raids from the German line were frequent.

``Don't go the the stool,'' the departing Sergeant advised. Fritz has a rifle trained on that seat.''

``Good,'' Joe said. ``Good. A dangerous lavatory is good for constipation.''

``You won't joke so much about it in a few days.''

``Well, why didn't you make the ------ house safe for democracy?''

``Wait and see.''

``Fritz must enjoy the joke he's been having with this division.''

``We'll turn the joke sour.''

There was faith in the Battalion's power to do better.

German ascendancy was soon challenged. Trenches were deepened, bagged, timbered, parapets were broadened, a thread of wire in no-man's-land became an entanglement. Working-parties got chopped about but progress was made. New Zealand nickel started to bite at German heads and the German heads were the heads of snipers. Soon New Zealanders were walking with heads erect in dry trenches instead of crawling about on their bellies in slime. Fritz surrendered his superiority grudgingly and fought back, raiding and being driven off. As always, when good divisions fronted one another, the front achieved a bloody stalemate. ``\textit{Esprit de corps},'' the General probably said, if from his remoteness he could even imagine the desperate bloodiness of the conflict.

John Guy was only a spectator of the front line hazards. He was fortunate, for his curiosity might easily have made his head vulnerable to sniper's nickel.

``Guy.''

``Yes, Sarge.''

``Attach yourself to Headquarters as gas guard.''

And since in modern war the higher one's rank the deeper and safer one's dug-out, he congratulated himself. While his comrades tasted blood and filth he inhabited a strong, gloomy, well-retired barn. He slept all day and stood by the gas gong at night. Standing by the gong, he directed the bearers who brought out the dead under cover of night.

``Over in the brick shed. They get them out on a trolley during the day. Who was it?''

``Poor old Cyril. There's two more coming.''

``What! Three to-day?''

``Yes. The bastards got three.''

Snipers' corpses were usually unmangled. A short-range bullet is a clean destroyer. Sometimes even the aperture through which death had entered closed over.

``Just put him down on the floor.''

On his first night there had been lain down the beautiful, unblotched face and body of a boy who had enlisted when he should have been kept by the fireside. Death had winged its way and left a mark like a small blue dot. In the night Guy heard strange noises and he entered, lighting a match. He nearly vomited. A dozen rats had gone scampering from the face and hands of the corpse. They paused and peered at him from the dark corners of the shed, showing their shiny, hungry eyes. The face of the boy had become a death's head with all the tender flesh eaten away and two grisly ears awaiting their turn. The flesh of the hands had been torn away as well. The bared teeth of the boy showed pearly white, or maybe white smeared with blood.

``Get out, you beasts. Get out!''

Maybe some day a writer will write a greater essay in praise of the rat. The rat was to the battlefield what a screaming gull is to a harbour. It insured against pollution. Corpses rotted and bloated only where death was too insistent for even rats to live; in other areas the rat avoided pestilence. But soldiers were too near to see beauty in battlefield hygiene. Soldiers were too short-sighted to see the useful purpose.

Guy didn't grudge Fritz his victims, but he grudged the rats. Scenting flesh, they had gathered and they awaited his retirement so that they might return to the feast. They were resentful at his interruption.

``Get out, you beasts. Get out!''

They retreated when he was noisy and sneaked back when he grew silent. And he couldn't yell all the time. Boys were too good for rat food, but the contest was unequal. In the end the rats won. They were more patient and their stomachs, unlike his, were not nauseated. He retreated at last to let them fight and glut their paunches.

``Lift the bodies on the bench,'' he advised next night.

But when he hastened into the room the rats jumped down and ran in all directions. When he struck a match he saw the sickening sight he had seen the night before. Again they peered at him from hole and corner as they awaited his departure. And suddenly he knew that if they didn't eat the fresh corpse they would burrow down into its shallow grave and eat it later on. Suddenly he knew that the rats might eat even himself and that thought made his spine twitch.

Sickening, he left the shed again and stood by the gas gong in the crisp air. He listened to the machine-guns traversing and searched the sky for familiar stars. Far away, Very lights circled up and down, making war's cruelty beautiful. He tried to think only of night's cold, dark beauty while in the shed rats fought for the choice portions of heroes. War was indecent.

Next day they slung the table by wires from beams and unloaded the dead upon that. But the rats were not to be denied. They dropped down from above and had their fill of human flesh. He gave up trying to defeat them. They were in a league with the guns and there was no escape from their teeth.

After that, for a week, when his system was trying to woo sleep he suffered from a nightmare in which rats gathered around ready to devour him, a nightmare he could frequently repulse only by giving up attempt to sleep and lying in alert wakefulness. And his mind played with him during hours of wakefulness, too. For he saw men sorely wounded lying caught in barbed entanglements, and he saw a horde of greedy mouths come out of the night and tear at the living flesh that could not resist. And he knew that this must have happened in a thousand cases, that living wounded men must have had their bodies, maybe even their faces, eaten away before their comprehending but helpless eyes. Was not France full of bleached skeletons; and all the skeletons could not have been the skeletons of men whose flesh had only been eaten cold. Hungry rats would tear at the helpless living and at one another in mad desire to have the choicest parts. Hungry rats would not be deterred by feeble groans.

Guy was tired of fighting rats that were as sure as fate. He let them fight for the lips and cheeks of the newly dead, he let them bare the bones of the hands. He moved over to another gas guard, a less imaginative fellow, who had come to France via Egypt, and found an\ae sthesia for the horror of his thoughts. As he watched the soft magic of the Very lights that cast a shimmering glow over the battlefield horror, he listened to his companion talk of the parade of vice he had witnessed in Egyptian brothels. He listened willingly. He encouraged the loquacity of the unreticent. He listened to white and negroid love being valued. He heard of the awful tricks played publicly on sex for a few piastres. He heard of the tiny boys who played their parts as touts. And he listened and questioned greedily, for he wanted his companion's voice to continue so that he might not be cast back upon the greater abomination rats played with the boys in the brick shed. He escaped from the horror of the shed to the glamour of the brothel, and the change was good.

``And they did that in front of everyone?''

``In front of an audience. A public performance.''

Machine-guns chattered, turning flesh to rat fodder. Very lights circled. Was the soldier so vicious when he escaped wittingly from the shed to the brothel? At least the brothel yielded a perverse flagellant joy.

In Le Bezit rats had plenty of food. Not the bodies of grisly old men who had lived a full life, but the firm flesh of boys. They knew full measure. The rats got on his nerves.

``I wish we could tie a few statesmen on the table.''

``Until they declared peace?''

``Yes. A few plump bodies.''

``The rats would know them. They'd be too round and clean and they'd taste of scented soap.''

``It's beautiful to die for King and Country.''

``Now, as I was saying, I met a little fellow who asked me to follow him for a little jig-a-gig.''

Wise soldier who refused to keep his mind too long on the major indecencies of war.

When moral and material ascendancy had been re-established, the Division was relieved by an English division as good as the other had been bad. So the rats grew corpulent on the rosy pink of English boys again, instead of on the more swarthy southerner. The Division left the trenches safe and deep and no-man's-land war at a stale-mate. Their successors kept it there.

``We are always moved on when we've made a good job of a sector.''

But all moving divisions grouched like that.