\chapter*{\textsf{Comedy or tragedy}}
\addcontentsline{toc}{chapter}{Comedy or tragedy}

W\textsc{hizz} B\textsc{ang}!

Guy stirred uneasily in his sleep.

Whizz Bang!

The far-away scream of a light shell half awakened him.

Whizz Bang!

In his half wakeful drowsiness he started to relive his daylight run.

Whizz Bang!

It was as if he were being chased again upon the cobble road.

Whizz Bang!

Gradually the bursts cleared drowsiness away altogether and he started to listen.

Whizz Bang! Bang! Bang! Bang! A salvo.

After each interval the shelling grew closer.

Whizz Bang!

The battery was having a fresh lot of fun with another runner, Guy realized. The thought must have simultaneously flashed into other minds whose sleep had been interrupted.

Whizz Bang!

The Captain raised himself on an elbow, gazed understandingly at the other recumbent faces, spat out a phrase again:

``Damn the Colonel. The fool's at it again.''

Whizz Bang! Bang! Bang!

Nearer and nearer and nearer.

Whizz Bang!

After a silent interval a voice outside among the bricks called a directional inquiry. The gas guard beyond yelled an answer to the runner, who emerged at a run from the refinery and doubled like a rabbit up the road to plunge into the cellar. The man who burst into the cellar was yellow and sweaty and nervous.

``The swine have been sniping me with whizz-bangs.''

Guy laughed outright. He couldn't control himself. There was an hysterical concern about the messenger that was comic. It was amusing to think of some other poor fool ducking and diving as he had done himself while a light German battery tried to snipe him. By Jesus, it was funny when it wasn't oneself.

But men in safety always roared at the undignified haste of men escaping shells. Once Guy had seen a Minnie floating in the air above four men, and the four men's urgent scramble from under had set a whole platoon laughing. To tie death to a human's coat-tails and set him running was funnier than tying a tin can to a puppy's tail. The awful urgency made the incongruity gargantuan. Anxiety could be so comic.

``What the hell are you laughing at?'' The runner was not light-hearted.

``At your funny run.''

``Can't see much fun in it.''

``I did it myself a few hours ago.''

``Did you stop to laugh?''

``Not on your life. But Fritz laughed, I'll bet.''

``You've got a perverted sense of humour.''

``Perhaps.''

The captain read the message and the runner talked.

``The Colonel thought Guy had been killed or wounded, so he sent me to find out and bring another message. He wants a reply.''

``To hell with the Colonel!''

Heavy shell started to drop around the bivvy again, and the runner who had just brought the message visibly twitched.

``Um. If I don't send a message back the Colonel will send the whole Battalion, and every bloody German gun will be trained on us. I wish the Colonel was a runner.''

He wrote a note and folded it.

``Here, Guy. You've had a spell. When the shelling eases you go with this. I wouldn't send anyone if it wasn't for the Colonel. Fritz will think we're mustering an Army Corps here.''

``Um. And if I don't get there?'' Guy grinned.

``Use your head and your feet. They don't drop a barrage on one man. They'll let you alone this time.''

``They drop too much for me,'' said Guy with a comic ruefulness.

As Guy buckled on his webbing and picked up his rifle the runner who had arrived started to laugh, a deep, hysterical laugh.

``Ho! ho! ho! ho!''

Guy's mournful appearance unlocked the doors of uproarious mirth. Everyone in the cellar, jumpy with nearby concussion, yelled aloud with laughter.

``It's not so bloody funny.''

They roared the louder when he protested. He looked so sorry for himself. He was caught up on the tide, and stood grinning by the bag that hung over the entrance.

``Ha! ha! ha! Ho! ho! ho!'' Deep-throated and shallow-throated it came.

``Balls!'' He called the word at them and they rocked.

The Captain grinned and added a phrase.

``To the Colonel.''

``Ha! ha! ha! ha!''

He turned and ran up the steps, his back pursued by the joyful syllables. The whizz-bang battery gave him hell. The traffic had keyed it to enthusiasm. The observer knew where he was going, and trained his guns accordingly, tried to turn him from destination as a greyhound does a hare from cover. He ducked, dived, ran straight, zigzagged, hesitated on his knees, spurted, sweated, cursed aloud. And he was frantically scared all the way when returning because his back was to Fritz. His back was broad, naked, unprotected. Run toward danger, and danger and menace are overpowered. Run from danger, and fear disintegrates, almost paralyses.
