\chapter*{\textsf{From Fleux Baix to Le Bezit}}
\addcontentsline{toc}{chapter}{From Fleux Baix to Le Bezit}

B\textsc{ut} spring was coming and the Division moved on to livelier quarters. The period of rest and refitting was at an end. A new regiment had arrived to blood itself in the quiet sector. The Division farewelled the Germans opposite by a severe raid that left New Zealand and Saxon blood on the snow. Like everything in France, Fleux Baix was half forgotten as backs were turned upon it. Villages, men and events were like that. A battalion moved on to new incidents, a diminishing number of men recalled the past. An \textit{estaminet}, a face gone down to the grave, might be recollected a time or two, but soon there was no companion with whom to indulge the recollection, and a million new concussions kept the mind focused on new emergencies.

Fleux Baix had been too quiet to be inhabited for long. Death had garnered an odd soldier but wastage to the cemetery had been a mere trickle. Spring was at hand and good divisions had to be moved to where blood spurred torrentially.

``Where are you going?'' inevitable question.

``Don't know,'' inevitable answer.

``Quick March,'' commanded the Captain.

Battalion marched heavy with equipment, still heavier with conjecture. It marched along the cobbled roads and journey and the day ended.

``We're here.''

``Where's here?''

``Romarin. Just behind Hill 63.''

Out of the line and yet in billets, harassed occasionally by German artillery, the Battalion spelled. In that moment men took stock. The German army had cleared from the Somme, leaving the British army in possession of a desolate quagmire. Men in France did not agree that retreat spelled victory. They knew what it was to flounder in a midden, what it meant to fashion a new trench line far away from billets, canteens baths.

``Save us from victories of that sort.''

The opinion flared into violent wrangle, for there were a few souls who, with the newspaper men, believed that every captured hill or swamp was the key to victory.

``War of attrition is all right to read about.''

For propaganda by cloaking Headquarters with infallibility nearly lost the war. The soldier might protest against war of attrition, but his protest was quenched in mud, drowned by the roar of guns that shattered his dissenting flesh. Murder on the large scale lent to large-scale type.

``War of attrition. I hate that phrase,'' Guy said.

``Correspondents are compelled to ring verbal changes or get sacked,'' Joe advised.

``To find a new way of saying 'Heavy bombardment of the Western Front.' ''

``Exactly.''

``Now this 'war of attrition' is the newest.''

``What does it mean?'' a third voice asked.

``Well, it's this way,'' Guy replied. ``We kill German and the German kills a Britisher. If the last man left alive is King George, we win. If the last man left alive is the Kaiser, we lose.''

``Bit exaggerated, but it's like that, I suppose.''

``Listen,'' Guy was pleased with himself and his audience. ``Did you ever see a small boy with a stick and a new pocket knife, with plenty of wood to whittle? Well, we are the wood.''

``Speak for yourself.''

``Well, I am the wood.''

``Yes. And they can do anything to you except put you in the family way.''

``If the war goes on babies will be created by numbers.''

``I hope they make me O.C. commanding Procreation. I'll give lessons in straight shooting.''

Men often toyed with that idea. More voices hastened to contribute.

``Marksmen's badges for the successful procreators.''

``Woman with large families will insist on equality of sacrifice.''

``Motherhood Battalions and a wound stripe for each baby.''

``But the Brass Hats will want to be the Papa Battalion.''

``No. We want cannon fodder, not generals.''

``For me I would refuse to breed cannon fodder.''

``For me I don't believe you could if you tried.''

With such jesting men revenged themselves upon high rank and high society, a tame revenge which helped the time to pass, akin to an impish poking out of the tongue at a teacher's back. Men were not criminal but human when they clicked heels, said ``Yes, Sir,'' and murmured an inaudible ``Bloddy bastard.'' If the body's fate was determined they could bait determinism with the illusion of free will.

Yet in Romarin there was hope. Spring and the spring year of the Allied effort was at hand. Allied material equipment had grown, was still growing. At last the tide was setting convincingly against the Hun. The army could still be infected by the stupendous addition to mechanical effectiveness and believe that in some magical way British guns and bayonets would blast and throw aside German bodies right up to the gates of Berlin. With such sheer weight of material even incompetent generalship seemed not to matter. How were the Other Ranks to know that High Command would use their mechanical superiority to create a bog in which to drown their own divisions. Above all that spring, the Other Ranks had a sombre confidence in themselves, a belief that man for man they were as good as the enemy. That did not mean that men in the prime of life were enthusiastic about dying. It merely meant that there was no demand to quit.

``What a good time in Berlin, boys.''

``Fr\"{a}ulein will receive us as conquerors and not as paying guests.''

``We wouldn't take liberties with their women. That would be un-British.'' said a fastidious one.

``But you don't take liberties even with your own. I don't mind being un-British for a little while, anyhow. When in Berlin do as the Huns do.''

In Romarin men were soon seeking outlets for carnal desires. Some few unlucky men discovered the haunts of a clandestine prostitute, a woman who sold her favours as well as other merchandise. The area was too near the shells for the organized brothel, but clandestine prostitutes dared official prohibitions and shell-fire. Many men braved the enchanted chance of infection and paid the high prices.