\chapter*{\textsf{The hysterical hero}}
\addcontentsline{toc}{chapter}{The hysterical hero}

A\textsc{nd} he was in front and alone. He did not know what his comrades were doing. He had looked neither to the right nor to the left and his comrades had vanished. For a moment or two this did not trouble him. For a moment he was not of the British army but a man gone to war. But a sudden sense of loneliness compelled him to take stock. He was full forty paces ahead of the nearest New Zealander and did not know whether his late comrades had steadied their impetuosity or whether they had been struck down in some shell-hole. not that it mattered. He could hear bullets spitting about him but his attitude of mind seemed to deflect them from his body. On his left was a deep sag in the advancing line. And men were going to earth. His eye noted and his mind wondered. Why?

But as he asked himself the question he saw it all. Ahead on the left there was a strong point and Hunnish machine-guns were mowing down men as a swinging scythe mows tall grass. Fanwise, New Zealanders were lying dead, spread-eagled around the circle of traversing machine-gun barrels.

``What the hell? What the bloody hell?''

He yelled the phrase aloud to himself. He was enraged at this evidence of German stupidity, Hunnish bloody obstinacy. Could not the silly swine understand the futility of such resistance? He would kick the whole bloody nest of them out of the way. Berserk fury hurled him along. He Hurried to the left and refused to lie down or crouch in a sell-hole, fortified as he was by a conviction that nothing could touch him. Not that it mattered if anything did. Life was not worth living, he would not be able to breathe, unless he brushed this nest aside. he must go on or his heart would stop. He would die of helpless rage if halted in immobility.

``What are you waiting for?''

The reason was so apparent that crouching men vouchsafed no answer, although he knew he bawled the question.

``German strong point,'' someone answered.

The country was dotted with the corpses of other hysterical fools who had believed in personal invulnerability. The crouching men were rational humans who had not been caught up into soldiers. But as he looked three men stood erect at the far end of the sag and ran forward.

Chug! Chug! Chug! Chug! Chug! Chug!

One at a time they were cut down, crumpled forward, somersaulted, stilled.

``The bloody hell. I'll fix those bastards.''

He started to run at the strong point. ``Another candidate for the V.C.'' Strangely enough, he heard Bill yell those words. ``Bloody rot.'' What bloody rot! Who the hell wanted anything more than the joy of kicking a German in the teeth? They couldn't hit him, anyhow.

Straightforward he ran. He heard bullets whine and crack and he didn't even wonder whether they were whining and cracking around his ears. The cracking only heightened his frenzy. He ran and ran and ran. His steps were giant-like, so soon did he cover space. And as he closed the whining and cracking ceased. He did not know whether the guns had gone out of action or whether he had outflanked the extreme angle of their traverse. He ran into a flank trench that led to the rear door of the strong point. Stepping across three wounded Germans, he ran around a concrete wall to a door, a deserted door. And as he reached the door cunning came to the aid of frenzy. With phenomenal speed he reasoned: ``If I go inside with a pointing bayonet I'll get bayoneted. I must be careful.''

He threw down his rifle and rushed inside with empty hands. There was a swarm of wounded and unwounded men in grey, ready to surrender the moment resistance was of no further avail.

\emph{``Kamerad! Kamerad! Kamerad!''}

And at the great moment of his soldier's life, when life and death had been the stakes and he had won, he behaved as a soldier. He rushed at a German officer and seized the first-class Iron Cross that decorated the breast, inveterate souvenir-hunter even in the charnel-house. With the toe of his boot he assisted the German officer into the open. The German rankers grinned, but there was a sickly something in their amusement. An officer in Guy's regiment had come from somewhere and had joined him in the pill-box, and the Germans, completely surrounded, were pleased to obey. One by one the officer and himself rooted the Germans out. Guy chased one man out, and as the prisoner reached the open he was confronted by a New Zealander with a fixed bayonet. Alarmed for his own safety, Guy cursed his countryman.

``Out of the way, you bloody fool. Let the swine out. There's a rabbit-warren here. Kill him and how do we get on with the rest? I don't want to bloody well get butchered.''

He found himself possessed of an Iron Cross, first class, ten watches, five rings, and three pairs of binoculars, which he added to his already crushing burden. When he emerged he was envied for his souvenirs. But soon life was to be worth more than a million souvenirs.

``I'm a bloody pawnbroker.''

``You look it. Were you in there without a gun?''

``Of course.''

``You're a lunatic as well.'' An opinion that was to persist.

He had no time to persuade to the contrary, to cause others to see the logic of his method. And he had no desire. Elated with success and souvenirs, he was off again, chasing the barrage while men still remained to search the strong point. On he must go. The Hun was willing, the line was bending. As he ran he threw away his entrenching tool and stuffed the canvas container with his looted watches. He put the Iron Cross in his pocket, where it rested against Omar and the yellow lock of hair. Forward he went. And because his company had attained its objective and the next objective was for others, he went alone. In his eagerness he was unconcerned about his unaccompanied state.

The next wave had not yet launched itself and was only now moving to position for the movement. He was lucky for the guns were playing heavily on the lines four hundred yards ahead, keeping down the heads of gunners, and the smoke of battle concealed him. The crackle and whine of bullets did not appear to be aimed particularly at him. A tank was going forward too, lumbering on, ready for the next wave. He outpaced it and waved a hand at a loop-hole. He saw an object to his right in a shell-hole, and got down and crawled on all fours toward it, jumping and rushing headlong in the last few yards. It was a packing-case he captured, an empty packing-case. He nearly cried from pent-up emotion and frustrated rage. He wanted to kill men and he had subdued a wooden box. The hysterical emotion that could deflect bullets and shell could not conceal him from the satire of that packing-case. The ravening beast had pounced on a ball of wool.

But what bloody right had a packing-case to be in a shell-hole on a battlefield? It seemed as if it had been deliberately placed to laugh at him. It was more out of place than a floating kerosene tin in the trackless ocean. He kicked the case with venom, one foolish kick, as though a box had a damageable personality, as though it were a sneering face.

``Bloody box!''

Mars, the jester, was laughing gleefully.

``Guy!'' A voice was calling him back.

But he hurried on, heeding the voice of the beast within. He wanted to get at the Hun, the rotten swine who had left an absurd case in the middle of a battlefield. Impetuously he blundered ahead. And Mars was genial to him. He was being spared the crushing impact and was being saved for prolonged torture. A machine-gun in a strong-point would have knocked red-eyed fanaticism out of him and would have left him bleeding and grunting to death with a box of live pigeons strapped to his back, but there was no immediate strong point. Forward, forward, forward he tramped alone, to slake his killing lust, to revenge the absurd mockery of a blasted packing-case deliberately left in a shell-hole on a battlefield.

And then he saw something in grey in a shell-hole ahead, an alive something, and he bore down upon it. Was it a section of men or was it a machine-gunner? Was it a wounded German only half broken on the bloody anvil? Whatever it was he would capture or kill or brush it aside. Over the soft churned earth that gave beneath his feet he moved, and it seemed that behind his bayonet point was the irresistible momentum of the whole army. His point was the apex of martial pyramid. Whatever it was it would give or be pierced. He thrilled to the killing joy of his sharp, clean steel, inwardly gloating over its penetrative force.

But the thing in grey was behaving queerly. It neither got up to fight nor held up its arms in surrender. It jumped about and yet it didn't shoot at him or get ready to meet him point-to-point. It did none of these things because what he found clad in grey in the shell-hole was no longer a man, because it had lost all sense of life and martial glory. Once the thing had been a man, but now it was an incredible botch, a nightmare fancy. The top of the thing's head had been lifted neatly as a hungry man splits off the top of an egg. The grey brain was exposed to heaven, to sun and dust. Two eyes hung down the face, pendant on the end of threads. Against this nightmare creation he had careered in killing elation; against a heap of botched flesh he had charged with naked bayonet. The botch confronted him, but with grey brain and pendant eyes instead of with steel and nickel, and he almost collapsed in defeat. The more primal drove his own bestiality back into its lair. The thing should have been dead, but tenaciously it clung to life. If the soul was disintegrated, the flesh persisted. Bereft of consciousness, it stood on its feet. The loathsomeness of all history was incarnated in the botch that had been released from man by the thunder of a gun.

``Ugh!'' He was sick at heart as he contemplated.

Hysterical elation fled. He grew suddenly limp, unheroic, weak. The heavy equipment which, perched on his shoulders, had seemed to add velocity to his forward momentum, grew instantly and intolerably heavy. Its weight bore down on his sick spirit. God, he was tired, tired, tired. Funny, but he hadn't thought of being tired before. Tiredness had not come gradually.

``It should be dead.'' That, in truth, was what he said, automatically using the neuter, for ``it'' was neither man nor beast, but just a heap of flesh. ``It should be dead.'' His voice had an accusing tone. This thing lacked all sense of decency. And ``it'' had to come at him on top of that stupid packing-case. The guns were spawning grotesques.

He would shoot at the beast. According to all the rules the thing was beyond misery, merely flapping about like a headless fowl, its twitchings like the prolonged twitchings of the legs of a dead centipede. But he would shoot it to void his own misery. It must cease to obstruct his progress. Something dead he could go over, anything living he could go through. This thing alive and dead was impassable. It must fall to earth and become merely a dead man. He put his rifle in line with breast, but it seemed to divine its fate, so it fell down on its hands and knees and whined like a dog. The eyes fell out and from the face, and swung on their threads. And the insane grotesque made him sane. He knew compassion allied to nausea. He grew naked martially. The emotional garb of hysteria was stripped away. No longer was his soul and body armour-plated. He was flesh and blood that could be botched. His mind had been the mental counterpart of what confronted him. He had been drunk with the primeval slime, ready to out-claw the most vile. But he had grown sober and vulnerable, as frail as a man in grey had been under the thunder of a gun.

``What the hell am I doing here, anyhow?''

It whined and grunted as he walked away, but he left it to whine and grunt with the sun beating down on the exposed brain. He could have killed a man, but he couldn't kill this that was already dead and in hell. It could yelp and whine and twitch until it fell down to stillness. He was weary and could scarcely lift one foot beyond the other.

``We called for you when you went on.''

``I was after souvenirs.''

``You mean medals.''

``Blast medals.''

He reached down to show the ticking watches that filled his entrenching tool cover, to find them all lost, spilt as he had stalked freaks and packing-cases. And nothing short of compulsion would have sent him back to look for them.

His impregnable carcase had grown brittle. Watches could go to hell. The whole lot could lie out in some battlefield shell-hole ticking their way to silence, as short of human meaning as the thing in grey. Life was more important than a thousand watches.

``How many did you say?''

``Ten.''

``Where?''

``Somewhere over there.''

``Go and get them.''

``You can have them.''

None moved to find the gift.

From the moment of zero until the incident in the shell-hole had seemed a flickering succession of incidents, but as he stood he grew aware that morning was late. In those hours he had gone from civilization to slime. He had been back a million years. He had indeed been the supreme beast of his Battalion. Two records of his bestiality had been incorporated in a dispatch by a commissioned observer, but of this he was unaware. Under the caress of high explosive he had revelled and wallowed in the primitive. The most sensitive had become the least civilized.

``What's the time?''

``Only ten. We'll get hell directly when the German batteries start to blow us out.''

``It seemed only a few seconds but now it seems a few million years since we hopped off.''
