\chapter*{\textsf{Bull-ring philosophy}}
\addcontentsline{toc}{chapter}{Bull-ring philosophy}

``C\textsc{an} we go look at some of the nearby pill-boxes, Sir?'' For ``Sir,'' the title blown to hell at zero had returned. The officer had discarded his private's tunic. The army was again divided into officer's and men.

``Where to?''

``There must be a lot of dead Huns to rat.''

``All right. Don't stay away too long. See if you can find me a good pair of binoculars.''

Guy and the batman went curio-hunting. They found pill-boxes full of dead Germans, men shot and bombed after desperate resistance. The bodies sprawled everywhere and the pockets were nearly all turned outwards.

``Some lousy swine has been here before us.''

In an obscure cellar they found four badly wounded, festering and groaning in the stuffy heat. Some hurrying man had thrown a Mills bomb upon them on his way forward the day before. The four were conscious. Afraid of death in the cellar, they yet were suspicious of Guy and the batman. Lingering death was preferable to violent.

``They think we'll kill 'em.''

``They'll die if they are left here, anyway. They're too weak or too scared to crawl out.''

``So would I be.''

``Let's carry the poor cows out.''

``Righto.''

Tugging and heaving and ignoring groans, Guy and his companion carried them to the rear of a concrete wall by the side of an old road that ran through the heart of Messines where they would be on the track of bearers. They rigged up shelter from the sun's fierce heat. One died or grew unconscious after they got him out. The four yielded a poor return in souvenirs, all they had was their wounds and fear. A tin of filthy water from the bottom of a nearby hole they consumed greedily. A second tin was left full again within reach. The iron rations of a dead New Zealander were divided between the Huns.

``If they live a day or so they'll be carried out.''

As they went on their way a rare mortal to be seen in such an advanced zone appeared. A red-tabbed Staff Officer came threading his way along the cobbles. Guy hailed him.

``Will you direct a stretcher when you go back, Sir?''

``What for?''

``Three wounded Huns.''

The officer laughed.

``Three wounded Huns. Knock 'em on the head.''

``Knock 'em on the head yourself,'' answered the batman.

The red-tabbed officer resented such insolence, but there were no means of applying discipline in such an area. Privates could curse even High Commands temporarily.

``Carry the swine yourselves.''

``We can't leave or we would.''

``Where are they?''

``Over there.''

The officer examined the men and touched one roughly with his foot. While the eyes of all dilated, he fingered his automatic. Almost regretfully he turned away.

``Like to shoot 'em on the nest, Sir?''

``Bloody insolence.''

The officer turned away and strode along, but paused to call back.

``If they dodge their own bombardment this evening they'll be lucky enough to deserve being carried out.''

He was right, for the Hun put down a rare bombardment that evening. Not that the wounded men lost much energy dodging Prussian shell.

As Guy and the batmen went back to their post they met the leg-weary bearers returning from the Red Cross station, scarcely able to drag one foot in front of another. Guy sensed the unfairness of his suggestion as he made it.

``Three wounded Huns over there.''

``Well, what of it?''

``Any chance of getting them out?''

``If they live long enough. We've got enough of out own now.''

This wasn't trench warfare with its ordered flow of life and death. This was battle, ebb and flow of nations, of armies, of hatreds. Bearers had enough to do. The day before more than enough. Bearers couldn't risk too much German shell to save German lives that had been shooting at them a few hours before. Expediency determined current morality. But the were not given the right to refuse,  for a German salvo landed and voices shouted.

``Stretcher-bearers.''

That was that. The Huns would lie until their own shells got them, until their wounds gangrened, until they died of loss of blood, or until the sector grew quiet and enemy could be salved without too much risk. Fighting men could not run errands of mercy for their own let alone for a few Huns. Guy made light of the problem when he could see that goodwill was of no account.

``Should have knocked 'em on the head.''

``Knocked what on the head?'' They had approached a group.

``Three live Huns.''

``Where are they?''

``Behind the pill-box.''

``I'll fix the bastards.''

A grim figure started off with a rifle and bayonet. Guy followed him. The batman followed Guy. Some tried to call the killer back, some sneered. It was a matter of temperament.

``Don't!'' Guy entreated, outraged.

``Get out of the bloody way.''

Well, he couldn't go to war with a comrade over three Huns. It would have taken bullet or bayonet to make the killer desist. He looked on. A bullet flattened one, while the others, wounds notwithstanding, wriggled and squealed. But a bayonet ran one through as he turned on his side, making Messines echo with his fear and pain. The other was pinned to earth as a pin holds a fly to paper.

``You swine!'' the batman cried.

``Saves the bearers.'' Guy shrugged his shoulders and accepted what he could not revoke.

``Your a swine too, Guy, to talk like that.''

``No use crying over spilt milk.'' And his laughter unnerved the killer.

``Shut up.''

``But you shut them up.''

``But a Hun killed my brother.''

``So you killed a whole family.''

``Shut up, Guy,'' entreated the batman.

And Guy knew that the killer had only applied the logic of the bull ring. And he had been tortured into hysteria by the loss of a brother. He had done in cold blood what Guy had been prepared to do in passion. Where was the difference? ``Don't show 'em mercy.'' ``They'll take it for weakness and shoot you in the back.'' ``The more food a German eats the less for a Britisher.'' ``The only good German is a dead one.'' The philosophy of the barrack square had been applied. And the man who had killed was no coward. He was brave, could be soft-hearted too, to a comrade. And he would have expected and would have taken a German thrust in similar circumstances. War had made him arbiter of life and death and had biased his judgements by killing his brother.

Guy and the batman discussed it all irrelevantly on their way to headquarters.

``One of 'em hung on to his iron rations.''

``A long journey.''

``Rotten business, Guy.''

``Perhaps. Get used to it. Idea seemed worse that the act. Thousands dead all over the hill. What's a few more or less?''

``You're a funny cow.''

``I don't know. The artillery shell men in cold blood. Wounded or not. I'd have killed anything yesterday for fun.''

``But------''

``But be damned! I don't believe in militarism and I ran at a machine-gun. It must be queer to lose a brother.''

``You believe in war.''

``Never did.''

``Well, what are you doing here?''

``Damned if I know. Damned if I know.''

``Funny cow. Funny cow.''

``Well, what are you doing here?''

``I wonder at times too.''

``What's a Hun more or less? Life's cheap.''

Again in the evening the guns unleashed the steel hurricane. Again curtains of flame fell from the clouds to the heads of men. Again men forgot to be philosophers as they crouched and clawed at earth for protection. Shells pounded against concrete and howled around ears. But Guy was safe, although he grovelled abjectly nevertheless. If his mind told him that he was safe in his deep hole behind the concrete protection, his nerves shrank at each concussion all the same. His was the lucky place this time, and he knew no hurt, while other groups of men were killed and maimed. And his nerves steadied a little by a mild tipple at the neck of a fresh bottle the runner had brought for the Captain. The mouthful of spirit gave the body detachment. Being an officer gave one the title to ample whisky. Being with an officer, he caught at an odd drop.

Ration fatigue brought rich freight. Men filled their stomachs with the food of the dead and wounded, placing ample slices of meat between slices of bread. For the twenty-four hours soldiers could be killed with full bellies. For twenty-four hours consolidation work went on apace. Fritz harassed the workers and sent back a regular stream of casualty, but stalemate was again at hand.

In the dusk before dawn on the fourth day they moved out. It was not necessary to relieve them, for the most advanced units had consolidated a fresh front. The guns were ungenerous and spreadeagled many of the company on its way back to safety. With backs to Fritz and reprieve ahead, the nervous exhaustion of all triumphed. Each howling shell was menace for a naked back. Tired men ran, ashamed of speed, unable to restrain the emotion that hurried them onward. Having faced Fritz resolutely, they ran like curs from their own fears. They were nearly safely returned from a long journey. Who wanted to get caught now that they were being allowed to escape?

``Steady!'' yelled the Sergeant, just as nervous, but short in the leg. ``Steady!''

They sat in the old British line and waited for the stragglers.

``Might get a little leave now.''

``Don't be optimistic. They can do everything to a man except put him in the family way.''

The remnant of the Battalion gathered and marched down the road to Romarin in easy order. Smeared with blood and grime, covered with whisker, they had no answers for curious interrogation. Talk would come after sleep, when beer and wine loosened the tongue.

In camp they swallowed the hot food of live and dead men. Better still, they got the dead men's issue of rum, burning liquor that drugged overtired nerves to sleep. At evening they emerged from huts to bloat their paunches and relieve themselves, to have a further rum ration and sleep again.

Guy fought off drowsiness to read the story of the offensive in the \textit{Daily Mail}. It was verbally ``a crushing victory.'' ``Messines Hill commanded Flanders!'' Messines was the key to Belgium!'' ``Another turn of the key and the door to peace would be open!'' He had read it all before of every offensive. Being a private he knew that flat land could be held as tenaciously as hills.

``What does the paper say, Guy?''

``Heavy bombardment on the Western Front,'' he grinned, tossed the sheet to the questioner, and lay back.

Soon the sound of heavy breathing possessed the hut.


