\chapter*{\textsf{The incubation of chaos}}
\addcontentsline{toc}{chapter}{The incubation of chaos}

H\textsc{e} dug trenches, helped repair roads, carried riveting and wire on burdensome fatigues into trenches in front of Messines. Carrying was competitive, so that two hour's work spelled exhaustion for the day, whereas three hour's easy going would have left all hands fresh. He had to race and be not a whit less tough than the other men. Weaklings could lag. He would foot it with the strong or fall. If he wasn't strong he could endure. He carried bags of earth away from the entrance to mines, chasing the vanishing back of a comrade around a traverse, running full tilt into the back around the next traverse as movement sped, or jerked to a standstill. The earth had to be carried away from the mines so as not to betray location.

Warm weather brought increasing concentration of man and machine, a spate of new guns, of new gun positions, of new ammunition dumps. Balloons, aircraft, anti-aircraft guns multiplied as the heavens were delimited into zones of deathly propietorship. More balloons were festooned in one sight than had existed in the sight of all men in the pre-war world. Waspish planes dived and darted and stung and spat with nickel pellets, balloons turned to torches, parachutists swung in the heavens, other balloons ascended to replace casualties. As a million frogs come crowding from ponds when winter recedes, so life imbued with deathly intent came crowding to the sector on the heels of departing snow. Preparations which had trickled over the snow streamed over the slush, quickened to gallop over the dry roads.

Shells ripped out the entrails of horses and mules and scattered them along the roads, smashed horse and driver and limber and cobbles into paste. A squad ran and shovelled animal flesh aside, carried human flesh to the seeping hillside, dumped the contents of limbers on nearby fields. And behold the preparation crowded on more furiously to recover lost time, wheels ground particles of flesh and pools of blood into mud, into dust. Men punched on wayside piles of mule flesh and hurried it of so that the odour of putrescence might not offensively linger.

He saw a battery knocked out in the morning. Before next morning four batteries had taken up fresh neighbouring places. Dumps blew up. Two new dumps replaced each one destroyed. On, on, came man and horse and motor with shell, sand-bag, bullet, bomb, cement, road metal, and timber. Blood and iron converged from the ends of the earth to be incorporated in inferno. Military cosmos grew ambitiously pregnant with martial chaos, for the machine laboured to give birth to death. Men were infected by the velocity of concentration as shells are infected by cordite, grew ready themselves to be hurled into destruction to add to destruction.

Poplars along the roads budded in response to the caress of spring, but ungenerous war shot bud away, smashed limb away, hammered trunks down with high explosive. Fruit trees in a deserted farmyard tried to blossom and set fruit, but coal-box shrapnel shrieked and yelled satanically, and blossom, tree and orchard vanished. Barns, sheds, houses, were vomited up and came down rubble, dust. Green grass in foolish optimism attempted to repair the green carpet, but high explosive bit into and churned up the turf so that the carpet grew brown in spring, brown, with here and there a red patch, for the guns that refused a living to the green scattered the rich and the brilliant and the perishing crimson.

Yes. Spring in the salient from Messines to the sea was like spring in the frog pond. Another cycle of life and death had commenced. Everywhere heaven and earth reverberated and throbbed with activity as men, like moths, hastened to the flames. A railway line was fabricated so that an armoured train could bring long-range naval guns to where they could direct fire at backward areas. Short howitzers, squat, broad-throated, with their noses away from the earth, were place along the rail-side, howitzers that shot bolts into the ether from which they would fall with avenging force into Messines, bolts out of the blue that lent martial might an appearance of aid from the Almighty. Batteries, batteries, guns long and short and heavy and light, the dummy guns in wood to draw fire, abounded.

And concentration was only yet an infant. From the ends of earth order was hastening produce, from the four corners ships were hurrying with men, with food and steel, with romanticism and heroism and the conscripted dissenter. All roads led to the bloody carnival. For spring was about. Cosmos was indeed incubating chaos.

``We'll blow Fritz off the bloody map.''

``France will rock when the barrage starts.''

``Guns wheel to wheel if they were lined up.''

``Shells will bump one another in the air.''

``But is Fritz doing nothing?'' For Fritz was a tradesman at the game of murder too.

And Messines was still on its green hill, but not on so green a hill and not so complete a Messines. Yet scarcely a rubble-heap, for the guns were not yet belching in unison. Shells were falling over there, but the withering blast was not yet. And in Messines men laboured feverishly with concrete to turn foundation and cellar into pill-box, to shelter brittle Saxon flesh with iron reinforcement. John Guy could still forward at Messines across parapet if he didn't look too long or twice from one place.

Fury was straining at the leash, awaiting release. He had a brief exhibition of accumulating fury at Nienve Eglise. A New Zealander threw a bomb at a shadow in no-man's-land. The shadow threw a bomb back. A nervous German afraid of a New Zealand raid littered the sky with SOS rocket brilliants. Save our souls. Save our souls. Save our souls. Prayer not to God but to the vomiting gut of cannon. Deep in a bivvy three hundred yards from the front line Guy sat with his ear to a phone. He was the only steady man in the bivvy because he was the only one with a job to do. The Captain was nervous because he had nothing to do except wait passively, make decisions. Responsibility is easier in action than passivity. The German SOS barrage thundered down, smashing trenches, cutting the phone line to the front.

``Dead,'' said Guy. ``Dead as a door-nail.''

``And the battery wire?'' the Captain queried.

A faint buzz of inquiry came from the battery and Guy buzzed back acknowledgment.

``O.K. so far.'' ``Sir'' got lost in moments of excitement.

A Sergeant came down from above the the bivvy.

``Heavy bombardment on our front lines, Sir.''

``Any sign of our SOS rocket?''

``No. Only German rockets so far.''

But SOS rockets had been destroyed in the past by shell-fire before they could be fired. The Captain was worried. Before he could send a runner to the front a raid might be over. And the runner would scarcely get there in the bombardment, anyway. If he waited it might be too late to protect his men.

``Hell of a bombardment, Sergeant.''

``Yes.''

From the battery the signaller continued to call and Guy to answer. A huge explosion shook the air outside the bivvy and clods cluttered the doorway. The Captain, twitching with indecision, made up his mind.

``SOS,'' he said.

The issuing of the order gave him self-control. Once more he was at the centre of things, a directing instead of a vacillating mortal. Decision was like nerving potion.

``SOS.'' Guy didn't wait to buzz, but bawled the order at the telephonist.

At the far end he heard the recipient yell it excitedly. ``SOS. SOS. SOS.'' Guy nodded assurance to the Captain. And the straining fury was unchained. From hundreds of guns thousands of shells streamed into the night. And then the distraught, sweaty runner tumbled in from the front line, glad to be out of howling space.

``Well, what is it?''

``One of our bombs put the wind up Fritz.''

``No raid?''

``No, Sir.''

``Anyone hurt?''

``Five or six wounded. A shell landed in a bay.''

``Call off the batteries.''

So inferno lapsed into peaceful night, the world grew pacific. The quietness was accentuated by the rar-away pitter-patter of machine-guns. All around the area gun and machine-gun was cleaned and recharged and death was again at the ready. The Captain started to twitch once more. SOS had been precaution rather than an error. A financial fortune had been spewed out of guns in a few minutes. SOS. Guy turned the letters on his palate, his soul tasted their flavour. It was pleasant to know that he could erect a wall of molten steel between himself and the enemy by bawling that command in a phone. Down in the battery none would hesitate but all would haste up to cram shell into smoking breech.

``Get the casualties down before Stand To, Sergeant.''

``Yes, Sir.''

The quantity of preparation caught him into an elation. He enthused about impedimenta. Preparation piled on preparation acted like a spiritual intoxicant until men made light of part failures, grew lyrical about accumulation. Mechanism assumed a monstrous but convincing personality. Under the lash of guns wheel to wheel the German line must bend, must break. The fever of concentration got into men's blood and bones, reintegrated hopes disintegrated by past defeats. The fever was the higher as men knew that victory would come from their own corpses. What fascinated was what destroyed what it fascinated. Their blood was to be part of the mine that would erupt on a red morning. As shells did so would soldiers destroy themselves in the great moment of life.

``We'll blow the bastards off the map.''

``Yes. We'll blow the bastards off the map.''