\chapter*{\textsf{Pont de Neippe}}
\addcontentsline{toc}{chapter}{Pont de Neippe}

I\textsc{n} Pont de Neippe the Battalion billeted when it came from the line. They lived in a huge ch\^{a}teau surrounded by a moat of filthy water. In the rear of the ch\^{a}teau Guy found two New Zealanders at work on a monument of soft stone. Chips littered the ground.

``Hello! What are you doing?''

``Well, this stone is sacred to the memory of a bloody old Froggy who died a century ago. We're removing his name.''

``What for?''

``Put our own on. Bill Jones, his mark. Why waste a good stone on a rotten old Froggy?''

``Your not dead yet.''

``We soon will be.''

Cheerfully they laboured.

``I've heard of men digging their own graves, but not of pinching an inscription off a tombstone.''

``Why not?''

``Go ahead.''

He had no answer. He left them. He could see more humour than sacrilege. Soldiers often threw coffins out of vaults to provide safe earthy sleeping places. Corpses didn't flinch from high explosive.

And the chiselers didn't have the last say. A German shell in due course dug the ``bloody old Froggy'' up to contemplate the wreckage of his tomb.

Soldiers lost their sense of the sacredness of property. In a room in the ch\^{a}teau there was a mirror many feet wide and deep over a carved mantelpiece. A dozen men could see to shave in it at once. In the corner of the room was a richly brocaded couch, a grand piano, the only furniture left. The piano and couch fed the flames, yielded a moment's heat. The mirror fascinated Guy.

``Could do with a piece of glass.''

``Surely you wouldn't.'' Someone sensed his intention.

``Why not?''

``I don't know. I supposed it'll get smashed some time.''

``The Cullinian diamond was too big so they cut it up. This is only a mirror. Why not?''

``Go ahead for me.''

He went outside and got a pick. Some held their breath. ``Crash!'' The pick drove forward, and cracks spread from the impact like lines from the heart of a spider's web. He selected a good square edge piece for himself when the glass came clattering down. From the top floor to the basement men came to get a piece of glass. The Lieutenant came to investigate the vandalism.

``Who smashed the mirror?''

``I did, Sir.''

``Why?''

``Why not, Sir?''

The Lieutenant shrugged his shoulders, grinned, bent and selected a piece, departed, saying, ``Why not?''

They were never far from the clouds of black coal-box shrapnel that burst and took steady toll of traffic about Hyde Park Corner. If they were not in the line or at Neippe, they were living in bell-tents on the wooded rear of Hill 63. The canvas roof was safe, for Fritz did not shell the wood, but prodigious labour was being expended to provide strong bivvies against the sky when the enemy changed that policy.

At the foot of Hill 63, from Red Lodge to Hyde Park Corner, was a strip of road along which hammered incessant traffic.

``Why Red Lodge? The red paint on the lodge or the blood on the road?''

As if it mattered. But after living in that locality men invariably thought of blood on the muddy road when they thought of Red Lodge. Traffic streamed along, the thousands going and coming from the front lines, the tide of fatigue and preparation for offensive. Limbers rattled furiously along, no driver loitering, mechanical conveyances accelerated. The only conveyance that gave leisureliness to the traffic was the wheeled stretcher dripping blood as a fresh corpse was propelled to hillside clay. Shelling grew in frequency as spring advanced, more and more blood was splashed on the dying road. Fritz started to drop odd shots along the road between Hyde Park Corner and Red Lodge all through the night. Splashes of red were pounded and obliterated by steel and rubber tyre and hobnail.

On the flat ground near Hyde Park Corner there was a swampy graveyard. Stiff wet clay was close to the surface, and only in summer could decent graves be excavated. But however shallow the excavation, the corpse was nevertheless launched into a puddle. The water of the swamp defeated the rats until the ground dried out in late spring, but German batteries sometimes dug up mud-plastered bodies quicker than they had been interred. Rodent eyes gleamed appreciatively and bellies were glutted. When the cemetery was full dead men from the front lines were trundled along on the road to a hillside cemetery by Red Lodge. The wheeled stretchers would come with a sprawled human covered by a grey blanket, and blood might drip down all along the way. When a higher officer had been smashed down the stretcher might even be covered with a Union Jack, a flag somewhat discoloured by mud and red blood. There was never a flag when killing had been vast, for ceremony prevented expedition.

On the slope of the hill one of the greatest gambling schools in the world existed. Men from the bell-tents and far from any amenity, turned their eyes to two spinning, twinkling pennies that glittered in the sun. Their eyes were to the gamble, their backs to the ungainly dead men that were trundled along. The game of pitch and toss with bodies played by High Commands was ignored, as the bodies that would be future stakes hazarded on another game. Never a gambler in coin, John Guy loved to watch the play of emotions among the gamblers. Sometimes he would watch the living with an eye for the dead. The hundreds of voices would call the bet.

``I lay heads.''

``I lay tails.''

``Fifty francs heads.''

``A hundred francs tails.''

The school would grow excited when a courageous gambler dared. Optimists betted on the thrower of the coins, on ``heads,'' pessimists on ``tails.'' Sometimes a man would succeed daringly, doubling up and breaking the ring with a long run of heads, ``throwing a trot.'' The man who spun would be worth ten francs, twenty francs, forty francs, eighty francs, up and up until it was thousands of francs, for the betters were in hundreds if the contribution of each was small.

``Two thousand francs heads.'' The exultant note in the winner's voice who was a few minutes ago worth five.

``How much this time?''

``The lot.''

The ring would gradually set the gambler the full amount. Dirty paper notes, silver francs, copper centimes, all would be carefully checked by the ring-keeper. The coins would spin upward, fortunes would change.

Pat! Pat! Soft impact of coins on clay.

``Tails!''

What a roar of exultation as the two thousand francs came back to the crowd and the gambler who had ascended to the heights shrugged his shoulders and walked from the ring.

``Come on, come on, come on!'' the ring-keeper would yell as he gave the new chance to spin to the biggest better.

``Come on, come on, come on, come here in motor-cars, go away on crutches or wheel-barrows.''

Was it coincidence that the sentence should fall against the deafened ears of a dripping body going by beneath its blanket?

``I bet heads.''

``I lay tails.''

Coin flashing and money coming and going as unconcernedly as the bodies, eyes for pennies, backs for bodies. Cannon fodder had no time to concentrate on what had ceased to be cannon fodder, on what had become rat fodder.

``I lay heads.''

The gamblers were not callous. They were wise. Why bother about the irremediable? If an eye did chance to see, a shoulder could shrug indifferently, a voice mutter:

``Another?''

What was another to the million others to be? The man on the stretcher was lucky enough to have a grave. In the spring the dead would rot in shell-holes.

``Hullo, Joe. Broke?'' for Joe backed out of the school.

``Yes. They're killing a few to-day.'' Joe's eyes were on the stretcher. The corpse was sewn tight this time.

``That's the third.''

``The bloody hat trick.''

``Bloody's right. Hat trick's right.''

``Death throwing a trot. Ha, ha!'' Joe used the gambling parlance.

``A run of heads,'' Guy agreed.

One day a boastful burial party of one came along the road crying the news like a town cryer, eloquent of his novelty.

``I've got three in the sand-gab.''

``Jesus!''

``Three hands and a few spare parts.''

``How many graves?''

``Do you think I'm a bloody sorter?''

He buried the problem in a common plot, leaving the solution to Gabriel on the Resurrection morning.

``I lay heads.''

``I lay tails.''

What the hell did a problem like that matter? When a man was dead he was very dead.

But gambling in coin was frowned upon by High Commands, was a sin. A military policeman might break up the school altogether, though it would form again.

The stream of bodies grew and grew as spring exploded more and more cannon. The cemetery in the swamp was abandoned because German guns disinterred too assiduously, and all the bodies went to the hillside behind Red Lodge. Spring rains penetrated newly dug soil, and a red seepage came out of wounded bodies and moved down the hill, so that the hill itself seemed to know ghastly wounds. The rats could burrow in the hill, but the guns played the rodents a trick and started to kill faster than rats could eat. The sharp teeth and furred bodies were overwhelmed by the avalanche of dead heroes. And war was still in the trench.

``The rats will die of gluttony before the advance.''

A fellow left the gambling school with five hundred francs, and a black coal-box chopped him down on the way to its enjoyment.

``Hard luck. Never spent a penny.''

Getting killed wasn't hard luck. Getting killed with five hundred francs was.

``Who got the money?''

``Some lousy blighter on the Red Cross.''

``Damned hard luck. No wine, no mademoiselle, five hundred francs.''

``You've got to stop, you boys,'' an R.S.M. commanded the school.

``But two up's fair.''

``But, boys, it's the Head's orders, the Brass Hats.''

``The head I'm concerned about are on the coins, I'll lay heads.''

``Well, it's not crown and anchor,'' said the R.S.M., backing away tactfully.

``Crown and anchor's like war,'' the ring-keeper orated. ``It gets a man all the way. Gives him no chance.''

There were a lot of fresh stains on the road.

``Come on, come on, come on. Come here in motor-cars, go away on wings.''

``But I haven't a feather left, Bill.''

``I warned you, boys,'' the R.S.M. called back.

``And we heard you, Sir.''

``I lay heads.''

``I lay tails.''

`` `And he that throws three down into the field' ''---the melodramatic Omar elocution of Joe could not be repressed until he sprawled in death too.