\chapter*{\textsf{Etaples}}
\addcontentsline{toc}{chapter}{Etaples}

E\textsc{taples} was one of the greatest canvas cities the world has 
ever seen, will ever see. For in future wars men will not shelter under 
canvas but inhabit city sewers and cellars, so that adequate drainage will 
become a military problem. In Etaples the soldier was given his final 
training by sergeants on loan from the front line. Etaples was the 
toughening location for the hot-house recruit. In the Bull Ring the 
sergeants were not parade-ground militarists, but martial realists. Drill 
was a vivid business. Bayonet drill lost parade ground prettiness. To 
hell with steel points flashing and twisting in mid-air. Here steel 
points had to be directed at imaginary ribs, at guts and throat, the 
butt had to swing up at brittle jaw.

``Why the Bull Ring?''

``Because the sergeants are mad bulls and you've got to bloody well hop to 
keep clear.''

``That explanation will do, Robbie.''

But John Guy appreciated the purposeful reality of Etaples after the 
regulated artificiality of Sling. To become a killer was easier than to 
become automaton.

Filthy little Etaples was the leave place of the private soldier. Poor food, 
poor beer at exorbitant prices were provided for his delectation. Officers 
and gentlemen and privileged sergeants got leave passes for Paris Plage, 
a few miles away, where there were flaring brothels. The private was kept 
from Paris Plage on disciplinary and not on moral grounds. It was ``prejudicial 
to good order and military discipline'' for officer and private to stand in the 
same queues. The army could not admit any form of democracy even in passions. 
That would have brought the officer down to the herd level.

Officers' brothels and officers' latrines and officers' messes, sergeants' 
brothels and sergeants' latrines and sergeants' messes, and privates' 
brothels and privates' latrines and privates' messes are integral parts of 
military discipline. And the private may drink beer but the officer may sip 
at whisky.

In Etaples Red Caps hit first and enquired afterward, on the principle that 
it was better to break an innocent head than to have to break a hundred or 
two in a riot. Military justice must be swift, rough consequently, and 
frequently blind. Where men are fashioned to kill a little maiming occasions 
no inquiry. Flesh has few rights except that it has the rights of high rank.

Temperamentally against all governments, he was loud in protest against the 
prohibition. If he had no ambition to stand in the brothel queue, any place 
out of bounds had attractions. He was abnormally curious. Coming from a land 
where regimented, commercialized vice was unknown, he wanted, tourist-like, 
to observe the real thing. The bright and the wicked has glamour even for 
the virtuous. If the wealthy tourist goes to the wicked den to tickle a jaded 
palate, what an attraction Paris Plage had for palates so denied that they 
could swallow anything.

``To hell with the Brass Hats.''

Yes, but quietly, for in Sling he had learned a lesson and steered for safety. 
If he let his inward sentiments explode vocally, he did not let them influence 
his outer life, wherein he was like many a religionist.

One day, later in the war, a Red Cap in Etaples hit a private so hard that the 
private died, and although he was not the first, on this occasion comrades hit 
back with terrific violence. In the battle of Etaples which followed the 
incredible occurred: the casualties were military police. Sense of injustice 
too long endured, of men beaten up in guard-room and prison, flared up, and 
Red Caps went down to the grave like ninepins. Military communiques did not 
report that offensive, but over-night High Commands realized how near armies 
can come to revolution. A spark and tyranny of authority gives way to tyranny 
of the mob. Authority had its day when the fever expired. When the timid deserted 
the audacious, the audacious expiated on the wheel, in the cells. But headquarters 
learnt a lesson, and Red Caps were banished from the canvas city and reinforcement 
pickets policed the area.

In John Guy's day the hand of the Red Cap subdued the unruly. When an outlaw had 
been crucified to a wheel in the freezing air, or punched a few times in a 
guard-room, he became amenable.