\chapter*{\textsf{The road to Etaples}}
\addcontentsline{toc}{chapter}{The road to Etaples}

C\textsc{alais} was dark and wet. The cobbled road seemed to wander aimlessly. 
Marching in the dark to a strange destination gives movement a sense of futility. 
Drenching rain, muffled curses, the clatter of equipment, the weakness of the 
sea-sick, high, old-fashioned brick fronts, interminably they stumbled until 
an eventual destination seemed miraculous. And destination was point of fresh 
departure. They crowded into a train, so tightly packed in the dingy compartments 
that each man sat stolidly to resist encroachment on his precious space.

War is the mother of rumour.

``We're going to the base camp at Etaples.''

``We're going right to the front line.''

``We are going everywhere, nowhere, anywhere. Whom shall I believe?'' Guy asked.

``No one except a Brass Hat.''

``In future we shall never know where we're going, Robbie.''

``Only the consignor will know that.''

``We'll be dispatched for slaughter like sheep or cows to a meat works.''

``You've got it wrong. We're the butchers and the Germans are the sheep and cows.''

They had no right to try and guess destination for they were rankers. The officer 
commanding might know, but soldiers were merchandise without even an attached label. 
And it seemed they were not going anywhere at all, for after haste to board, the 
train remained at the platform hour after hour. Tired men became fretful at waiting. 
Sleepy voices declaimed haphazardly, like guns firing shots at night. But that was 
the war method of conversation. No searching discussion but an ejaculation of phrases,
of grunts, of sounds, words from throats that trusted utterance might find a billet, 
that didn't care if utterance spent itself on unhearing void.

``Here for the duration.''

``First seven bloody years are the worst.''

``They can do anything to a man except put him in the family way.''

When tired how irritating it was to be a trucked beast, to be delayed, or hurried 
without being aware of reasons. Whence! Whither! Wherefore! When! Why! Vouchsafe 
these crumbs to querulous intelligence and the way grows easier, for journey 
acquires meaning. One becomes an understanding part of the force that directs 
one's body. In time they would all grow used to movement with intelligence 
blindfolded, would come to relish any delay as respite. Who, long in France, would 
fret at delay on the road to meet a bullet, a bayonet point, an exploding shell? 
Indeed ignorance of direction came at last to spice the obvious with prospect of 
variety, came to lend colourful wings to rumour. But that was not so at once.

Rain went and night came, clear night brutally cold. The train moved from Calais 
to rolling downs, where a moon threw silver light upon deep snow. Giant pines stood 
silhouetted and threw long shadows over the cold blanket. Awake, Guy yet dreamed. 
Down under in New Zealand's January warmth, insects would sing to the moon, the 
fresh scent of cut hay, of wheat and oat sheaf, would cause nostrils to distend, 
warm air would invite love-making and rollicking beneath hedge and stook.

``Christ, it's cold!''

That reiterated phrase brought his mind from dreams of summer to winter's brutal 
realities. Night was so cold that none got down at stops to relieve bladders, lest 
movement expose the few warm surfaces to cold air. The pain of cramp and insanitation 
was preferable.

``Christ, it's cold!''

Even civilian life is crammed with such repetitions. Way is longer when destination 
and route are unknown, for the mind cannot humbug the body into passive endurance by 
telling it how much has been has been traversed. Very slowly the train moved across 
the frozen foreign lands. And the some man who had been before recognized a landmark. 

``We're going to the Bull Ring at Etaples.''