\chapter*{\textsf{Eyewash after chaos}}
\addcontentsline{toc}{chapter}{Eyewash after chaos}

``P\textsc{arade} for inspection in ten minutes.''

For all good things, such as the right to loaf around, come to an end. For a couple of rounds of the clock life had been disturbed only by meals. Dead men's rations had been eaten up, arrears of sleep made good, restocked \textit{estaminets} had been visited once more by the wealthy, cocoa and a biscuit had been bought for a penny in the Y.M.C.A. if means were slender. Maybe men had gathered in groups to refight the battle to possess Messines Ridge. Some few afforded the price of paper so as to read the meaning High Command and War Correspondents assigned to recent heroism. And now organization was on the way again.

``Fall In!''

The sorry remainder of the company came haphazardly from their hutments and fell in. Parade was not a tour of inspection by the Captain, but more of a greeting, a gathering of survivors. The Captain's service dated to the commencement of the war, and he was curious to see if there still survived an early face. He walked along in front of each platoon chatting pleasantly, pausing to inquire about a vanished familiar, bestowing words of praise, of cheer, upon everyone. It was a strange assembly. Men were caked with the blood and filth of battle, buttons were dulled, chins unshaved, the company was slack, indifferent.

The Captain's eyes sparkled with glee when he saw Guy splashed with red from head to foot, when he saw the dirty red smear on Guy's bayonet. He was sure the dirty red on the steel was German blood, and he was proud of Guy's achievements. He had read a dispatch in which Guy had been singled out for reckless gallantry, and he was sure gallantry had gone hand-in-hand with copious blood-letting. He was proud of every man who brought honour to his company.

``What's that?'' He pointed to the bayonet. ``Blood?''

``Yes, Sir. Blood.''

``How many 'Uns did you spit?'' In his gladness the Captain mislaid an aitch.

Guy smiled, disinclined to admit that the blood was of friends, that individually he had shoved a point into no gut. Blood guilt sat lightly on his shoulders when the carcase was not in front of him.

``Well, come on. How many?'' the Captain prompted, with a comradely twinkle in his eyes.

``Bayonet rusty through sticking it in a clay bank, Sir.''

``Oh, come now.''

``All I did, Sir, was to boot 'em on the backside.''

``Um. Um.'' The Captain eyed the stained bayonet.

``They say you did good work, but if you won't talk about it I can't make you.'' He passed on, smiling.

Inspection came to an end.

``Parade!'' yelled the acting Sergeant-Major.

That command banished goodfellowship from the Captain's face, so much so that the company grew serious, made an attempt to ``stand to attention.''

``At ease, Sergeant-Major.''

They stood at ease. The Captain harangued.

``You have had a rough time and you have acquitted yourselves as men. I am proud of the company and of the men in it. I am sorry so many well-known faces are missing. We are going out of the line for a week or two to reinforce, to refit, to rest and to drill. From to-night I expect slackness to end. At to-morrow's parade there will be a close inspection. You will get your packs and cleaning gear at once, and there will be no further excuse for muddy boots and uniforms, for unshaven chins, for dirty buttons. The company has been getting slack about saluting and special attention will be paid to this matter. Offenders will be harshly dealt with. The sooner we get back to normal the better impression we shall make upon our new reinforcements. It is your duty to make them understand that you were good soldiers in the line because you were good soldiers out of the line. Because of the detail we have acquitted ourselves well in Messines and in other stunts, and the war is not going to end for a little while yet. Saluting and clean buttons and a well cared for rifle are better for companies than undisciplined courage. Let me repeat. No slacking from to-day. Mud and blood off clothing and boots and webbing from to-day. I expect you all to be clean shaven from to-day. I am pleased to see so many good soldiers among the survivors. Dismiss the company, Sergeant-Major.''

``Parade, Dismiss!''

``Some say Good Old Captain,'' murmured a voice.

The response came stealthily form a dozen throats.

``And we say b------ the Captain.''

The following day the Battalion retired to a quiet village.


