\chapter*{\textsf{Enter fear}}
\addcontentsline{toc}{chapter}{Enter fear}

A\textsc{nother} wave assembled and moved on as the barrage lifted, on, on, on, to bomb, to bayonet, to die. He saw them go and only shrugged his shoulders. The desire to go with them or in their wake was non-existent. Madness had gone. The marvel was that it had ever possessed him. A few minutes before and life and death had been incidents in a game. But now life did matter and death was about to howl and pounce out of the sky.

``Dig!'' someone said. ``Dig for your bloody life.''

``Is it so urgent?''

``Fritz will have all this country measured to an inch. Taking the hill was easy. Holding the bloody hill, or living on it------'' The speaker left much to inference.

He started to dig as soon as he could dump his equipage. He drove the shovel deep into soft earth. As elation had driven him forward, so now a sense of impending retribution drove him downward into the safety of soil. He didn't pause to watch the surge of the advancing wave or to see whether someone found ten ticking watches. He delved. The very absence of shells, of rockets around his ears, unnerved him. For the moment German gunners were smashing up the advancing line. But still air was unearthly after chasing on the heels of detonation. A sweep of his eye showed him the ground in front of the pill-box he had subjugated populated with dead, face down dead, and as he delved he knew how stupid had been his reckless sense of invulnerability. He had been mad enough to believe he could deflect nickel bullets almost by shrugging them off. What luck had been his to survive his madness? It seemed so long ago. The act was clear, but the reasons for the act were drunkenly blurred. He had been drunk. He gave up bothering and dug and dug and dug. And made little progress. No shells came. Shells were for the last wave which was facing high explosive and uncut wire. It seemed he dug for hours but could get no shelter. The loose shell-churned earth fell back as fast as it was thrown aside. It seemed he would never get cover.

``Dig, boys, dig! Shells are coming!'' The voice was prophetic.

Shells came.

Zzzzzzzzzzz------Z! Crash! overhead.

A spinning fragment that had the range to a yard smashed the prophet's cheek. Blood sprang from a gash that ran across from mouth to temple.

``Good-bye, boys.'' The Lieutenant jumped for the rear with hand uplifted to the agonied face.

Zzzzzzzzzzz------Z! Crash!

The shell dug into the ground at his hurrying heels and hurried him on a few yards, where he fell down for ever. He lay crumpled and dead, killed by the shells that came to make good his prophecy.

Spades delved, earth flew, men sweated. A coal-box shrapnel, brilliantly accurate, split in the air above, so that all lungs gasped at the black smoke. Someone lost an ear and someone sat down to contemplate with a melancholy curiosity a gash of blood that ran down his chest.

``I'm off,'' said the earless soldier.

But he too found an accelerating shell which pitched him at right angles across the dead Lieutenant. The melancholy man with the bloody chest smiled, tried to speak, spat up a mouthful of blood, subsided for ever. The German guns, retreated in preparation for the well-advertised offensive, had the land measured to an inch.

``It's too hot here.''

They gathered up their equipage and ran forward to a broad abandoned German trench.

``No good, boys, Fritz will have this measured too.''

They retired thirty paces and set out to join a few deep shell-holes into an effective strong point. Again they could not have chosen a worse spot, for an old German trench ran from Messines to the broad trench in front, alongside where they dug. They dug in the angle of the T. Fritz expected trench and sap to be used, so that if proximity to cover gave an illusion of security in the event of counter-attack, actually it added hugely to danger. But shells seemed to be pattering down everywhere, and they had no time for rational choice. The other platoons were digging in away to left and right. They would have been far safer out in open shell-holes, far away from ranged targets, but men only achieve such balanced judgements in beer-house post-mortems. When wisdom and death coincide, wisdom is of no avail. The human family seeks danger under cover rather than safety in naked exposure. A shelled man dives under a sheet of calico and resists detonation by shutting his eyes.

``Someone's going to get killed,'' Guy said sardonically, as he spurred his muscles and his shovel to the task.

``Someone is. And the soft earth won't bloody well stand up straight. We can't get a parapet.''

Zzzzzzzzzzz! CRASH!

The speaker lost all concern about parapets, dropped, chopped about in a score of places, splashing the crimson.

``Someone has been killed,'' Guy murmured to himself.

``Up with him boys. He'll do for a bloody parapet,'' a wise survivor bawled.

``Bloody parapet's right.''

They dug and dug and threw the soft earth out over the dripping corpse, but the body sagged and the earth fell back.

``Take him a while to get stiff.''

They couldn't get a protective breast-work. All they could achieve was a deep hole with sloping sides, protection from sight but no impediment to falling fragments of shell. They joined two holes and crouched in fear. On the edge of a hole Guy placed his pigeons.

An Australian artillery observer came running overland, paying out a coil of signal wire as he ran. He came from an advanced battery to establish a forward observation post. On the way he had lost his comrades. He paused in the shell-hole to test his wire, but it had been chopped in a hundred pieces.

``Think I'll stay here a while. Man wants a bit of company.''

The basket of pigeons came sliding down into the hole after the Australian, and as the basket and the Australian steadied, a coal-box shrapnel salvo burst above. A piece of flying steel cut off the Australian's leg. As Guy opened his eyes from detonation he saw a mass of tangled flesh and pieces of jutting bone. And the heart started to squirt two streams of blood up into the air as the Australian writhed. The blood circled over and dropped through the wicker basket, painting the pigeons carmine.

``Quick!''

Speaking and jumping, Guy caught the jagged stump with both hands, applying pressure. Someone took a clasp-knife and cut the few strands of flesh that held the leg to the body away. The bootlace was taken from the useless boot and the leg was thrown into a shell-hole. Tightly two men tugged at the bootlace until the lace bit deep into the flesh and the flow of blood ceased.

``Think I'll have to bloody well stay,'' the Australian managed to say in his agony.

``Keep still, old man.''

``Any stretcher-bearers?''

``Not just now. Carrying someone out.''

``Will they be back soon?''

``Yes, I think so.''

``If they don't come soon I'm a gonner.''

``Nonsense. You've got a ticket for home.''

But it wasn't nonsense. He must die because he must lie out in a shell hole all day. He would stew in the boiling sun. He would get wounded again. His leg would gather the filthy battlefield mud. His only chance was to be carried away at once so that arteries could be tied and wounds dressed antiseptically. None could jump to save his life. It wasn't their business. They had to sit in a hole and get killed themselves. For what are a few lives to the gods of war, beside a few yards of conquered earth? To succour a comrade was a base desertion of all the offensive meant. Life was easier to abandon than military objectives. And bearers were few and cargoes were many. And bearers who struggled and plodded until they could scarcely move their own feet would have no time for the Australian. They had to make intelligent choice. They wanted to carry a man back with a chance of life, not kill themselves dragging back a man likely to die in transit. What the Australian knew was not nonsense.

``Wooden leg of blue gum.''

``Not even a jarrah coffin.''

``You're all right.''

Easy to say ``You're all right'' if your own life wasn't ebbing drop by drop despite the cruel tourniquet. Belgian and French soil were dry and hungry for blood. It took a large quantity of blood to stain a little earth, but a very little ebbed could ooze out a lot of life. Drip, drip, drip, all day the leg would go, and never an odour of native wattle or eucalypt to soften the stench of high explosive.

Zzzzzzzzzzz! CRASH!

Guy was dry. Guy was scared. He took a swig at his water-bottle, precious liquid to be carefully apportioned. There was an Australian everywhere, in every gambling joint, brothel, Y.M.C.A., \emph{estaminet}. And he watched one die. But he couldn't concentrate on the Australian. The guns wanted more than a leg or two. They wanted corpses, piles of corpses. They wanted his life if they could get it. Shells, shells, shells, rained down and split the sky above. And he twitched and he winced and pressed himself against the earth and awaited his turn.

Out of the hot sky a shell seemed to scream and strike at his core, seemed to try and get to the centre of being before it exploded, so that he would fall completely asunder. He cringed down. He shut his eyes. His heart stood still to see whether it was worth while starting again. He was buffeted by a ripping, splitting concussion. He heard the hurrying swish of travelling fragments. Cautiously he opened his eyes, as if keeping them shut to the last moment prolonged security, as though frail lids could break the momentum of steel. He erected himself and took stock, and he was in the midst of the dead. Four men were dead. One had a dented head, one a foaming throat wound. And he was splashed with blood and unharmed. A living companion moaned at the pain of a deep arm wound. And the Australian had got rid of the bootlace and the gusher arteries were giving the pigeons a fresh bath. Somehow he got the Australian's leg tied again, but life had dimmed a little more.

``Good-bye,'' said the wincing arm casualty. ``I'm off.''

``Congratulations.''

But as he topped the hole a shell howled in fury at him, and dropped him back dead, heavily across the Australian. Guy dragged him away.

``You poor Aussie bastard.''

The Aussie smiled feebly at his commiseration.

``Any bearers?''

``Soon now.''

``I wonder.''

Distracted with fear, Guy crouched down amid the dead and awaited his turn. And then he decided to move. He must have company. He couldn't prevent the gradual ebb of Australian life. He couldn't drag a one-legged stalwart to a safer place. There were no safer places. He picked up his pigeons and tumbled into the second hole. With him went bad luck. A coal-box salvo followed him. And again he was unscratched and death was all around. One survivor lay with a leg split longitudinally from ankle to hip. Guy tied the leg together as best he could, his dirty, earthy hands penetrating into the warm, slippery wound, so that he had to wipe them on his trousers.

``Any bearers about?''

``Soon.''

The questioner knew that the answer was a lie, always knew because as like as not he had just been lying to some other man.

``Soon, you say?''

``Yes, soon.''

The questioner grew quiet and grim in the concussionless moments. Between each stinging lash of exploding steel he had time to wonder whether he would journey to first aid or to eternity. And the human family was too busy killing to bother about his salvation. The hot sun beat fiercely down, its burning rays in league with the guns.

``Any bearers?''

Three men went down at the next stroke. Three walking wounded reduced to two when a savage shell reached out and tore an arm and life from an escaping victim. And Guy was not hit, except by clods and pieces of flesh. He was covered with blood, the blood of a dozen, his face smeared where his sticky hands had brushed. His buttocks were red with blood, red with earth. His pigeons, as tenacious of life as himself, were red, red with the blood of men dead and dying.

Once very long ago, after zero---and that seemed a hundred years before---he had been swollen with an heroic madness, with the blind mania that sends an imbecile at a concrete wall with lowered head. He had believed himself powerful to bend and break the German line, to conquer an army corp. He had been garbed in and impregnated with irresistibility. But that was a past epoch. Now he was timid, puny, frail, so brittle of body and spirit that he cowered back before the crash and yell of each bursting coal-box. And shell after shell hammered down at the head of the sap. German gunners had the range to an inch. In the detached area of the battery were cool logicians aware that wretched men would leave the exposed safety of open country to cower around the illusive protection of a sap. So all day exploding shell left fresh corpses, rent the corpses that were fresh. The survivors should have realized that comparative safety lay sixty or seventy paces right or left, but survivors were too busy twitching and grovelling and bandaging to have time for nice logic, too scared to stand erect and see where the safe places were. For the howl of shrapnel with death in each burst disintegrated thought before dismembering the body. Cringing to earth, what else could men cringe to? And for all the flinching men knew death was as alert to right or left.

Into a third hole away from the dead Guy went with a survivor, and again the sky gave of its thunderbolts and he was alone with dead and his pigeons. Tortured by fear, demented by hellish concussion he sat, and the argument of Romarin came back to him and he jabbered satirically to his pigeons:

``I think the angels are guarding you and me.''

Maybe they were.

In a lull he discovered four men in a very narrow piece of sap, each man on his hands and knees in terror. There was a deep, narrow place for him, so he took his pigeons and lay flat against the earth. The sap quivered under the hammer-strokes of high explosive and shrapnel, and he quivered in response but was unharmed. He did not bother to try and look beyond and see whether the last wave had attained its objective. High strategy had lost its meaning. He had forgotten all foolishness of that sort. What mattered was to continue to live. He heard a voice coming from the man nearest to him.

``You mustn't be afraid.''

The voice must have come in a lull.

``You mustn't be afraid.''

If he was scared, something told him, he would never sing out about it like that. But how many men and women have prayed to themselves, standing apart and trying to quell the monster within. He took up the refrain once.

``You mustn't be afraid, my pigeons.''

And then advance must have ceased, for the bombardment of back areas intensified into steady and obliterating drumfire. He shook and twitched as the song of concussion came and went as does the rolling of giant sea. The roll of explosion was ever there but it slowed and quickened, intensified and subdued. Ra, ra, ra, ra, deep-throated as though earth was groaning in agony under the iron lash, went with the song of the guns. And the hero or madman was afraid for his carcase. Canisters of steel ripped and split and scattered themselves and all else about. He who had wanted to go on to Berlin and shove his bayonet through the Kaiser's neck, who had wanted to kick and destroy and trample down writhing bodies, he was skulking in fear at the bottom of a narrow trench. Perhaps the Corporal was right and he was yellow. But if he was afraid it was of shells, of the fury of men and not of God. Strange thoughts came as he heard the voice again.

``You mustn't be afraid.''

Oh, yes. John Guy was no bloody hero after all. He was afraid. He was abjectly afraid. He was losing the self-respect that would cause him to conceal his fear. And he heard a yell from his neighbour that chilled his spine. He jumped to look. Mournfully the fellow looked at his abdomen. A splash of red spread outward from the point at which a piece of shell had entered. The dying man clutched at the straw that all reached for.

``Any bearers?''

``Directly, old fellow.''

He stumbled back again to his pigeons. He couldn't do anything and he didn't want to see death creeping down. From his funk-hole he heard the new prayer the dying man was muttering to himself.

``Any bearers?''

How was the advance going? He didn't care. He didn't care what happened to a few yards of earth. He didn't care whether the war was lost or won. He cared but for life. He cringed in perpetuity before the shattering force of murderous drumfire. He didn't want the top of his skull lifted off. He didn't want his leg chopped off. He didn't want to look in sorrow at a hole in his guts. But a shell would get him. A shell must get him. A shell had got everyone. It was a matter of time. A little patience and he would be dead. He was sure high explosive would search him out. As he became sure he became more resigned. Him and his bloody pigeons. Suddenly he was too weary to bother about it all. There were limits to fear, to the power to cringe back from concussion. The poor devil was still mumbling. Gut wounds let life out too slowly.

``Any bearers?''

The voice was weaker. Death was creeping down like sleep upon a drowsy child. And sleep was creeping down on Guy. Shells were falling. Ra. Ra, ra, ra, shells, shells, shells, shells. His shell was coming. Shells. His shell. His very own shell. Ra. Ra. Ra, ra, ra. Ra, ra, er---ip, er---ip. Ra, ra, ra, ra.



