\chapter*{\textsf{Hell-fire for orators}}
\addcontentsline{toc}{chapter}{Hell-fire for orators}

M\textsc{artinets} never tire of proclaiming themselves good fellows. 
Judge, jury, and lawmaker, for there is relativity about military 
offence, they proclaim to all the world that military courts from 
orderly room to courts martial are the fairest in the world. It is 
insubordination for the offender to express a contrary opinion 
publicly. After all, most dictators believe that they are jolly 
good fellows.

Outside the orderly room culprits stood to attention. Inside the 
Colonel kept his kidneys near the warm fire. He frowned. He was 
annoyed. The amount of crime was high. Why couldn't men devote 
themselves to King and country without shirking? Men were parading 
with buttons undone, were leaving their gear out of alignment in 
the hutments, two had omitted to shine buttons, one to shave, one 
had a dirty rifle and a dirty spirit, one had called a Corporal 
``Jim'' and added, ``You're a sneak,'' for good measure. Two had 
been caught talking in the ranks. The army was going to the dogs.

Outside the orderly room Angley, master of ceremonies, pranced. 
His eyes danced with glee when he sighted Guy.

``What! Again!''

``Yes, Sir.''

``Yo ho. I knew I would see you again. You're the type. What excuse 
have you got?''

``None, Sir.''

``Um. You're learning. You're in the army now.''

``Obviously,'' Guy thought. ``Obviously, irreparably.''

The Colonel loved warm kidneys but he had to make justice summary. 
Delay irritated him. He hated argument, liked the fellow who 
capitulated with, ``Nothing to say, Sir,''  and took his gruel. Why 
should any offender object to a few days C.B.? There was no time 
for rambling procedure. There was a war on. Men had to drill, not 
to argue about it. Even if innocent the punishment was light, or 
so it seemed to the Colonel. A certain amount of criming was good 
for discipline.

Outside in the cold air the sentenced whispered, ``The Colonel is 
in bad fettle this morning,''

``Two days C.B.! Three days C.B.! Forfeit a week's pay!'' The Colonel 
assessed each crime and decided its penalty as the charge was read, 
although he went through the formula of asking the culprit what he 
had to say. The Colonel sat in relaxed comfort while the offender 
stood to stark attention. Alertness could shorten, dreaminess lengthen 
the sentence. Polished arguments were not as important as glittering 
buttons.

``Private Guy.''

The sharp order that his name constituted jolted him to alertness. He 
had been trying to piece together a good sentence, but the necessity 
for physical starkness disarranged his ideas.

``One step forward march! One! Two!''

He found himself with an escort on each side. Stepping and wheeling and 
turning, he was suddenly in front of the O.C. The charge was read out.

``Anything to say?'' The Colonel's attitude requested ``No, Sir.''

How unfair it was. The Colonel should have said, ``By numbers, say.'' 
Concentration on martial precision had scattered Guy's thoughts. Guy 
opened his mouth, relaxed it, raised a hand, lifetime's gesture to 
excited expression.

``Sir.''

But Angley's voice shot down his offending gesture.

``Stand to attention. Put that hand by your side!''

Guy started to poker-like rigidity, body in perfect order, thoughts 
in chaos.

``Sir, I was at the orderly room, and I thought------''

``Never mind what you thought. Give us facts, not a speech.''

``Yes, Sir, I------''

The hand wandered again from the seam  of the trousers.

``Stand to attention!'' roared Angley.

``What were you doing in the orderly room?'' The Colonel took the 
initiative.

``Found him in bed after reveille, Sir. I dressed him down for 
slacking.''

``But, Sir------'' The hand would move with the vocal chords.

``Stand to attention!''

``Was your rifle dirty?''

``Yes, Sir, but------''

``Never mind the but. Did you argue on parade?''

``Yes, Sir, but------''

``Silence!'' roared Angley.

The Colonel looked so thoroughly annoyed that Guy sensed danger 
in continuance. His persistence was an irritant. He should have 
kept quiet from the start. There is penance and absolution for 
military sinners, but hell-fire for ranker orators.

``UUM---making a bad start. What other convictions?''

Guy gave a start when the orderly read his list. He couldn't say 
that at least two hundred men had been in the Capetown disturbance.

``Insubordination and mutiny at Capetown, Sir.''

``We can deal with men of your stamp. You can't do as you like 
here. A little pack-drill and guard-room will complete your 
education. Two days in the guard-room and forfeit a week's pay. 
If your sheet had been clean we'd have treated you lightly.'' The 
Colonel dropped his eyes to the next charge sheet.

``Left Turn! Quick March!'' Angley roared.

``How did you get on?'' Guy's mate in the file whispered.

``Top weight, I think.'' The sporting phrase came to his lips.

He was marched away to the guard-room, searched, relieved of his 
equipment, thrust into a cell. He sang with mock cheerfulness to 
ease his bitterness.

\begin{verse}
``Rule, Britannia; Britannia, rules the waves. 

\ \ Britons never, never, never shall be slaves.''
\end{verse}

He was dazed. Everything has occurred so swiftly. A mild succession of 
mishaps and the disastrous Capetown conviction had supplied overwhelming 
evidence of recalcitrancy. He was a deep-dyed military criminal. But 
Angley and the Colonel were swine. And the sore spot in his chest was 
no better.