\chapter*{\textsf{Any bearers?}}
\addcontentsline{toc}{chapter}{Any bearers?}

A\textsc{nd} then he was no longer afraid.

Fear, like heroic exultation, was transient. The body had its limitations. Fear had been the product of uncertainty. Fear had been occasioned by chance of life. But life was gone, definitely. Soon he would be dead, and there was no uncertainty about that. He must be killed, for he was no different from other men. This new certainty brought profound relief, an acquiescent passivity. Why struggle? He would just wait his turn. He grew sorry, rather than afraid for himself. Fear had ousted exaltation and now weary resignation ousted fear.

``Only shrines are protected by angels.''

The thought expressed in the words gave him a moment's perverted joy. How he had routed the stout padre! Shrines were safe while men were battered on the bloody anvil. The parson's God had capitulated to Satan. Parsons were medicine men. Exulting the spirit of the bayonet. God was the god of battle. He, John Guy, soon to be dead, wanted to run God through the innards with a bayonet, bomb the serenity from the heavenly countenance. If shells made him yellow he was not afraid of God. He drove his knuckles against his head for reasoning was getting sluggish. But parsons were dirty liars. Shells. Shells. Shells. Detonation succeeded detonation, the steel ripping, ra, ra, ra, ra, ra. The earth quivering beneath his flinching body, for the flesh still cringed after the mind had resigned itself. Cosmos reverting to chaos. He sipped water and talked to his pigeons about the parson's God. Shells. Shells. Shells. Shells. Ra, ra, ra. Ra, ra, ra, ra. Ra, ra, ra. Ra, ra, ra. Ra, ra, ra. He had no sense of time, no fear of eternity. He had lost all interest in battles. Shells, shells, shells. He talked to his pigeons, two birds bathed in drying blood.

And in the midst of reverberation and ready for death, he fell asleep. The sleepless night before, the hysterical advance beneath the crushing burden, the ebb of nervous energy. he was fearfully tired. He fell asleep ten thousand times. He fell asleep between the concussions. he fell asleep, instantly asleep, to be instantly jarred to wakefulness, to fall instantly asleep again, catching precious split seconds of rest between violent awakening. In some lulls he fell asleep for fractions of a minute. Once he opened his eyes and felt refreshed and had a sip of water. There was plenty of water in the bottles of dead men. His pigeons seemed little worse for wear. He nibbled a piece of chocolate.

The water and the chocolate restored his faculties slightly. He stood and looked ahead. There was a terrific barrage half a mile away. That was why drumfire had slackened around about. Men were being butchered under the smoke. He felt so safe that curiosity or humanity drove him to call up his surviving companions, to pay a visit to men dying in shell-holes. He hurried lest Fritz start again with the coal-box.

``Any bearers?''

``Directly, old fellow.''

``I know all about that.''

``You'll soon be in England.''

``Be in heaven.''

``England's better.''

If his lying brought no faith his presence brought comfort. Dying men got tired of inhabiting shell-holes with dead men. They would be dead long enough and wanted living companionship. The sun blazed down on naked wounds. Men died in silence or grunted and groaned in intolerable pain and achieved heroism not by mouthing noble precepts but by smiling at living comrades. The grunts were involuntary. Giants toppled by thunderbolts showed a grim resolution for comradeship's sake. \emph{Esprit de corps} in the teeth of gathering darkness. Such grunts as passed their lips were beyond the power of man to subdue.

``Any bearers?''

He stood beside the Australian, who looked at him from a face pale through loss of blood, drawn with suffering.

``Soon, Aussie.''

He did not say ``No bearers.'' The Australian's life was dimming and the evasive negative was half-humorously accepted. Maybe pain was escaping as well as life, easing death. Perhaps as flesh lost its hurt the Australian could dream of wattle and eucalypt.

``Say, Enzed.''

``Yes?''

``When my fellows come tell them I missed. Napoo Mademoiselle from Armenti\`{e}res.''

A faint grin seemed to conquer pain altogether as he said the words. And then he wriggled and rubbed his stump in the mud and died. John Guy reached for the Australian's pay-book so that if a miracle occurred and he survived he could write to the Australian's home. For the dying man had been cut off so far from his fellows. But a coal-box shrapnel yelled again and he abandoned the breast-pocket to dive for his funk-hole, where he talked to his pigeons. ``Tell them I missed.'' What an absurd prelude to heaven.

The guns had started again on the backward areas with their rending ra, ra, ra, shattering earth, shattering men alive and dead. Shells, shells, shells, shells. Again he fell asleep between concussions, but ra, ra, ra, they went, forcing him back to wakefulness. Sleep, concussion, sleep, concussion; he had lost sense of time and place, care for victory or defeat, concern for his life or fear of death. Shells, shells, ra, ra, ra---ra, ra, ra---ra.

He awakened to another long lull and ventured again. He found three men on the lip of a shell-hole where they had fallen hours earlier. They had been peppered with nickel.

``Hallo.''

``Any bearers?''

``Soon.''

The three were too weak and near to death for them to shade their faces and heads from the pitiless sun. Away from the sap they had lain for hours, intimidated by concussion and scorched by heavenly rays. Safer than if they had staggered to the sap but sure to die.

``Any in sight?''

``I'm afraid not.''

A man gave him back a sigh. One had a scar across his forehead. Foreheads have compelling affinity for bullets, but this bullet had slanted across instead of going mercifully straight. All had bullets in the guts. One had an arm torn off by a shell fragment, the amputation so contused that life ebbed with wretched slowness. They knew no pain. All had gone beyond pain except when they moved. Their steel helmets were sprawled about, their heads exposed to sunrays. Who could clap steel helmets on such heads? They were all dirty yellow of face.

``No bearer's any good to me. It's too late.''

``Nonsense.''

And despite lack of an arm or a tourniquet life stayed until sundown.

``You'll get a wooden arm.''

One of the men with a belly full of nickel had to be the ex-Salvationist with whom Guy had argued so frequently and violently about God, his friend and bitter verbal enemy. Why should he have to wonder if the dying man was still brimful of faith? Greasy sweat was on the waxy face and the lips were parched and dry. His warrior God had got him peppered in the guts.

He started to move one of the wounded from the exposed lip to the bottom of the crater, but his action restored cruel agony to the numbing body so he left all in peace. After all, what did it matter whether they died slowly on the bottom or were hurried away with a burst of shrapnel on the lip?

``A drink of water,'' the Salvationist broke his silence.

``Sorry. A drink will kill you cruelly.'' The man's guts were perforated.

``Just a drop.'' On the road to heaven the Salvationist could feel the parching pain of hell.

They all clamoured and he shut his ears. Water might leak through the guts and restore terrible aches. Water wasn't good for abdominal casualties. They would die anyway, but there was always miraculous chance.

``Just a mouthful.''

He shook his head.

``I'd sell my soul for a good swig.''

``But I can't let you sell your life.''

He moistened a rag and wiped the black lips that cracked in the sun.

``Thanks.''

Good God, they thanked him.

The drumfire came again and he went back to earth.

``You'll tell the bearers where we are?''

``Yes. Yes. Of course.''

But they were to die in the evening because the were too far from medical aid. The mercy of bearers was infinite but their strength was human.

He crouched down again, caked with blood, while the earth squirmed under the ra, ra, ra, ra, ra, of ripping steel. And he thought of his argument about God a few nights before. And he hoped the man beyond had no doubt. To be weighted with doubt at the crisis would be awful. He hated the warrior God. He did not hate the Pope's God, for the Pope wanted peace. Some day, if ever he had time, he would think about the problem again. Some other time when life was not so browbeaten. Shells, shells, ra, ra, ra, ra, rararara, ra. Shells and moments of sleep got intermingled. Sometimes he would mutter inanities to his pigeons. He was not afraid of shells when half asleep. They were bound to get him. Shells fell for a thousand years and he slept between each impact and concussion. Ra, ra, ra. Shells. Ra, ra.
