\chapter*{\textsf{In which the infantry have a jolly good time, parley voo}}
\addcontentsline{toc}{chapter}{In which the infantry have a jolly good time, parley voo}

O\textsc{ut} of the pay-office and along Southampton Row he swung with a vigorous step, thrilled with the glory of success. So happy was he that instead of crossing the road to avoid him, he saluted a Provost Officer with all the punctiliousness and comic deliberation of the drunk. His financial plunge had made the hectic high finance of millionaires seem cheap. Millionaires played with and risked their money, juggled with and multiplied figures in a book. But each Bradbury note to Guy was a title to life. He had gambled and he had won eight days of life. Millionaires grow richer and more dyspeptic. He had eight days, money, gargantuan appetite. He had won the very right to breathe, to eat, to exist, to buy love. For eight days he could run from the nervous sickness that dogged his progress. But no, the Bradburys couldn't cure him at once. His joyful emotions melted suddenly into tears and he had to stand and appear to examine a shop window. What a goat he would appear if he walked down the Row crying. And he didn't want to report to a doctor now until his leave and cash ran out.

He bought papers and devoured the morning news. He want again to the barber and lounged back to enjoy the keen razor. He had a cup of tea and a few biscuits. When the pubs opened at midday he repaired to a bar and drank two glasses of heavy stout for the purpose of slowing up the emotional reactions that precipitated him through nervous crisis to tears. He used his stomach to steady his soul. He lunched well in a cheap restaurant, a sixpenny tip overcoming all meat ration restrictions. He spent the afternoon aimlessly wandering, enjoying London by following his nose. He was in the mood to admire everyting, even the Albert Memorial, although he knew that according to the rules he was not supposed to admire that. Still, he didn't know what the rules were. He had read somewhere that it was to art what a fireman's blackpan is to a saloon dinner. That it contained a little of everything. However, he was like the fireman inasmuch as he was hungry for any fa\.{c}ade not built of sand-bags. Primitive souls are easily satisfied. His admiration was really occasioned by his state of mind, and that made the world a purple land. He was embarassed in his savagery and contemplated the memorial with frank joy, unconcerned about the cultivated verdicts of ``Folk who Knew.'' Maybe his sentiments were affected a trifle by the midday stout.

He had the evening meal at the Club and did not book a bed. Although he had sought no bed-mate during the day, he had ambitions to find such another as the jolly girl who had shown white teeth on his apple the previous evening. Such a one might come to him out of the shadows. Yet when the shadows fell he resisted all accosting and went to the Y.M.C.A. at the Adelphi.

The Y.M.C.A. was dull. Apart from the fact that it was officered by women and not by men, it seemed a part of France,  of the Army. He sat at a desk and wrote a letter, and then, tiring of the scratch of the pen, of the effort to prevent his mind from escaping to the world outside, he threw the pen down, toyed with a cup of tea, and got out as someone started to distribute hymn-books. Walking toward the Club he chummed with another New Zealander going the same way. As he chummed two girls came out of the shadows toward them.

``Good night, boys.''

``Good night.'' They both halted.

Each girl appropriated a man. Guy's chum accepted his fate and moved willingly away. Guy peered at the girl to see if she was plump and fresh and jolly like the one that remained in his mind. The girl was frail but not bad-looking. She took his arm and set his feet moving as though she had proprietary right to direct him.

``I say------''

``Come on.'' She clung fiercely to his arm, although hers was the initiative.

Unwillingly his feet moved. He wanted the right to exercise a choice of sorts.

``I'm not dangerous.''

That unnecessary phrase set his feet moving more readily. Dangerous. One rough shake of his arm and she would be gone. She was slender, average height, pretty in a way. She was brunette, not yet twenty years old, and she had what he believed was a Cockney accent. She was a wisp of a child to be street-walking, frail enough for him to have his doubts. But her grip of his arm had about it a desperate possessiveness. She was not yet old in the profession. She wanted to get quickly off the streets and with a meal ticket. He was bread and butter, clothing and rent. Nor was he coarse and drunk but seemingly kindly, so that bread and butter might be fairly easy to earn.

She was depraved, dangerous. He chuckled at the thought as he contrasted her clinging possessiveness with himself in full fighting kit.

``What are you laughing at?''

``Cold night.'' He ignored her question.

She misunderstood his reference to the night. She thought he wanted liquor.

``I don't want anything to warm me up. Madame will give us a cup of coffee.''

Her pal had different ideas. Guy heard her telling her man that there were certain essential preliminaries, a show and a few drinks.

``I'm not a cavalry horse to be saddled up at a moment's notice.''

So the other two went looking for amusement and stimulation. Guy got familiar and put his arm around his companion's waist as they went their way. He insisted on pausing outside the soldier's Club. He had some business to attend to. The frail little girl might be a decoy for a very tough place.

``One moment. I'm going inside.''

``Don't run away. I'll make you happy.'' She was afraid he was quitting.

``I'm coming back.''

``What do you want to go in there for?''

``But I won't be a minute.''

``You're giving me the slip.''

``Here. Here's five bob as a guarantee. Wait two minutes. Now don't bolt.''

``Well hurry. It's not good for a girl to loiter.''

John Guy ran up the steps of the Club and handed over all his cash except two pounds and received a receipt. He had eight days to go. The office attendant smiled with an awareness that irritated him. He ran into the gloom again, and the girl caught his arm and hurried him away. She guessed wrongly at his reasons and admonished him.

``You want no safeties. I'm quite healthy, thank you very much.''

He soon lost his sense of location and she brought him to a terrace of very high houses.

``In here.''

They walked up the steps and entered the door and went upstairs to a warm, comfortable room. He caught and kissed her in the dark on the way.

``Now then, passionate. No hurry. You've got all night. You diggers have no patience.''

``We have to wait for years.''

``What's your name?''

``John. And yours?''

``Ann.''

When the light had been on for a few minutes a maid came to inquire what they wanted. They ordered coffee and it was brought to them as they sat side by side in bed.

Perhaps it was a night's pleasure he found. Certainly it was a night's companionship in a clean bed. He was so tired from walking pavements, so mentally tired from the strain of arriving penniless and making a foray on the pay-office, so weary of war, that he could not have made it too hard for the frail girl to earn a living. Profoundly he rested, a long night of dreamless sleep. For Ann shook the thunder of guns from his senses. What did matter was that the little creature alongside him was a woman, that a woman's company was something a man had yearned for, not merely out of lust but out of the will to live. Very little of the night was spent in other than innocent, life-restoring sleep. And after that deep night he felt profoundly grateful.

``The trouble is you men hate us afterward.''

``Don't be silly. You're a good pal.''

For after a night of sleep rather than of passion he fell in love with her almost as a big brother might. While she protected him from the thunder of the guns he protected her from the law and the hunt for new clients. He grew sorry for the girl who trudged the pavements, afraid of police, afraid of men who turned out to be swine, happy when safe in her room with an average man. In that act of compelling him to be sorry for her Ann achieved a healing miracle, for she separated him from the morbid brooding about himself that had been his lot for months. He grew to worrying about her future, becoming far more concerned about her than she was for herself.

She was contented as she lay in bed chatting next morning. She put her arms around and tried to kiss him to silence when he started to ask her questions and probe the past. She was used to last night's rake growing moralist and giving good advice in the morning. She didn't mind making him a father confessor if by doing so she could secure his financial goodwill. So she was sweet and informative for business reasons when she found he was just over from France. She wanted him by the week. He wanted to enfold a woman. She was prepared to be encompassed provided she had her arms around a meal ticket.

Maybe all men do give advice to the so-called fallen, bid them mend their ways, in the morning. Morality and high-sounding precept coincide with departure. Maybe a good deal of other regarding is just an unconscious effort to assert propietorship over even a prostitute. The sinner would have her be moral, to other men. And Ann was prepared to assent to any proposition. Her philosophy was pliable when the bed was warm and the day's rent assured. Nevertheless he was not inflamed by her eyes or lips or breasts or her ardour. She was restful. That was enough.

``You do ask questions,'' she protested.

But she told him her story. A waitress turned munitions worker because munition work was patriotic. The work had been too heavy for her and illness had reduced her earning capacity. A girl friend who had also been patriotic and who had left the streets to make shells. The girl friend had missed the odd excitements of prostitution. Patriotism was as dull in a munition factory as in the trenches, a grey business, and as poorly paid except that one was a shareholder. And overseas soldiers with money came pouring into London. Canadians, Australians, New Zealanders, all drawing allowances that were munificent alongside Tommy's poor dole, wanderers who had left their women but not their sex on the other side of the ocean. The advent of men on leave who came to enjoy respite rather than visit homes made prostitution easier and more profitable.

So Ann's pal had gone back to prostituting herself to make soldiers happy after prostituting her talents to aid in their dismemberment. Maybe the shell-making was the least human form of prostitution if the most honourable. She had persuaded Ann and thereafter they had hunted together and shared the good and bad luck. And Ann had found more happiness in prostitution. She was healthier, happier, better fed, better clad, slept in warmer if in more sinful beds and, unlike civilians, soldiers were not given to sneaking off in the morning without paying. Occasionally they stayed for a week, taking in music-halls and other places of amusement, using the girl as guide to London.

``You'll stay a week.'' She pressed herself closely to him.

``We'll see.'' He tried to avoid her.

He was not sure he had the desire for a week. Certainly he had not the money. So he hesitated. Still he was disinclined to make a straight refusal. He talked instead about her profession, more anxious to probe the secrets of her mind than of her body. She told him that many boys in khaki were innocents but determined not to die in innocence. She was afraid of drunks, but sometimes there was no choice. Drunks made a mess, attracted the attention of the police, annoyed the landlady.

``So you'd rather make love than shells?''

``Stop asking questions.''

``But would you?''

``I'd rather make love than talk about it.''

But he was curious. He came from a land where commercialized vice is rare. He wanted to know. She yielded to his persuasions and told him of her experience, of other girls, of girls who decoyed men to places where the men got a sore head and little else, of men who imposed on girls and got their love on a credit basis with no means of suing for the debt. She wasn't afraid of disease if she was carful and didn't get drunk. She objected to men taking similar precautions, for that reflected against her. What she feared was a run of wet nights when men refused to loiter and talk, when her clothing got wet. The good beats were around the leave centres of overseas troops. Tommy was passionate, but Tommy was poor.

``What will you do when you stop doing this?''

``I don't intend to stop.''

``But when you get older?''

``Stop asking questions. Ain't you happy?''

``Yes, but------''

She stayed his words with her lips. She refused to project herself into the future. The bed was warm, and a host of sun-tanned soldiers were marching down upon London. The war was a generous tide that brought a rich harvest on its bosom. The war was a continuing quantity. Peace was an impossible nightmare that might shut off the supply of men. And overseas soldiers unused to commercialized vice, and with a brief eight days, had not time to haggle over shillings.

``Soldiers won't always come.''

``I hope they come for years.''

``For years?'' He was amazed.

``For years and years and years.''

Was it surprising that war went on and on, that peace-makers were criminal or at the very least suspect? For war was a vested interest to all except the soldier. War was honourable. Peace was a thief in the night. The profiteer made excess profits on the shell that killed the soldier, the war correspondent on the phrase that told of his death. Brass Hats climbed to sublime heights of promotion and importance on war of attrition, honours falling to them according how they felled men. Statesmen fashioned glowing perorations, while soldiers sacrificed themselves in reeking slimes and foul ditches. The Sam Brown parson whose sermons had emptied the village church revenged himself on the scoffers and had them marched to his service in martial battalions. Society ladies found a new amusement called war work. Mothers who believed that they had made the great sacrifice by sacrificing their own sons, insisted that other sons of less sacrificial mothers rot in the same foul pit. The poor prostitute made a living out of the soldiers' loins. And she was maybe the only one who gave value, who healed by making men half human again. She was the lesser prostitute, for she only sold her body to live, but others with fat livings sold their minds. War was dug in. The whole populace had a financial interest in murder.

``You hope they come for years?''

``Why not?''

It was the prostitute who showed him in what a trap the soldier was caught.

``For years and years?''

``Why don't the police leave us alone? We harm no one.''

``You're a very moral person after all, Ann.'' He hugged her, still feeling very much the big brother.

``What funny things you say.''

She kissed him a few times and he forgot his moralist and big brother pose.

They slept together again, two soiled children, for he was soiled with trench rottenness, the foulness of a dead-house. And the killing beast was softened and humanized and rested by contact with the girl of the streets. He pillowed his head against hers, her hair brushed against his nostrils. He forgot she was a prostitute because he discovered she was a woman. Men are only selective when they have time and choice of many woman. She grudged him nothing, and brought him nearer to sanity. If every morning he was to be big brother and moralist, every evening he was to be a soldier and a man.

The blind was drawn when he roused himself again, and Ann was standing by the bed in a flimsy wrap smiling down at him. She was assiduously polishing her nails and looked so dainty that again he ceased being a moralist. She had slipped out to tidy herself, to smooth her hair, to powder her cheeks, to be as alluring as possible in the daylight. She had to make a good impression. She had her man. She was satisfied to want him for the duration of his leave.

``Hullo. I went to sleep again.''

``Happy? You had a good sleep.''

She bent over him to kiss him good morning a second time.

``Am I dreaming?''

The clean, fleecy bed, a girl with mischief in her eyes, the right to stretch, to yawn, to let sleep go gradually, all this was better than a violent awakening in a grubby billet, a rubbing of sleep from eyes with grimy hands. And he credited her with all the change because she was the most vital portion of it. He yawned loudly, letting his throat pronounce his sensual joy. He was refreshed as though he had slept an age. His nerves were soothed. He stretched a second time, and as she bent to kiss him he threw back the covers and drew her down, wrap and all. She was frail, delicate, if harsh and indelicate was her profession, no more to be judged on account of it than was a soldier with a fanatical bayonet.

``Quick, quick, let go, breakfast.'' She struggled free.

A girl came in with a tray, a girl void of embarrassment and of good looks, or she would have been in a bed instead of carrying a tray. She put the tray down and said a well-mannered ``Good morning,'' as though she catered for the most respectable people in Britain. There was coffee and eggs and toast for each for which he had paid the previous evening. He ordered at once a second plate of toast.

``A tip for the girl,'' Ann urged.

``You girls look after one another,'' he complied.

``Thank you very much.''

Together they ate and made great play of breakfast in bed. As they sat side by side a couple of other girls knocked a rat, tat, tat, on the door. ``Come in,'' he said, before Ann could answer.

They came to exchange greetings.

``Morning,'' he grinned.

``Late this morning,'' a prettier than Ann replied.

``Get out,'' said Ann. ``No poaching,'' reaching an arm around her man.

``I'm not a thief. Ta, ta.'' They went.

``Were they alone last night, Ann?''

``No. They are this morning, though.''

``Surely they're not after me.''

``No. But some girls would steal a man out of your bed.''

He was amused at the openness. Being a soldier he could sit up in bed and grin. He did not have to dress early and slink out to save a civil reputation. He jumped from the bed and went to the bath. It occurred to him that there had been a soldier in nearly every room of the three-floored house, a whole company of the line. But the house had been quiet. Girls came and went with a carelessness not attainable if the house had been given to civilians. The uniform killed the individual. Civilians had to think of neighbours, of employers, were compelled to make vice furtive. Soldiers could walk openly up the stairs and march bravely out in the morning, and in a few days even the girl who entertained them would not know them.

``Like your back soaped?'' She fired a shot at him as he went.

``Will it be extra?'' he bantered.

``Absolutely free.''

She scrubbed his back. He remained so long laving his body with the suds, caught up in his new adoration of the cult of cleanliness, that she came again.

``Thought you had fallen down the sink.''

When he was buttoned up she instantly had her arms around him, a pleasing armful on a grey morning.

``You'll stay. Your mate is staying. We'll have a good time.''

``He's not my mate, although he's a New Zealander. I only met him a few minutes before I met you. I haven't much money.''

``How much?''

``Eight pounds.''

``Enough if you let me spend it.''

``For how long?''

``Nearly a week.''

``We'll try for a few days.''

She was overjoyed. She knew where to get reasonable meals cheaply, where to see good shows at the lowest figure. The room wasn't too expensive. She wasn't an aristocratic whore, so was not too mercenary in her own demands.

``Are you sure I've got enough?''

She sealed the bargain with a hug, and playfully nibbled at his ear. Then she broke away to tell the maid that the linen would do for the night before he could change his mind.

He told her ashamedly where he had left most of his money, and she let him go to the Club to get it. She was not eager to let him out of her sight, but there was no option. She would tidy herself while he was away. He ran down the stairs and turned to take stock of the exact number and location of the house, for it was one of a confusing row. After collecting all except a pound he hurried back, growing confused about the number as he stood in front of the terrace. But he went confidently up the steps to the house he thought it was and knocked. A huge, course fellow came to the door, evidently the landlady's husband or lover, hefty enough to quell anything that looked like a rough house.

``Well?''

The hefty brute put his hand on his hips and glared. Guy wondered if the tough thought he had returned to recover stolen valuables, he glared so fiercely.

``I want Ann.''

He hadn't bothered about the second name.

``There's no Hann 'ere.''

``Oh yes, there is. I didn't sleep with a ghost.''

``Hann? Oo's Hann, anyway?''

``That's her. Ann who. I forget her second name.''

``Look 'ere, young fellow, take my hadvice and shove off. You've come to the wrong 'ouse.''

And Guy wasn't sure the big stiff wasn't right. He was doubtful about the name and couldn't swear as the the house. He felt like a damned fool.

``A---a---a------'' He searched for something to say.

``Hay what? 'Ook it!'' the door slammed violently.

`` `Ook it.'' He repeated the bludger's accents disgustedly to the reverberating door.

Daunted, he tried the next house. A demure young lady opened the door. Her presence, her wholly moral appearance, silenced him. How could he question her?

``A---a---a------''

She waited with her hand on the knob, smiling. He didn't know she was a street-walker too. She was so much like a Sunday school teacher.

``My mistake. Wrong house.''

``We don't mind paying guests.''

She laughed at his confusion, but he hurried down the steps and again the door slammed on his back. He stood on the kerb for a while and saw  the coarse fellow peering at him from behind a curtain. After a little while the door opened and Ann herself appeared. She was surprised and overjoyed to see him.

``Why didn't you come in? I thought you had slipped me up.''

``I knocked at the door and a big swine told me to 'ook it.''

``Hush,'' said Ann. ``Hush. Don't call him a swine. Come in.''

The big coarse swine was in the passage as they tripped upstairs.

``Why didn't you say? Ow did I know?''

``Go to hell, you stupid cow. Do you think I want to spend my leave standing in the gutter?'' There was more trench than culture in the remonstrance.

``Hush! Hush! Hush! Be quiet!'' Ann knew the swine had a winning set of muscles.

But he was slow to anger. It paid to pacify the girl's clientèle.

``Sorry, New Zealander. Sorry.''

Ann got herself ready and they went into the streets. The sun was shining. Jauntily as picnickers they stepped along the pavement. All the world could see. He didn't care nor did the girl. A soldier with a whore. All the world was busy making profits and forgave the soldier as it despised the whore. Her profits were too slight to be honourable. He was on leave and he was young and the war was going on for ever. She was as blithe as a sparrow and nearly as simple, and her comment on the passing show was spicy. If she had cracked all the same jokes a hundred times before, to him their content and their accent came as stimulating novelty. He bought two fourpenny apples and they munched as they perambulated the streets. He had sloughed off the years and was a child. They even bit at one another's apples. People could approve or disapprove. He was more of a boy than a sinner and Ann was more of a girl than a whore.

Wealthy folk may yet get bored in marble palaces amid beautifully gowned and exquisitely perfumed courtesans. They need a little hunger, a little continence, enforced abstinence and suffering, to restore appreciation of simple values. Not ``is she beautiful'' counts with a soldier, but ``is she a woman,'' for all women become beautiful, magdalen or virgin. Soldiers are not civilians so belong to no scheme of things. They forage by the way. And boys out of hell manage to be grateful and not contemptuous to daughters of the streets.

He passed dozens of officers but only saluted when they were police officers. To hell with salutation. He stared defiance when they looked questioningly to assess his intention so as to be ready with the courteous response. And some few officers were hungry for subservience if the fellows turned the head away. Ann made good and exaggerated the omissions.

``Don't you salute?''

``They can go to hell.''

``Leave it to me.''

She saluted with offensive elaboration, laughing when an officer grinned and nodded, putting her fingers to her nose or poking out her tongue when the officer scowled.

``Wouldn't sleep with him. He'd remind me of his high rank every few minutes.''

She was great company. Perhaps an infantry private and a Flossie had a lot in common for the body of both was in pawn.

