\chapter*{\textsf{Eve of offensive}}
\addcontentsline{toc}{chapter}{Eve of offensive}

H\textsc{e} heard the wail as the heavy shell high up sighed its way toward Warneton. He saw a huge upward splash of smoke and powdered earth in Warneton as though a breaker had hit an iron shore sending spray flying. But this wave was of steel and bit remorselessly at brick walls, at slate roofs, at terraces of houses. Between la Bass\'{e}e Ville and where he stood he could see the sugar refinery into which New Zealanders were shortly to penetrate, and shells were falling there too, spurting black shadows that silhouetted like trees against the skyline.

``Warneton is getting hell.''

``Is it?'' The man who replied was uninterested, his ``is it?'' was merely a courtesy.

``Yes, hell!''

For John Guy, possessor of a fine pair of German binoculars, found Warneton detailed enough to be interesting. He observed from a post at the top of a support trench where he could lounge forward in comfort. Below him a comrade cleaned a rifle.

OOOO------ooo! would wail a shell in the air overhead and the earthen spray would dash up angrily in his vision. And when the spray had blown aside or had subsided to earth the shell would be seen to have obliterated a tower, to have destroyed the corner of a house, to have smashed in a roof, to have spreadeagled a poplar. Steel breakers were greedy in their bite. A permanent army order commanded other ranks to surrender looted German binoculars for the benefit of officers, but the order was a dead letter in the New Zealand Division. Officers were human in their attitude to rankers' souvenirs.

``So Warneton's getting hell.'' The rifle-cleaner repeated the words automatically, wholly devoting his attention to the rifle.

Guy didn't reply. He was feasting hungry eyes. What a world a pair of binoculars opened up. Looking dead to the front he could see Warneton falling asunder. Turning to the right and looking across the river Lys he could examine the low, flat land still in the possession of the Hun. The flat land across the river was near enough to be searched with the naked eye but the binoculars made it intimate. It was unshelled and green, knee-deep in grass and wild flower, a haven for eyes tired of the churned brown of shell-holes. The extraordinary contrast of grass and brown earth on opposite banks of the river evoked articulation.

``Brown earth and shell-holes on this side of the Lys. Luscious pasturage on the other.''

``Eh? What's that?'' asked the man at his feet, sliding his bolt and ejecting the yellow cartridge.

Guy didn't reply. He moved the binoculars in a traversing sweep piercing the blanket haze. He saw clean houses with smokeless chimneys, frail houses that the concussion of high explosive could hammer down were it not that the frail white house had become but a mask for a strong steel and concrete kernel. On one side of the Lys death swallowed houses and shrubs and roads and grass and men and earthen trenches, reducing all to vile chaos. On the other side the country was serene and beautiful. It didn't matter that serenity was a mask, beauty an illusion, that down on the flat as on the opposite bank men came and went on lethal purpose. His careful searching failed to pierce the mask and for a moment he contrasted the banks and toyed with illusion. He thrust aside the fact that life would shoot and mow down khaki life if the river was crossed, that men in grey would magically appear amid grass meadow and wild flower practising trigger pressure, that the machine-guns would spit from the whitewashed houses and barns. It pleased him to toy whit the idea of running away from war to holiday. But death stood alert in the meadows if death was banished from sight. And death could splash green grass crimson.

In this support line he could stand right above the parapet and search the low country. The enemy on the flat was too far away to snipe and used no machine-gun in the day-time, for British eyes observed for flat land targets. After roughly sweeping the whole country he set out to examine it in detail. The grass was rank, the wild flowers very tall, the hedges extravagantly thick and unkempt, uncurbed by the trimmer's slasher. The whitewashed houses were as cleanly to the distant eye as they must have been on the day when war swept down, compelling inhabitants to crowd trumpery possessions on to hand-carts, wheel-barrows, perambulators, sometimes into horse-drawn vehicles as they all hurried away before the sweeping tide.

On one side of the river was unnatural silence, on the other unnatural violence. Death on each side garbed differently. On one side death spreadeagled, on the other embalmed. On one side a frozen stillness, a death without rottenness, a country transfixed in immobility. On the other a chaotic rottenness foul to nostrils, foul to every sense. On one side chimneys and no fireside smoke drifting away, on the other the angry clouds of black or white smoke ascending or drifting away from shattered house foundations. There were shell-holes and not a vagabond blade of grass for miles, across the canal lush green and no brute to crop a mouthful. There were stumps of trees scarred of trunk, shorn of limbs and leaf, trees like stumps of decayed teeth in a battered mouth, shattered remnants of plantation, and across the way apple trees in blossom and fruit, green poplars. As he swept the glasses from bank to bank his mind travelled from pre-war to explosion. Across the river life had been hushed by invasion, where he stood it perished catastrophically. Yet in his chaotic world he could see cheerful, grimy men, in the enchanted land across the way stealthily men crouched in lairs. Cannon fodder was noisy amid rottenness, silent and whispering amid wild flower.

A patrol or two had ventured across the canal and had vanished, lost in the unruffled stillness, and on the following day green grass had appeared just as green and deserted wild flower just as lovely. The white walls of houses had swallowed up or devoured piecemeal flesh and blood bodies. The country was still serene even when a bloodied and terrorized survivor of a party scrambled back to bear frenzied witness to meadow stealth. Some said the Germans had a secret tunnel through which they could come beneath the river to raid and patrol, through which they might escape with the prized booty of a bloody and groaning prisoner. Naked, unashamed violence was merged into lethal stealth.

``Give us a look.''

The man at Guy's side, finding Guy's silence more impressive than his utterance, laid down the rifle and reached for the glasses which Guy surrendered regretfully. His companion searched Warneton in front and then turned to the east bank.

``It's devastated. But there must be a lot of trenches behind the hedges.''

Binoculars gave Guy eyes, extended the range of his interests, were a toy that made trenches less dreary. Even when a policy demanded that no one look across a trench top, he could look up at swarming planes, at observing balloons. Death extended into all dimensions. Miles to the front, miles to the rear, miles to the flank, miles into the heavens he could look and everywhere was death, organized and being organized. Beneath the very feet miners might be burrowing. Binoculars took their owner out of his ditch while he peered, but everywhere the eye ran into death. The stillness of deserted farmlands made him sad for the plight of common men who had to abandon clean grass to sit in charnel filth amid bursting shells.

``There's going to be an attack from here to Ypres.''

Guy nodded.

``Our Battalion's going to the sugar refinery in front.''

Guy nodded.

``Get hell on the flank there. We'll get shell and machine-gun fire from the quiet sector alongside.''

``But we're just going to jump about half a mile. Just one quick wave.''

``Do you know, I'm as scared as hell. I wish I could run away. My nerves are gone.'' There were four casualty stripes on the arm.

``No wonder. I'm all in and I haven't been hit.''

``But I haven't been nearly as much in the line as you. I get hit at once every time. Got shrapnel in each thigh, a scalp wound, a peppered back. Last time I got blown up and was silly for a week. Now I'm waiting for a fresh dose. I can't sleep. I get more scared every time I come back.''

``Why don't you run away?''

``Too scared to do that, too.''

``You won't get hit again. Someone else's turn.''

``I know all about that. I've been waiting for it since my return. If I could sleep.''

Ah, yes, if he could sleep. Lying against a breastwork in the cold air, his fear subtly communicated itself to Guy, who had enough to carry already. Guy shivered and felt a chill down his spine. The sky out of which shells came was very broad and his back seemed instantly very frail.

``I'm going down.''

He moved to a deep, filthy dug-out where a dozen men were breathing vile air, air densely clouded with the smoke of cigarettes.

``There won't be a wheel-barrowful of La Bass\'{e}e Ville to-morrow.''

``Enough to spreadeagle a good few of us.''

He lounged on a damp floor and dragged a sheaf of leaflets out of his equipment. Because his company was not going across he had retained his pack. What a funny collection he had again. He took up a Fabian pamphlet and grinned at it. ``Socialism for Millionaires,'' by G.B.S. He put it away. The incongruity of that leaflet in a soldier's pack was what kept it there. It had no place, so perversely it was retained to read at some time. ``When Peace Comes.'' He grinned again at a reconstruction leaflet he had gathered in a moment of optimism. Peace wasn't coming and millionaires didn't want to be Socialists. ``Admirable literary tastes,'' he confided to himself as he lit an issue cigarette. \emph{White Fang}, a filthy copy he had found in a dug-out, as next to hand. ``Too long to be read straight away. After next stunt, perhaps. Rotten to be killed half-way through a book.'' Two other leaflets told him how to replan the world if he lived, but none told him how to live. He fell back on \emph{Mirror of the Sea}, which took him away for a time from the filthy dug-out. The fresh salt winds rattled about his ears. In a sentimental mood he opened his breast-pocket and looked at the yellow ringlet within the covers of \emph{The Rub\'{a}`iy\'{a}t}. And then big shells started to pitch into the soft country above, so he abandoned literature for nicotine.

He emptied his pack of issue cigarettes. Ten packets he had saved, for he smoked only in the line. He settled down and smoked cigarette after cigarette, packet after packet, until his senses reeled and his tongue grew sore, absorbed smoke so that his nerves grew too dull to wince at concussion. The heavy shells fell around the old German dug-out that the German gunners now knew to be inhabited by New Zealanders. Heavy shells that slid a long way into soft earth on each impact before exploding, impacts that caused the dug-out to oscillate. But such dug-outs were not easy to reach. A furious bombardment of British back areas grew with evening, Fritz being anxious to render attacking concentration impossible. Confused with the narcotic, Guy found his nervous responses casual.

``Going up to relieve myself,'' his comrade with the four wound stripes announced. ``Don't want to go, but------''

``You're all right. Die in bed. Tell the kids what you did in the Great War, daddy.''

But as Guy reached for another cigarette he heard an intimate shriek and crash as a shell burst around the dug-out entrance. The voice of the gas guard called down the shaft.

``Stretcher-bearers.''

The bearers went, but the fifth stroke had been the last. They collected the pay-book and a disc, and threw the corpse over into a shell-hole to await a burial party. Guy heard them joking as they came down.

``Would have been a heavy swine to carry out. Thirteen stone easily.''

``Shell-hole's as good as a cemetery. He won't know about his new address.''

``Told me coming in he felt he was for it.''

``Fritz couldn't miss him.''

Guy smoked and smoked, and some of the words remained in his mind. ``Heavy b------ to carry out.'' The tone had been comradely. He marvelled at a war which shot four times at a man and then dropped him finally, that kept another permanently racked. How many times more before he got his? Before some humorist would find ``Socialist for Millionaires'' and ``When Peace Comes'' in his pack. What laughter there would be. And, Jesus, it was funny. But he didn't want anybody to find and laugh at his golden ringlet of hair.



