\chapter*{\textsf{Calm}}
\addcontentsline{toc}{chapter}{Calm}

I\textsc{n} the morning equipment was overhauled again for the last time, although the final inspection was lax. Men were going into battle, not on to parade, and were trusted to adhere to instructions. Ground-flares, emergency rations, a second set of bandages and iodine, picks and shovels, an extra bandolier of ammunition, the weight of martial impedimenta grew until it threatened to crush mobility. Trenching tools were slung across the crutch instead of the backside, special equipment was apportioned.

``Here,'' said the signalling Sergeant, handing Guy a basket of pigeons. ``Strap those on your back. Don't cook 'em and eat 'em.''

``Righto.''

``Takes a bloody lot to kill.''

``Takes more to bloody well keep alive. Poor shoveling is more deadly than poor shooting.''

None, not even the most overweighted, objected to the shovel. All knew the value of earthwork.

``Here,'' said the Sergeant, handing out a couple of Mills bombs per man.

He had parted with all his treasures except two books. For a time he had carried a miniature library. He retained his two favourites in his fighting kit. Absurd sentiment to overburden himself in this way. He would never read under fire. He had a copy of Conrad's \textit{Mirror of the Sea} he had stolen from the Y.M.C.A. library, a copy grimed but until his theft unread, because its pages had been uncut, a book out of whose pages had blown the salt tang of the sea to refresh nostrils accustomed to Flanders filth. The volume was a friend, a talisman, a reminder of imaginative ecstasy that civil life had once permitted. And there was a copy of Omar with a girl's photographs and a lock of golden hair. He never told the Omar crank that he had a copy in his own pocket. He was afraid Joe might want to look at it and laugh at its photos and its lock of hair. For he posed as something of a realist and not as a romantic. He often touched the lock of golden hair that was gummed against a verse that told of the perishability of all things but woman and wine. He sneered at his sentiments but he never threw the book away. For deep down underneath, so deep that none could guess, so deep that at times he was unaware of its continuity, he treasured the memory. Perhaps one day a miracle might occur and he might go marching back, sullied a little, but back. He sneered at his sentiment but carried the ringlet.

``Company will parade for final inspection and instruction.''

Company did parade, not in line, but loosely round the Captain, and the company listened to final instructions which were merely a repetition of all other instructions. And the parade was unusual, for the Captain and his Lieutenants had suddenly become as his men, were dressed in ranker's uniform and fittings, and off their martial pedestal were fellows in arms after all. Fine feathers were good things to swagger about in, but prosier garb was reputed to have a higher survival value. If in bygone centuries the most gaily caparisoned knight led his forces in person and valour, in modern warfare the prettiest plumes are farthest in rear, deepest in the dug-outs. Men grovelling around in mud had to achieve protective colouration. An officer showed common vulnerability by wearing a batman's tunic.

They stood around listening as a gathering of idle loungers might listen to a street-corner speaker. John Guy acquired a new interest in his Captain, who seemed suddenly human, a fellow-sufferer, no longer an instrument of militarism for militarism's sake. Stripped of his uniform, he was just a man and not very lucid. His tones were halting, his oratory dross.

``Any questions to ask?'' Silence greeted the invitation. ``Well, good luck. We can't all survive. I know the company will do its job whatever happens. Dismiss the parade, Sergeant. No. Wait a minute. Let me repeat one instruction. No one must leave the line to help carry away a wounded man. That is the bearers' job. If we all got tender-hearted there'd be no one left to man the line.''

Common sense, Sir. The objective of offensive was not avoidance of death but possesion of territory.

``Parade, dismiss. Don't go away.''

Parade shuffled away. Men hung around camp in the hot morning and afternoon sun. Day dragged away. Money had been spent, or civilians and wine were no more. Organization fretted for zero. A few men sat in the Y.M.C.A. hut writing a postscript to farewell letters. A few might have gathered to be fortified again with prayer. Devout Catholics received the sacred rites of the Church.

At early evening the cooks ladled out the ration of stew, and the water-cart offered a last chance to fill bottles. Conversation was fragmentary. Over all was a fitful brooding. Addresses were again exchanged by men who wanted comrades to write a letter in the event of death. Death would cut with a broad blade.

``Will you write home if I die?'' a dozen asked Guy.

``Righto.'' A request was never refused.

Nothing in life could be planned beyond a letter to the bereaved.

Messines was a barrier to the future.

``Will you write a letter for me?'' the Salvationist asked.

``Why not the Padre?'' Guy grinned as he assented. ``Sorry I was a bit fresh last night.''

``The Padre will. But I want a note from a fellow who is going over the top.''

``But to-morrow you'll be in heaven and I'll be where notepaper scorches.''

``Well, you don't need to write if you end in hell,'' said the Salvationist genially.

``Sure, I'll write all right.''

``Do you want me to write for you?''

``I'm born to die in bed.''

``You look more like death by violence than any fellow I know.''

``Go on. I'm yellow.''

``Forget last night. And we are friends?''

``Of course we are friends. Friendly verbal enemies. We'll continue the argument after Messines.''

``After Messines.'' The tone of the voice seemed heavy with fate. ``Something tells me there'll be no after Messines for me. I feel Messines has gone the number of my mess.''

``Nonsense.''

``No, really. I do.''

``How dreary the day is.''

The air was hot, heavy, and men hung around idly.

``The sooner we get it over the better.''

``That's how I feel.''

``Did you ever wait outside a dentist's?''

``Did you ever stand at the door of hell?'' Guy concluded with a grim jaw.

``I like you, Guy.''

Guy shook hands and walked away.

He liked the Salvationist.
