\chapter*{\textsf{Insubordination}}
\addcontentsline{toc}{chapter}{Insubordination}

R\textsc{eveille} sounded its beautiful call. The morning was cold, heavy with fog. 
The lovely notes were obscenely greeted.

``Damn the bloody reveille,'' said a drowsy voice.

``Hock the Kaiser,'' a second replied, thrusting back the blankets.

``B-last the Kaiser,'' said the first, exploding upward out of bed.

Soldiers stirred, cursed, rolled on the broad of their backs, wrestled with the 
desire to linger on and the need to jump. More and more burst upward like 
high explosive, reaching for the trousers.

``Time to milk the cows.'' A dairy farmer threw a phrase at the hutment.

Phrase after phrase was thus thrown into the hut, not to be listened to 
but to assert wakefulness. That raw morning last moments filched from 
war gods had flavour. But man after man caught at his towel and was off 
to the ablution shed.

Someone always lingered. John Guy lingered that morning. He had a mild 
chill. His chest was sore from an old weakness. He sniffed at frosty air 
as it drifted through the open door. He listened to hear if the sore spot 
on his chest crackled. No. It was just a soreness. He was always afraid 
of that old sore. There were enough bullets in France without taking 
a weak body along. Generally he was early in the race for the ablution 
shed. He was aware of the risk of lying on. Some disciplinarian might 
pounce. But sore of chest, he revelled dangerously.

Tick! Tick! Tick! Tick!

He watched the second hand of his watch jumping. He would luxuriate until 
the last moment and then with haste be ready for physical drill parade. 
But he heard another sound and knew himself for a fool. The front door 
of the hutment slammed in and Angley bounced across the threshold 
accompanied by the Orderly Corporal. John Guy jumped. As he bent for his 
trousers a stinging cane cut at his elevated rear.

``Serves me right,'' he thought as he lifted a leg to the trousers.

A second sideward swipe left another red weal.

``Drop them! Drop them! Drop them!'' Angley roared.

``Stand to attention.''

Men were hurrying back from the ablution hut. Angley aimed another cut at 
the shivering victim. ``Swish,'' went the cane. ``Ouch,'' said Guy.

All smiled if all hated Angley. A shirt-tailed, shivering victim defending 
his buttocks from a swinging cane is not a dignified sight.

``We've met before,'' announced Angley.

``Yes, Sir.''

There came another weal.

``Turn that up.'' Guy caught at the cane, foolish gesture, longing to hit 
Angley on the jaw.

``Ho! Ho! Turn it up, eh! Who the hell are you to talk to me like that?''

And Guy knew he had annoyed the bully.

``Turn it up. The private gives advice.''

The air circulated between the doors. Guy shivered and bent for his trousers 
as the raw air played around his body.

``Stand to attention!'' Angley headed off the attempt. ``Put your heels together, 
thumbs behind the seam of your shirt.'' Angley beatifically contemplated his 
misery. ``Having a little beauty sleep, eh? Thought you were at the Savoy. 
Well, Mr. Private, I'm sorry to disturb you. Just called in to take your 
breakfast order.''

Someone in the hut tittered. Angley was delighted. He loved public recognition. 
The call to physical drill sounded.

``Tell the instructor this man won't be on parade, Corporal. Say he's staying 
late for beauty sleep.''

Angley loved an audience for his wit, but preferred solitude for his brutality. 
Witnesses were discomforting. They might be embarrassing at some time. The hutment 
had emptied in double time, for Angley was present.

``To attention, I said!''

Guy was so cold and numb that his teeth chattered.

``Ah, I didn't want you to go out in the cold.''

Teeth chattered.

``Stop making a noise on parade.''

Teeth chattered.

``Swish,'' went the cane viciously.

``Stop making a noise,'' Angley yelled.

Guy tried to clench his teeth, but that only made the chattering louder.

``Your name?''

Angley knew his name. Twice on parade he had asked him that. Once 
because a suggestion of a sneer was on his face.

``John Guy, Sir.'' The victim was sure he was getting a heavy chill.

``When did you arrive here?''

``Last reinforcement, Sir.''

``Last reinforcement?'' Amazement in the R.S.M.s voice.

``Do you read the newspapers?''

``Yes, Sir.''

``You know what year this is?''

``1915, Sir.''

``Took you a bloody long time to discover there was a war on.''

The Orderly Corporal tittered, Angley glared maliciously. John Guy's 
teeth chattered. He gave no answer until the swinging cane cut at his 
numb legs.

``Yes, Sir.''

``Bit of a rotter to leave your country undefended so long.''

The cane was swinging. There was no escape. John Guy had heard others 
put through a similar catechism. But it took another cut to knock the 
appropriate answer out of him.

``Yes, Sir.''

``You think war is a holiday?''

``No, Sir.''

``If the Hun catches you in your shirt-tail he won't use a cane.''

``No, Sir,'' said Guy, thinking, ``THE HUN has caught me.''

``I think we'll have a little exercise.''

The victim was glad of anything that would make the blood flow.

``Lie down. Cover yourself up. Pleasant dreams, little one.'' Angley 
had a technique.

Guy did as he was bidden.

``Get up!'' Angley shouted.

Guy jumped.

``Lie down.''

He lay down.

``Up---down. Up---down. Up---down. Up---down.''

Guy jumped until his head reeled, aided by odd swipes with the cane. What 
would he not have given to swing a right-hander at the swine.

``Hurry, you cur! Jump, you rotter! Put a move on, you bloody slacker! I'll 
wake you up, you sleepy swine! Call yourself a soldier!''

Guy was more used to the invective than the cane. And Angley was 
God. So Guy did as he was bidden. When men came doubling back from physical 
drill Angley was satisfied. The culprit had spent half an hour drilling 
in his shirt-tail.

``Report to me at the orderly room at 12.30.''

``Yes, Sir.''

He aimed a parting blow and hurried for the door, but as he stepped out 
he was inspired. He turned and Guy was pulling on his trousers. One leg was in. 

``SHUN!'' Angley yelled. ``SHUN! Stand up. Stand up. Stand up. Let go of those 
trousers on parade. Let them go. Let them go!''

Guy let them go and they hung around one ankle.

``Mark time! Double Mark Time!''

The hut rocked uproariously at the incongruity.

``Parade, Halt! Parade, Dismiss!''

``Angley's a rotten swine but you did look funny doubling with your leg through 
your trousers.''

``Angley's so white he won't be able to wear trousers in France.''

Guy made the best retort he could. He was sure he had a chill and he had 
a few choice red weals on his buttocks.

``It would test your golden smile, Robbie.''

``Why did you get caught lying in?''

``Why the hell does anyone get caught?''

Before parade was dismissed at midday each man was given a new rifle and 
the old issue were collected. The rifles were filthy.

``Get them cleaned for afternoon parade.''

Everyone cursed at the encroachment upon time. The rifles had been used 
on the range and pulled through hurriedly by a departing reinforcement. 
The barrels had sweated out their filth. Even the outsides were dirty. 
Guy spent most of his hour outside the orderly room awaiting Angley's 
pleasure. There he was lectured.

``You know you are a slacker?''

``Yes, Sir.''

And so on. Not very viciously for the sun was shining, lunch had been 
good, alcohol had not lost its glow.

``You may go but don't let me catch you again.''

``Yes, Sir.''

``Fall in,'' said the Sergeant as Guy reached the hutment.

``Dirty rifle. Take his name, Corporal,'' said the Sergeant-Major who 
examined the squad.

``But I hadn't time,'' Guy argued.

``Everyone else had no time. You've made no effort.''

``But, Sir------''

``Don't `but' at me on parade. You're not a Billy-goat.''

``I had to go to the orderly room, Sir.''

``Um, a frequent offender.''

``I'm not getting a fair go.'' Guy tried to explain the injustice.

``Another word and I'll get you for insubordination.''

``I only want a fair spin, Sir.''

``We'll give you a fair spin. Colonel, put down insubordination as well.''

Guy's mouth opened to try again.

``Another syllable and it'll be the guard-room.''

Guy though it a bit thick, but the more he tried to explain the 
worse it became. He was enraged at his powerlessness.

``Hard luck,'' Robbie consoled when the Sergeant-Major moved to 
another squad.

``Shut up!'' yelled the Corporal. Already he was trying to snarl 
like Angley.