\chapter*{\textsf{Pagan death but Christian burial}}
\addcontentsline{toc}{chapter}{Pagan death but Christian burial}

H\textsc{e} went on fatigue for a moment, preparing for fresh offensives, improving communications to captured lines, carrying rations, material, munitions to infantrymen in forward positions. He helped to comb the country between trenches for the corpses of past offensives when night came down. The parsons had to ``do their stuff'' so the country was very carefully scoured. The fatigue parties took identity discs and pay-books from month-old corpses, putrefying and crawling remains, and deepened a shell-hole while the parson hurried along to cloak the odour of decay with the aroma of sanctity. The corpses were scarcely touched by rats for the rats were only gradually eating their way forward.

The whole area was foul with bodies, bodies swollen and bursting with the heat and the rain, bodies making sizzling sound of flesh alive in death. And since every man had to be accorded Christian burial the parson mumbled his words from behind a perfumed handkerchief and could not be blamed for that. For Christian death disgusted earthly membranes.

What a satire the last rites when everything human except the clothing had gone from the body, when the body was more unclean than the clay to which it was returning. Love and comradeliness and the will to obscenity had been battered to earth weeks before. Sight and taste and touch and voice and, mercifully, power to smell, for what wretch wanted to linger for a month in the odour of his own heroic decay, had departed on a fateful day. So now came the mocking of Christian burial for Bill Jones when Bill Jones had been blasted by a Christian gun. Christian burial when man made in God's image was become a pent-house of squirming life. Christian burial because the \textit{Army Manual} ordained that the parson must have his warm fire in a way-back billet to perform what was distasteful to all. A plague on this Christian burial. Was Guy the only one who thought that?

Such loveliness as existed in burial existed when dead men still wore outward semblance of living men, existed before Christian guns warped and blasted their bodies, existed when the corpse seemed of a comrade. But what loveliness was there in a parson muttering from behind a perfumed handkerchief while all felt nauseated by sizzling putrefaction? Christian burial was indeed blasphemy. The Sam Brown parson was always sure that the patriotic soul, the vital flame long departed from the rotting body, had been blown from earth to heaven, elevated to serenity by cordite. Fortunately the parson could never get the service over too quickly. Shovels poised loaded with clean earth ready to blot out the rottenness of honour and glory the moment he stayed his mumblings. Fatigues never lingered around the location of Christian burial.

What a task for a new reinforcement was the combing of an old battlefield, the burying of such dead. It took such an experience to make honour and glory as nauseating to the soldier stomach as it was uplifting to the civilian soul. Newcomers recoiled outraged by the savage and careless obscenity of men who had been long enough in France to make a joke of life.

``If they dragged this corpse up the main street the maggots it would drop would upset the Lloyd Georges.''

``For Christ's sake shut your mouth,'' said a man who hadn't yet realized that obscenity helped to retain sanity.

``If everyone could get a good chestful of glory.''

``Sh------sh,'' the parson started.

``The Lord giveth and the Lord taketh away,'' was that filtering through the handkerchief.

``In New Zealand they fancy each fellow gets buried rolled in a Union Jack.''

``Yes. Last Post, military salute, white cross.''

``White crosses in the sun,'' said Guy, thinking of the sad, beautiful forest of crosses in Etaples. ``White cross and no rat burrows.''

``Judging from the smell this one will have to be disinfected in heaven.''

``Quiet on that,'' from a shocked voice.

``Hell is the place for a good soldier. The fire will cleanse.''

``But he's too green to burn easily,'' Guy muttered.

``For Christ's sake shut up. I think some of you fellows enjoy this job.''

``There are five Huns on the right. Dead ripe. Do you want to read the service over them?''

Some padres did. Padres who were not average.

``Got enough to do to-night.'' The parson practiced ca-canny.

``Come on,'' the Sergeant commanded to Guy and another. ``What Hun needs a gentleman's burial anyway?'' the Sergeant was another who saw little merit in Christian burial for month-old corpses.

A shovel dropped a clod on a body so moist that the clod splashed.

``Christ! he smells like a Hun.''

``Another too green to burn, and all Huns are for hell.''

``Shut up,'' said the Sergeant.

``Cook first. Burn afterwards.''

``Shut up, for Jesus' sake.''

``Why burial by parsons anyway, Sarge?''

``Military orders.''

``Wish the Kaiser had to sleep with this one.''

Joking instead of brooding, men clung to the remains of their sanity as they combed the shell-holes. If they had paused to say ``This might have been me,'' they could not have continued soldiers. Yet the odour of Christian burial could persist in the mind of a sick man for many a week, for many a month, something to crowd the silences. The memory could even start him wondering whether it wouldn't be better to put a bullet in his own brain and get buried in clean earth was not without fascination, though why he should bother about the sort of corpse he would be when he was dead was something that did not occur to him. Sick minds catch at the inconsequent details.

He didn't want a Sam Brown parson to mumble indecencies over his bloated and rotten remains. Rats would eat one if one was buried outside, but rats avoided such filth as he witnessed nightly. The soldier, after all, owed a debt to the rat. Was it outrageous to take one's own life, more outrageous than to kill? If one could only cheat the war gods without torturing one's comrades. That was it. Without torturing one's comrades. Suicide was escape, but not for the fellows who had to bury the corpse. Still he must cheat the Brass Hats, the Padres, the Maggots, the squirming  life that crawled over heroes.

``Huns. Phew!''

``All smell the same now. Only tell 'em by the uniform.''

``Get a good rum ration to-night.''

``Is the parson burying the soldier or is the parson burying the livestock?''

His mind pondered over such problems. He was becoming morbidly instead of boisterously obscene. He was growing serious and seriousness is always dangerous in a mad-house.

``Phew!''

``Yes. Pretty high. A General couldn't do better.''

``Shut up.''

``They can do everything to a soldier except put him in the family way.''


